### 2014-06-04 paper/paper.tex
starting to work on the paper... first day's worth of work

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass{sig-alternate-10pt}
\usepackage{endnotes,url}
%\usepackage{sig-alternate-10pt,endnotes}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{verbatim}
% Replaced by subfig package
% \usepackage{subfigure}
\usepackage{subfig}
\usepackage{epstopdf}
\usepackage{cite}
\usepackage{array}
\usepackage{multirow}
\usepackage{listings}

%\pdfpagewidth 8.5in
%\pdfpageheight 11in
%\baselineskip=12.5pt
%\setlength{\pdfpagewidth}{8.5in}
%\setlength{\pdfpageheight}{11in}
%\usepackage{pdfdraftcopy}

% To import classical math notation used
\input{mathnotation.tex}
% Package for date and time presentation
\usepackage{datetime}
\newcommand{\datation}{\today, \xxivtime}

\def\full{0}        % set 1 for a full tech report version
                    % set 0 for submission version
\def\shownotes{1}   % set 1 for version with author notes
                    % set 0 for no notes
\def\anon{1}        % set 1 to anonymize
                    % set 0 for acks and author names

%%%%%%%  Author Notes %%%%%%%
%
\ifnum\shownotes=1
\newcommand{\authnote}[2]{{ $\ll$\textsf{\footnotesize #1 notes: #2}$\gg$}}
\else
\newcommand{\authnote}[2]{}
\fi
\newcommand{\Ynote}[1]{{\authnote{Yan}{#1}}}
\newcommand{\Anote}[1]{{\authnote{Andrius}{#1}}}
\newcommand{\Nnote}[1]{{\authnote{Narseo}{#1}}}
\newcommand{\Vnote}[1]{{\authnote{Vijay}{#1}}}
\newcommand{\Dnote}[1]{{\authnote{Dina}{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\namedref}[2]{#1~\ref{#2}}
\newcommand{\tableref}[1]{\namedref{Table}{#1}}
\newcommand{\sectionref}[1]{\namedref{Section}{#1}}
\newcommand{\appendixref}[1]{\namedref{Appendix}{#1}}
\newcommand{\theoremref}[1]{\namedref{Theorem}{#1}}
\newcommand{\remarkref}[1]{\namedref{Remark}{#1}}
\newcommand{\definitionref}[1]{\namedref{Definition}{#1}}
\newcommand{\figureref}[1]{\namedref{Figure}{#1}}
\newcommand{\lemmaref}[1]{\namedref{Lemma}{#1}}
\newcommand{\claimref}[1]{\namedref{Claim}{#1}}
\newcommand{\propositionref}[1]{\namedref{Proposition}{#1}}
\newcommand{\constructionref}[1]{\namedref{Construction}{#1}}
\newcommand{\corollaryref}[1]{\namedref{Corollary}{#1}}
\newcommand{\equationref}[1]{\namedref{Equation}{#1}}
%
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{obs}[theorem]{Observation}
%

\providecommand{\vs}{vs. }
\providecommand{\ie}{\emph{i.e.,} }
\providecommand{\eg}{\emph{e.g.,} }
\providecommand{\cf}{\emph{cf.,} }
\providecommand{\resp}{\emph{resp.,} }
\providecommand{\etal}{\emph{et al.}}   %Removed trailing space here; usually want non-breaking space with following reference
\providecommand{\etc}{\emph{etc.}}      % No trailing space here either
\providecommand{\mypara}[1]{\smallskip\noindent\emph{#1} }
\providecommand{\myparab}[1]{\smallskip\noindent\textbf{#1} }
\providecommand{\myparasc}[1]{\smallskip\noindent\textsc{#1} }
\providecommand{\para}{\smallskip\noindent}

\newtheorem{axiom}{{\bf  Axiom}}
\newtheorem{defin}{{\bf  Definition}}
\newtheorem{proposition}{Proposition}

\usepackage{enumitem}
\setlist{nolistsep}

%Model parameters

\usepackage[breaklinks=true]{hyperref}
\frenchspacing
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{noTCP: do not extend, repurpose}
\ifnum\anon=1
\author{[Paper: \#XXX]}% \hspace{0.2cm} \ampmtime ]}
\else
\numberofauthors{2}
\author{
\alignauthor Andrius Aucinas\\
\affaddr{University of Cambridge} 
\and
\alignauthor Jon Crowcroft\\
\affaddr{University of Cambridge}
}
\fi
%for single author (just remove % characters)

% end author
\maketitle
% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}
\begin{abstract}
Too abstract.
\end{abstract}

\section{Introduction}
\label{section:intro}

Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large. The problem primarily arises from the various middleboxes that perform packet content processing withing the network to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems and is already well documented~\cite{UntoldMiddlebox2011,Qian:2012bj,Honda:2011ci,Guha:2005tb}.

The lack of extensibility has become acutely noticeable with the development of TCP extensions such as MPTCP and TCPCrypt, the former to make use of multi-homed endpoints, particularly useful in wireless networks and the latter to provide ubiquitous end-to-end traffic encryption. These and other extensions are primarily designed around the limitations of existing middleboxes, but they demonstrate two problems: 1) any extension design is extremely constrained by prevalent middlebox behaviors\cite{Honda:2011ci} and 2) we have already run out of TCP option space, the built-in mechanism for extending TCP.

Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}. TCPCrypt goes even further and repurposes the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}.

We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified but at the same time to be compatible with most of existing, deployed network infrastructure and safely fall-back to vanilla TCP if such modifications fail. At the core of the idea is modification of TCP header fields in a way that is not interfered with by middleboxes or so that such interference can be safely recovered or ignored.

With the focus on cellular and WiFi networks due to their renown middlebox peculiarities, we make the following contributions in this work:

\begin{enumerate}
    \item Provide an analysis of current networks behavior in the presence of unspecified TCP protocol behavior.
    \item Propose a robust method to negotiate endpoint-specific behavior across middleboxes.
    \item Demonstrate that the method is effective across a number of cellular and WiFi networks.
\end{enumerate}

\section{Candidate protocol changes}

Without using the options field and only changing the meaning of header fields in a backwards-compatible manner we only have a few options:
\begin{itemize}
    \item Setting header fields to specific values without setting the corresponding flags so that they do not have a meaning to legacy implementations.
    \item Assigning one or more of the reserved field bits.
    \item Defining semantics for specific values of currently valid header fields.
\end{itemize}

In this section we overview these different options, discuss their pros and cons and show how real networks interact with them. To that end, we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes and without congestion or flow control since we only exchange two data packets once connection has been established (as well as any additional ones generated by middleboxes). We also do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.

\subsubsection*{Unset Flags}

The only header fields that may or may not have a meaning depending on the corresponding flag are \emph{urgent pointer} and \emph{acknowledgment number},with the latter only not being set at the initial SYN packet.

Urgent pointer is both a good and a bad candidate at the same time: it is already recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be claimed that it is taking up precious space in the TCP header. On the other hand, it is common practice by NIDS to reset the field as it makes it difficult to track the application-layer data~\cite{seolma}. While we witnessed the field to be passed through on many of our studied networks, we also found many that do not (Table...), therefore it is a usable but not our preferred choice.

Acknowledgment field, on the other hand, is double the length and we saw it successfully pass through the networks in more cases. Nevertheless, unless specific value is signaled out of bound between the two endpoints to use the field with TCP Simultaneous Open, only the initiating end can use it and the responding side needs to acknowledge the understanding of it in some other method.

\subsubsection*{Reserved bits}

The main difficulty with TCP Reserved bits is that there are only 3 bits unallocated if we consider the three ECN bits added over 10 years ago~\cite{Floyd:up,Ely:uc} but still not universally implemented~\cite{}. Any further reserved bit allocation would therefore be extremely difficult.

Nevertheless, there are proposals that make use of them. One of the proposals for reducing Web latency~\cite{Flach:2013uy} suggests sending multiple copies of packets, but to avoid triggering duplicate acknowledgments add a reserved bit to the header. We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgment.

We tested middlebox behavior by separately setting reserved bits in the handshake (SYN and SYNACK packets) and during data exchange, but we did not observe different behavior for the two cases: the bits were either untouched or cleared, but we did not observe cases when such packets are dropped.

\subsubsection*{Value-specific semantics}

Our remaining option is crafting packets with specific values for port numbers, initial sequence numbers, window size and specific checksum...

\section{Related Work}
\label{sec:related}

TCP implements an \emph{urgent mechanism} at its core that allows the sending user to stimulate the receiving user to accept some \emph{urgent data}. The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is specified that it is not a mechanism for sending such data. Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and in general it is recommended against the use of the mechanism~\cite{Gont:2011vi}.

In the context of MPTCP one proposal to add a generic control stream is to map such stream into a separate sequence number space and exchange control data over established subflows, only modifying MPTCP DSS option to use one of the reserved bits to differentiate the control stream from the data stream. It does, however, assume that MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.

Conceptually close to MPTCP's subflows is SCTP's \emph{data chunks}, where a packet can contain multiple TLV (Type-Length-Value) encoded chunks to separate user data and control information. Despite its challenged deployment in the Internet, SCTP over DTLS is being used as the basis of WebRTC data protocol~\cite{Tuexen:wv} and we suggest that such chunking could be used on top of \emph{noTCP} once such behavior has been negotiated as described above.

%\section{Conclusion}
%TBD
%\input{conclusion}

\ifnum\anon=0
\subsection*{Acknowledgments}

\fi

%Only show things we really cite :)
% \nocite{*}
{\footnotesize 
\bibliographystyle{abbrv}
%\footnotesize
\bibliography{references}
}

%\appendix 
%\input{data}

\end{document}

### 2014-06-05 paper/paper.tex
writing in progress commit, moving machines

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{noTCP: do not extend, repurpose}
+\title{noTCP: extend, extend, repurpose}
 \ifnum\anon=1
@@
 One of the proposals for reducing Web latency~\cite{Flach:2013uy} suggests sending multiple copies of packets, but to avoid triggering duplicate acknowledgments add a reserved bit to the header.
-We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgment.
+We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgmentof the proposed scheme.
 
@@
 
-Our remaining option is crafting packets with specific values for port numbers, initial sequence numbers, window size and specific checksum...
+Our remaining option is crafting packets with specific values for initial sequence numbers, window size or a specific checksum value.
+Initial sequence numbers, however, must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security.
+
+Window size field is potentially a better candidate, especially if used only during connection setup phase, since data is not typically transmitted during the handshake.
+We could further claim that even during data exchange a few of the least-significant bits could be sacrifised for carrying other information, however that would prevent flow control from operating properly.
+Higher-order bits could instead be given up when window scaling option is used to compensate for the derease of signalled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links.
+Finally, we observed many middleboxes to modify the field and there is a risk of the middleboxes using the initially negotiated one for assigning resources for the duration of the connection~\cite{}.
+
+There are multiple benefits of using the checksum field.
+Firstly, it is (only) used for error detection as opposed to other fields that govern protocol behavior and is either not always necessary due to the layered design of the OSI stack (when error detection is done at the layer below or above), or can be replaced with extensions' separate (MPTCP DSS checksum) or even stronger (TCPCrypt MAC) mechanisms.
+Secondly, checksum operations are efficient as they only involve one's complement addition.
+And finally, it is possible to generate checksums in a way that the intended value can recovered by the receiving end after traversing common types of middleboxes that recompute the checksum.
+
+Typically~\cite{} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimise the resouce use.
+In our sample X networks only do that, as indicated by the fact that they allow packets with incorrect checksums through towards the endpoint (otherwise the packet would either be dropped or a valid checksum would be recomputed).
+It is not always the case and another alternative is to add payload to a packet to make checksum a specific value.
+It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards.
+A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{}:
+
+\begin{quotation}
+	Although these
+	examples do not show connection synchronization using data-carrying
+	segments, this is perfectly legitimate...
+\end{quotation}
+
+\section{Protocol negotiation}
+
+What is common to all candidate procol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information.
+Such information must instead be exchanged through other means - either implicit to the opcode, or explicit by redefining the meaning of part of the data payload or previously exchanged information.
+We threfore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.
+
+\subsection{Port-specific middlebox behavior}
+
+Interestingly, on certain networks with what appears to be proxies deployed the behaviour is dependent on destination port used.
+In particular, they appear to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified.
+In fact, most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped.
+
+\section{Discussion}
+
+The main points of discussion on \emph{noTCP} is the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
+Furthermore, any special protocol negotiation still needs to be standardised to ensure compatibility at the Internet scale...
 
@@
 In the context of MPTCP one proposal to add a generic control stream is to map such stream into a separate sequence number space and exchange control data over established subflows, only modifying MPTCP DSS option to use one of the reserved bits to differentiate the control stream from the data stream.
-It does, however, assume that MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
+Arguably the specified options are complex as well as extensible to enable it are precisely because there are very few ways in which TCP itself can be extended.
+the underlying assumption, however, is taht MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
 

### 2014-06-05 paper/paper.tex
added refs, a tiny restructuring, done for the day

@@
 
+\section{Methodology}
+
+To test what works on current networks we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes and without congestion or flow control since we only exchange two data packets once connection has been established (as well as any additional ones generated by middleboxes).
+We also currently do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
+
 \section{Candidate protocol changes}
@@
 In this section we overview these different options, discuss their pros and cons and show how real networks interact with them.
-To that end, we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes and without congestion or flow control since we only exchange two data packets once connection has been established (as well as any additional ones generated by middleboxes).
-We also do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
 
@@
 One of the proposals for reducing Web latency~\cite{Flach:2013uy} suggests sending multiple copies of packets, but to avoid triggering duplicate acknowledgments add a reserved bit to the header.
-We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgmentof the proposed scheme.
+We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgment of the proposed scheme.
 
@@
 Window size field is potentially a better candidate, especially if used only during connection setup phase, since data is not typically transmitted during the handshake.
-We could further claim that even during data exchange a few of the least-significant bits could be sacrifised for carrying other information, however that would prevent flow control from operating properly.
-Higher-order bits could instead be given up when window scaling option is used to compensate for the derease of signalled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links.
+We could further claim that even during data exchange a few of the least-significant bits could be sacrificed for carrying other information, however that would prevent flow control from operating properly.
+Higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of signaled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links.
 Finally, we observed many middleboxes to modify the field and there is a risk of the middleboxes using the initially negotiated one for assigning resources for the duration of the connection~\cite{}.
@@
 
-Typically~\cite{} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimise the resouce use.
+In traditional NATs~\cite{Egevang:tu} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use.
 In our sample X networks only do that, as indicated by the fact that they allow packets with incorrect checksums through towards the endpoint (otherwise the packet would either be dropped or a valid checksum would be recomputed).
@@
 It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards.
-A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{}:
+A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
 
@@
 
-What is common to all candidate procol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information.
+What is common to all candidate protocol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information.
 Such information must instead be exchanged through other means - either implicit to the opcode, or explicit by redefining the meaning of part of the data payload or previously exchanged information.
-We threfore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.
+We therefore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.
 
@@
 
-Interestingly, on certain networks with what appears to be proxies deployed the behaviour is dependent on destination port used.
+Interestingly, on certain networks with what appears to be proxies deployed the behavior is dependent on destination port used.
 In particular, they appear to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified.
@@
 The main points of discussion on \emph{noTCP} is the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
-Furthermore, any special protocol negotiation still needs to be standardised to ensure compatibility at the Internet scale...
+Furthermore, any special protocol negotiation still needs to be standardized to ensure compatibility at the Internet scale...
 
@@
 Arguably the specified options are complex as well as extensible to enable it are precisely because there are very few ways in which TCP itself can be extended.
-the underlying assumption, however, is taht MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
+the underlying assumption, however, is that theyt MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
 

### 2014-06-09 app/src/org/smarte/tcptester/RawSocketTester.java
work on standalone app UI - kind of ready

@@ -43,8 +43,23 @@ import com.stericson.RootTools.execution.Command;
 import com.stericson.RootTools.execution.CommandCapture;
 import com.stericson.RootTools.exceptions.RootDeniedException;
 import edu.berkeley.icsi.netalyzr.tests.Test;
+import org.apache.http.message.BasicNameValuePair;
+import org.apache.http.client.entity.UrlEncodedFormEntity;
+import org.apache.http.impl.client.DefaultHttpClient;
+import org.apache.http.NameValuePair;
+import org.apache.http.client.methods.HttpPost;
+import java.io.UnsupportedEncodingException;
+import android.os.AsyncTask;
+import org.apache.http.impl.client.BasicResponseHandler;
+import org.apache.http.client.ResponseHandler;
+import android.widget.ProgressBar;
+import android.widget.TextView;
+import android.content.Intent;
+import android.app.Activity;
 
-public class RawSocketTester extends Test
+
+
+public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
 {
     public static final String TAG = "TCPTester";
     private static final String TESTER_BINARY = "tcptester";
@@ -52,16 +67,20 @@ public class RawSocketTester extends Test
         "iptables -%c OUTPUT -p tcp --tcp-flags RST RST --sport %d --dport %d -d %s -j DROP && iptables --list";
 
     private SocketTesterServer mTesterServer;
-    private Context mContext;
+    private Context mActivity;
     
     private InetAddress mServerAddress;
     private int[] mServerPorts; 
     private ArrayList<TCPTest> mResults;
+    private ProgressBar mProgress;
+    private TextView mProgressText, mSubmittedResults;
 
-    public RawSocketTester(String name, Context context) {
-        super(name);
-        mContext = context;
+    public RawSocketTester(Activity activity, ProgressBar progress, TextView progressText) {
+        super();
+        mActivity = activity;
         mResults = new ArrayList<TCPTest>();
+        mProgress = progress;
+        mProgressText = progressText;
     }
 
     public void init() {
@@ -75,9 +94,10 @@ public class RawSocketTester extends Test
         mServerPorts = new int[]{80, 443, 993, 8000, 5258, 6969};
     }    
 
-    public int runImpl() throws IOException {
-        if (!RootTools.hasBinary(mContext, TESTER_BINARY)) {
[…]

### 2014-06-09 app/src/org/smarte/tcptester/SocketTesterServer.java

@@ -172,7 +172,7 @@ public class SocketTesterServer extends Thread {
         boolean result = false;
         lock.lock();
         try {
-            // Log.d(TAG, "Sending message " + RawSocketTester.bytesToHex(msg, msg[0]));
+            // Log.d(TAG, "Sending message " + bytesToHex(msg, msg[0]));
             if (!clientConnected) {
                 Log.d(TAG, "Waiting for client to connect");
                 connected.await();
@@ -207,7 +207,7 @@ public class SocketTesterServer extends Thread {
             if (bytesRead > 0) {
                 posOffset += bytesRead;
                 totalBytesRead += bytesRead;
-                // Log.d(TAG, "Receive data from socket, bytesRead = " + bytesRead + ", " + RawSocketTester.bytesToHex(buffer, totalBytesRead));
+                // Log.d(TAG, "Receive data from socket, bytesRead = " + bytesRead + ", " + bytesToHex(buffer, totalBytesRead));
                 byte cmdLen = buffer[0];
                 if (cmdLen == totalBytesRead) {
                     // Log.d(TAG, "Full command received");
@@ -222,6 +222,17 @@ public class SocketTesterServer extends Thread {
 
     public String getLocalSocketAddress() {
         return server.getLocalSocketAddress().getName();
+    }
+
+    final protected static char[] hexArray = "0123456789ABCDEF".toCharArray();
+    public static String bytesToHex(byte[] bytes, byte length) {
+        char[] hexChars = new char[length * 2];
+        for ( int j = 0; j < length; j++ ) {
+            int v = bytes[j] & 0xFF;
+            hexChars[j * 2] = hexArray[v >>> 4];
+            hexChars[j * 2 + 1] = hexArray[v & 0x0F];
+        }
+        return new String(hexChars);
     }
 
 }

### 2014-06-09 app/src/org/smarte/tcptester/TCPTest.java

@@ -17,8 +17,12 @@
 package org.smarte.tcptester;
 
 import java.net.InetAddress;
+import java.net.UnknownHostException;
+import android.os.Parcelable;
+import android.os.Parcel;
 
-public class TCPTest {
+
+public class TCPTest implements Parcelable {
     public static final int TEST_GET_GLOBAL_IP = 21;
 
 
@@ -73,6 +77,57 @@ public class TCPTest {
         this.extras = new byte[1];
         this.extras[0] = (byte)extras;
     }
+
+    public int describeContents() {
+        return 0;
+    }
+
+    private TCPTest(Parcel in) {
+        name = in.readString();
+        opcode = in.readByte();
+        result = in.readByte() == 1 ? true : false;
+        // in.readByteArray(extras);
+        byte[] tempSrc = new byte[4];
+        in.readByteArray(tempSrc);
+        try {
+            src = InetAddress.getByAddress(tempSrc);
+        } catch (UnknownHostException e) {
+            src = null;
+        }
+        srcPort = in.readInt();
+
+        byte[] tempDst = new byte[4];
+        in.readByteArray(tempDst);
+        try {
+            dst = InetAddress.getByAddress(tempDst);
+        } catch (UnknownHostException e) {
+            dst = null;
+        }
+        dstPort = in.readInt();
+    }
+
+    public void writeToParcel(Parcel out, int flags) {
+        out.writeString(name);
+        out.writeByte(opcode);
+        out.writeByte((byte) (result ? 1 : 0));
+        // out.writeByteArray(extras);
+        out.writeByteArray(src.getAddress());
+        out.writeInt(srcPort);
+        out.writeByteArray(dst.getAddress());
+        out.writeInt(dstPort);
+    }
+
+    public static final Parcelable.Creator<TCPTest> CREATOR
+            = new Parcelable.Creator<TCPTest>() {
+        public TCPTest createFromParcel(Parcel in) {
+            return new TCPTest(in);
+        }
+
+        public TCPTest[] newArray(int size) {
+            return new TCPTest[size];
+        }
+     };
+
     public String toString() {
         String ret = " " + name 
             + " " + src.getHostAddress() + ":" + Integer.toString(srcPort) 

### 2014-06-09 app/src/org/smarte/tcptester/TCPTestAdapter.java

@@ -1,0 +1,38 @@
+package org.smarte.tcptester;
+
+import android.widget.ArrayAdapter;
+import android.view.View;
+import android.content.Context;
+import java.util.ArrayList;
+import android.view.ViewGroup;
+import android.view.LayoutInflater;
+import android.widget.TextView;
+import android.widget.ImageView;
+
+
+
+public class TCPTestAdapter extends ArrayAdapter<TCPTest> {
+    public TCPTestAdapter(Context context, ArrayList<TCPTest> results) {
+        super(context, R.layout.item_test, results);
+    }
+
+    @Override
+    public View getView(int position, View convertView, ViewGroup parent) {
+        // Get the data item for this position
+        TCPTest result = getItem(position);    
+        // Check if an existing view is being reused, otherwise inflate the view
+        if (convertView == null) {
+            convertView = LayoutInflater.from(getContext()).inflate(R.layout.item_test, parent, false);
+        }
+        // Lookup view for data population
+        TextView testPort = (TextView) convertView.findViewById(R.id.testPort);
+        TextView testName = (TextView) convertView.findViewById(R.id.testName);
+        ImageView testResult = (ImageView) convertView.findViewById(R.id.testResult);
+        // Populate the data into the template view using the data object
+        testPort.setText(Integer.toString(result.dstPort));
+        testName.setText(result.name);
+        testResult.setImageResource(result.result ? R.drawable.ic_action_accept : R.drawable.ic_action_warning);
+        // Return the completed view to render on screen
+        return convertView;
+    }
+}

### 2014-06-09 app/src/org/smarte/tcptester/TcpTesterAbout.java

@@ -41,61 +41,27 @@ import android.view.MenuInflater;
 import android.view.Menu;
 import android.view.MenuItem;
 import android.support.v7.app.ActionBar;
-import android.content.Intent;
 
 
 import edu.berkeley.icsi.netalyzr.tests.Test;
 
-public class TcpTester extends ActionBarActivity implements View.OnClickListener 
+public class TcpTesterAbout extends ActionBarActivity 
 {
     public static final String TAG = "TCPTester";
-    private ProgressBar mProgress;
-    private TextView mProgressText;
 
     public void onCreate(Bundle savedInstanceState)
     {
         super.onCreate(savedInstanceState);
-        setContentView(R.layout.form);
-        int buttonId = R.id.connect;
-        ImageButton btn = (ImageButton) findViewById(buttonId);
-        mProgress = (ProgressBar)findViewById(R.id.testProgress);
-        mProgressText = (TextView)findViewById(R.id.progressText);
-        btn.setOnClickListener(this);
-
+        setContentView(R.layout.about);
+        
         ActionBar actionBar = getSupportActionBar();
         actionBar.setDisplayHomeAsUpEnabled(true);
         actionBar.show();
     }
 
     @Override
-    public void onClick(View v) {
-        new RawSocketTester(this, mProgress, mProgressText).execute();
-    }
-
-    @Override
     public boolean onCreateOptionsMenu(Menu menu) {
-        // Inflate the menu items for use in the action bar
-        Log.d(TAG, "Options menu created");
-        MenuInflater inflater = getMenuInflater();
-        inflater.inflate(R.layout.main_activity_actions, menu);
         return super.onCreateOptionsMenu(menu);
-    }
-
-    @Override
-    public boolean onOptionsItemSelected(MenuItem item) {
-        // Handle presses on the action bar items
-        switch (item.getItemId()) {
-            case R.id.action_about:
-                openAbout();
-                return true;
-            default:
-                return super.onOptionsItemSelected(item);
-        }
-    }
-
-    private void openAbout() {
-        Intent intent = new Intent(this, TcpTesterAbout.class);
-        startActivity(intent);
-    }
+    }  
     
 }

### 2014-06-09 app/src/org/smarte/tcptester/TcpTesterResults.java

@@ -1,0 +1,250 @@
+/*
+ * Copyright (c) 2014 Andrius Aucinas <andrius.aucinas@cl.cam.ac.uk>
+ * 
+ * Permission to use, copy, modify, and distribute this software for any
+ * purpose with or without fee is hereby granted, provided that the above
+ * copyright notice and this permission notice appear in all copies.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES
+ * WITH REGARD TO THIS SOFTWARE INCLUDING ALL IMPLIED WARRANTIES OF
+ * MERCHANTABILITY AND FITNESS. IN NO EVENT SHALL THE AUTHOR BE LIABLE FOR
+ * ANY SPECIAL, DIRECT, INDIRECT, OR CONSEQUENTIAL DAMAGES OR ANY DAMAGES
+ * WHATSOEVER RESULTING FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN
+ * ACTION OF CONTRACT, NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF
+ * OR IN CONNECTION WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.
+ */
+
+package org.smarte.tcptester;
+
+import android.app.Activity;
+import android.widget.TextView;
+import android.os.Bundle;
+import android.view.View;
+import android.util.Log;
+import org.apache.http.client.methods.HttpPost;
+import java.util.ArrayList;
+import org.apache.http.NameValuePair;
+import org.apache.http.impl.client.DefaultHttpClient;
+import org.apache.http.impl.client.BasicResponseHandler;
+import org.apache.http.client.ResponseHandler;
+import java.util.List;
+import org.apache.http.client.entity.UrlEncodedFormEntity;
+import org.apache.http.message.BasicNameValuePair;
+import org.apache.http.HttpResponse;
+import java.io.UnsupportedEncodingException;
+import java.io.IOException;
+import android.widget.ImageButton;
+import android.widget.ProgressBar;
+import android.os.AsyncTask;
+import android.support.v7.app.ActionBarActivity;
+import android.view.MenuInflater;
+import android.view.Menu;
+import android.view.MenuItem;
+import android.support.v7.app.ActionBar;
+import android.net.ConnectivityManager;
+import android.net.NetworkInfo;
+import android.widget.ListView;
+import android.widget.ArrayAdapter;
+import java.util.Collection;
[…]

### 2014-06-09 app/src/org/smarte/tcptester/RawSocketTester.java
cleaned up imports

@@ -16,46 +16,36 @@
 
 package org.smarte.tcptester;
 
+import android.os.Build;
+import java.util.ArrayList;
+import android.content.pm.PackageManager.NameNotFoundException;
+import java.io.IOException;
+import java.net.UnknownHostException;
+import android.content.Intent;
+import android.content.Context;
+import java.util.concurrent.TimeoutException;
+import android.app.Activity;
+import android.widget.ProgressBar;
+import android.os.AsyncTask;
+import android.widget.TextView;
+import android.net.NetworkInfo;
+import java.util.List;
+import android.content.pm.PackageManager;
+import java.util.Random;
+import android.content.pm.PackageInfo;
+import java.net.SocketException;
+import java.util.Enumeration;
+import android.util.Log;
+import android.net.ConnectivityManager;
 import java.net.InetAddress;
 import java.net.NetworkInterface;
-import java.io.ByteArrayOutputStream;
-import java.util.ArrayList;
-import android.content.pm.PackageManager;
-import java.util.Enumeration;
-import android.net.ConnectivityManager;
-import java.nio.ByteBuffer;
-import java.util.Random;
-import java.net.SocketException;
-import android.content.Context;
-import android.content.pm.PackageInfo;
-import android.os.SystemClock;
-import java.util.List;
-import android.net.NetworkInfo;
-import android.content.pm.PackageManager.NameNotFoundException;
-import android.util.Log;
-import java.util.concurrent.TimeoutException;
-import java.net.UnknownHostException;
-import android.os.Build;
-import java.io.IOException;
+
 import com.stericson.RootTools.RootTools;
 import com.stericson.RootTools.execution.Shell;
 import com.stericson.RootTools.execution.Command;
 import com.stericson.RootTools.execution.CommandCapture;
 import com.stericson.RootTools.exceptions.RootDeniedException;
 import edu.berkeley.icsi.netalyzr.tests.Test;
-import org.apache.http.message.BasicNameValuePair;
-import org.apache.http.client.entity.UrlEncodedFormEntity;
-import org.apache.http.impl.client.DefaultHttpClient;
-import org.apache.http.NameValuePair;
-import org.apache.http.client.methods.HttpPost;
-import java.io.UnsupportedEncodingException;
-import android.os.AsyncTask;
-import org.apache.http.impl.client.BasicResponseHandler;
-import org.apache.http.client.ResponseHandler;
-import android.widget.ProgressBar;
[…]

### 2014-06-09 app/src/org/smarte/tcptester/TCPTestAdapter.java

@@ -1,3 +1,19 @@
+/*
+ * Copyright (c) 2014 Andrius Aucinas <andrius.aucinas@cl.cam.ac.uk>
+ * 
+ * Permission to use, copy, modify, and distribute this software for any
+ * purpose with or without fee is hereby granted, provided that the above
+ * copyright notice and this permission notice appear in all copies.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES
+ * WITH REGARD TO THIS SOFTWARE INCLUDING ALL IMPLIED WARRANTIES OF
+ * MERCHANTABILITY AND FITNESS. IN NO EVENT SHALL THE AUTHOR BE LIABLE FOR
+ * ANY SPECIAL, DIRECT, INDIRECT, OR CONSEQUENTIAL DAMAGES OR ANY DAMAGES
+ * WHATSOEVER RESULTING FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN
+ * ACTION OF CONTRACT, NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF
+ * OR IN CONNECTION WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.
+ */
+
 package org.smarte.tcptester;
 
 import android.widget.ArrayAdapter;
@@ -8,8 +24,6 @@ import android.view.ViewGroup;
 import android.view.LayoutInflater;
 import android.widget.TextView;
 import android.widget.ImageView;
-
-
 
 public class TCPTestAdapter extends ArrayAdapter<TCPTest> {
     public TCPTestAdapter(Context context, ArrayList<TCPTest> results) {

### 2014-06-09 app/src/org/smarte/tcptester/TcpTester.java

@@ -16,34 +16,18 @@
 
 package org.smarte.tcptester;
 
-import android.app.Activity;
+import android.view.MenuInflater;
+import android.view.View;
+import android.support.v7.app.ActionBarActivity;
+import android.support.v7.app.ActionBar;
+import android.util.Log;
+import android.content.Intent;
+import android.view.MenuItem;
+import android.widget.ImageButton;
+import android.view.Menu;
+import android.widget.ProgressBar;
 import android.widget.TextView;
 import android.os.Bundle;
-import android.view.View;
-import android.util.Log;
-import org.apache.http.client.methods.HttpPost;
-import java.util.ArrayList;
-import org.apache.http.NameValuePair;
-import org.apache.http.impl.client.DefaultHttpClient;
-import org.apache.http.impl.client.BasicResponseHandler;
-import org.apache.http.client.ResponseHandler;
-import java.util.List;
-import org.apache.http.client.entity.UrlEncodedFormEntity;
-import org.apache.http.message.BasicNameValuePair;
-import org.apache.http.HttpResponse;
-import java.io.UnsupportedEncodingException;
-import java.io.IOException;
-import android.widget.ImageButton;
-import android.widget.ProgressBar;
-import android.os.AsyncTask;
-import android.support.v7.app.ActionBarActivity;
-import android.view.MenuInflater;
-import android.view.Menu;
-import android.view.MenuItem;
-import android.support.v7.app.ActionBar;
-import android.content.Intent;
-
-import edu.berkeley.icsi.netalyzr.tests.Test;
 
 public class TcpTester extends ActionBarActivity implements View.OnClickListener 
 {

### 2014-06-09 app/src/org/smarte/tcptester/TcpTesterResults.java

@@ -16,48 +16,35 @@
 
 package org.smarte.tcptester;
 
-import android.app.Activity;
-import android.widget.TextView;
-import android.os.Bundle;
-import android.view.View;
-import android.util.Log;
-import org.apache.http.client.methods.HttpPost;
 import java.util.ArrayList;
 import org.apache.http.NameValuePair;
-import org.apache.http.impl.client.DefaultHttpClient;
-import org.apache.http.impl.client.BasicResponseHandler;
-import org.apache.http.client.ResponseHandler;
-import java.util.List;
-import org.apache.http.client.entity.UrlEncodedFormEntity;
-import org.apache.http.message.BasicNameValuePair;
-import org.apache.http.HttpResponse;
-import java.io.UnsupportedEncodingException;
-import java.io.IOException;
-import android.widget.ImageButton;
-import android.widget.ProgressBar;
-import android.os.AsyncTask;
-import android.support.v7.app.ActionBarActivity;
-import android.view.MenuInflater;
+import org.apache.http.client.methods.HttpPost;
 import android.view.Menu;
-import android.view.MenuItem;
-import android.support.v7.app.ActionBar;
-import android.net.ConnectivityManager;
-import android.net.NetworkInfo;
-import android.widget.ListView;
-import android.widget.ArrayAdapter;
-import java.util.Collection;
-import java.util.Comparator;
-import java.util.Collections;
-import android.content.Context;
-import java.util.HashMap;
-import java.util.Set;
-import android.view.ViewGroup;
-import android.view.LayoutInflater;
+import java.io.IOException;
 import android.widget.ImageView;
-
-
-
-import edu.berkeley.icsi.netalyzr.tests.Test;
+import android.content.Context;
+import android.widget.ListView;
+import java.util.HashMap;
+import org.apache.http.impl.client.DefaultHttpClient;
+import android.widget.ArrayAdapter;
+import android.os.AsyncTask;
+import android.widget.TextView;
+import android.net.NetworkInfo;
+import java.util.List;
+import android.view.View;
+import org.apache.http.client.ResponseHandler;
+import org.apache.http.message.BasicNameValuePair;
+import java.io.UnsupportedEncodingException;
+import android.view.ViewGroup;
+import org.apache.http.impl.client.BasicResponseHandler;
+import android.support.v7.app.ActionBarActivity;
+import android.util.Log;
+import android.net.ConnectivityManager;
+import android.os.Bundle;
[…]

### 2014-06-10 app/src/org/smarte/tcptester/RawSocketTester.java
minor bugfixes

@@ -78,8 +78,7 @@ public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
         } catch (UnknownHostException e) {
             mServerAddress = null;
         }
-        // TODO: also add other common ones: 80 8000 8080
-        mServerPorts = new int[]{80, 443, 993, 8000, 5258, 6969};
+        mServerPorts = new int[]{80, 443, 993, 8000, 5228, 6969};
     }    
 
     protected Integer doInBackground(Void... none) {
@@ -91,6 +90,8 @@ public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
                 Log.d(TAG, "Installing binary failed");
                 return Test.TEST_PROHIBITED;
             }
+        } else {
+            Log.d(TAG, "Native binary already exists");
         }
 
         if (!RootTools.isRootAvailable() || !RootTools.isAccessGiven()) {
@@ -192,22 +193,21 @@ public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
         basicTests.add(new TCPTest("plain-URG", 5));
         basicTests.add(new TCPTest("ACK-checksum-incorrect", 6));
         basicTests.add(new TCPTest("ACK-checksum", 7));
-        // basicTests.add(new TCPTest("URG-URG", 8));
-        // basicTests.add(new TCPTest("URG-checksum", 9));
-        // basicTests.add(new TCPTest("URG-checksum-incorrect", 10));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 1));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 2));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 4));
-        // // basicTests.add(new TCPTest("Reserved-syn", 11, 8));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 1));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 2));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 4));
-        // // basicTests.add(new TCPTest("Reserved-est", 12, 8));
-        // basicTests.add(new TCPTest("ACK-checksum-incorrect-seq", 13));
+        basicTests.add(new TCPTest("URG-URG", 8));
+        basicTests.add(new TCPTest("URG-checksum", 9));
+        basicTests.add(new TCPTest("URG-checksum-incorrect", 10));
[…]

### 2014-06-10 app/src/org/smarte/tcptester/TcpTester.java

@@ -58,7 +58,6 @@ public class TcpTester extends ActionBarActivity implements View.OnClickListener
     @Override
     public boolean onCreateOptionsMenu(Menu menu) {
         // Inflate the menu items for use in the action bar
-        Log.d(TAG, "Options menu created");
         MenuInflater inflater = getMenuInflater();
         inflater.inflate(R.layout.main_activity_actions, menu);
         return super.onCreateOptionsMenu(menu);

### 2014-06-10 app/src/org/smarte/tcptester/TcpTesterResults.java

@@ -150,11 +150,12 @@ public class TcpTesterResults extends ActionBarActivity
 
         TCPTestAggregate(Integer dstPort, Integer passed, Integer total) {
             this.dstPort = dstPort;
-            if (passed/total > 0.8) {
+            double successRate = ((double) passed) / total;
+            if (successRate > 0.7) {
                 this.result = CLEAR;
                 this.description = "All traffic allowed";
             }
-            else if (passed/total > 0.01) {
+            else if (successRate > 0.01) {
                 this.result = FILTERED;
                 this.description = "Some extensions filtered";
             }

### 2014-06-10 server/server.py

@@ -263,4 +263,4 @@ processor.start()
 sender.daemon = True
 sender.start()
 
-sniff(prn=receive_queue.put, filter="tcp and dst 192.95.61.161 and (dst port 6969 or dst port 80 or dst port 443 or dst port 8080 or dst port 8000 or dst port 445 or dst port 993 or dst port 139 or dst port 5258)", store=0)
+sniff(prn=receive_queue.put, filter="tcp and dst 192.95.61.161 and (dst port 6969 or dst port 80 or dst port 443 or dst port 8080 or dst port 8000 or dst port 445 or dst port 993 or dst port 139 or dst port 5228)", store=0)

### 2014-06-10 app/src/org/smarte/tcptester/TcpTesterResults.java
another minor tweak in the way results are displayed

@@ -185,7 +185,7 @@ public class TcpTesterResults extends ActionBarActivity
             ImageView testResult = (ImageView) convertView.findViewById(R.id.testResult);
             // Populate the data into the template view using the data object
             testPort.setText(Integer.toString(result.dstPort));
-            testName.setText(result.description);
+            testName.setText("\t"+result.description);
             if (result.result == result.CLEAR)
                 testResult.setImageResource(R.drawable.ic_action_accept);
             if (result.result == result.FILTERED)

### 2014-06-10 paper/paper.tex
table of network characteristics

@@
 
-\section{Methodology}
-
-To test what works on current networks we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes and without congestion or flow control since we only exchange two data packets once connection has been established (as well as any additional ones generated by middleboxes).
-We also currently do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
-
 \section{Candidate protocol changes}
@@
 Urgent pointer is both a good and a bad candidate at the same time: it is already recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be claimed that it is taking up precious space in the TCP header.
-On the other hand, it is common practice by NIDS to reset the field as it makes it difficult to track the application-layer data~\cite{seolma}.
-While we witnessed the field to be passed through on many of our studied networks, we also found many that do not (Table...), therefore it is a usable but not our preferred choice.
+On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) to reset the field as it makes it difficult to track the application-layer data~\cite{seolma}.
+We present our observations of how often the field is filtered in Section~\ref{sec:network}, however although we witnessed the field to be passed through on many of our studied networks and hence it is a usable option, it is not without drawbacks.
 
@@
 We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgment of the proposed scheme.
-
-We tested middlebox behavior by separately setting reserved bits in the handshake (SYN and SYNACK packets) and during data exchange, but we did not observe different behavior for the two cases: the bits were either untouched or cleared, but we did not observe cases when such packets are dropped.
 
@@
 
-\section{Protocol negotiation}
+\section{Methodology}
 
-What is common to all candidate protocol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information.
-Such information must instead be exchanged through other means - either implicit to the opcode, or explicit by redefining the meaning of part of the data payload or previously exchanged information.
-We therefore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.
+To test what works on current networks we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes and without congestion or flow control since we only exchange two data packets once connection has been established (as well as any additional ones generated by middleboxes).
+We also currently do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
+
+\section{Network cahracteristics}
+\label{sec:network}
+
+In this section we look at the different network behaviours as observed by our testing tool.
+
+We tested middlebox behavior by separately setting reserved bits in the handshake (SYN and SYNACK packets) and during data exchange, but we did not observe different behavior for the two cases: the bits were either untouched or cleared, but we did not observe cases when such packets are dropped.
+
+\begin{table*}[t]
+{\small
+\begin{center}
+\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
+\hline
+    \textbf{Net ID} & \textbf{No filter} & \textbf{Global IP}   & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize ACK}  & \textbf{Normalize URG} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
+    % Univ Helsinki WiFi
+    WiFi education 1 &                   &                      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    % Eduroam
+    WiFi education 2 &                   & \checkmark           &                        &                            & \checkmark             &                         & \checkmark \checkmark  &                             & \checkmark          \\ \hline
+    % wgb
+    WiFi public 1      & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    % Finland internet.saunalahti
+    Cellular Finland 1 & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    % Virgin home
+    WiFi residential 1 &                 &                      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    % GiffGaff
+    Cellular UK 1      &                 &                      &  \checkmark (443, 993) & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+\end{tabular}
+\end{center}
+}
+\caption{Network behaviour observed through tests generating custom TCP packets}
+\label{tab:networks}
+\end{table*}
 
@@
 In fact, most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped.
+
+\section{Protocol negotiation}
+
+What is common to all candidate protocol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information.
+Such information must instead be exchanged through other means - either implicit to the opcode, or explicit by redefining the meaning of part of the data payload or previously exchanged information.
+We therefore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.
 

### 2014-06-10 app/src/org/smarte/tcptester/RawSocketTester.java
Fixed global IP address reporting

@@ -187,29 +187,29 @@ public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
 
     private ArrayList<TCPTest> buildTests(InetAddress serverAddress, int[] serverPorts) {
         ArrayList<TCPTest> basicTests = new ArrayList<TCPTest>();
-        basicTests.add(new TCPTest("ACK-only", 2));
-        basicTests.add(new TCPTest("URG-only", 3));
-        // basicTests.add(new TCPTest("ACK-URG", 4));
-        basicTests.add(new TCPTest("plain-URG", 5));
-        basicTests.add(new TCPTest("ACK-checksum-incorrect", 6));
-        basicTests.add(new TCPTest("ACK-checksum", 7));
-        basicTests.add(new TCPTest("URG-URG", 8));
-        basicTests.add(new TCPTest("URG-checksum", 9));
-        basicTests.add(new TCPTest("URG-checksum-incorrect", 10));
-        basicTests.add(new TCPTest("Reserved-syn", 11, 1));
-        basicTests.add(new TCPTest("Reserved-syn", 11, 2));
-        basicTests.add(new TCPTest("Reserved-syn", 11, 4));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 8));
-        basicTests.add(new TCPTest("Reserved-est", 12, 1));
-        basicTests.add(new TCPTest("Reserved-est", 12, 2));
-        basicTests.add(new TCPTest("Reserved-est", 12, 4));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 8));
-        basicTests.add(new TCPTest("ACK-checksum-incorrect-seq", 13));
+        // basicTests.add(new TCPTest("ACK-only", 2));
+        // basicTests.add(new TCPTest("URG-only", 3));
+        // // basicTests.add(new TCPTest("ACK-URG", 4));
+        // basicTests.add(new TCPTest("plain-URG", 5));
+        // basicTests.add(new TCPTest("ACK-checksum-incorrect", 6));
+        // basicTests.add(new TCPTest("ACK-checksum", 7));
+        // basicTests.add(new TCPTest("URG-URG", 8));
+        // basicTests.add(new TCPTest("URG-checksum", 9));
+        // basicTests.add(new TCPTest("URG-checksum-incorrect", 10));
+        // basicTests.add(new TCPTest("Reserved-syn", 11, 1));
[…]

### 2014-06-10 app/src/org/smarte/tcptester/TCPTest.java

@@ -86,7 +86,9 @@ public class TCPTest implements Parcelable {
         name = in.readString();
         opcode = in.readByte();
         result = in.readByte() == 1 ? true : false;
-        // in.readByteArray(extras);
+        int extrasLength = in.readInt();
+        extras = new byte[extrasLength];
+        in.readByteArray(extras);
         byte[] tempSrc = new byte[4];
         in.readByteArray(tempSrc);
         try {
@@ -110,7 +112,8 @@ public class TCPTest implements Parcelable {
         out.writeString(name);
         out.writeByte(opcode);
         out.writeByte((byte) (result ? 1 : 0));
-        // out.writeByteArray(extras);
+        out.writeInt(extras.length);
+        out.writeByteArray(extras);
         out.writeByteArray(src.getAddress());
         out.writeInt(srcPort);
         out.writeByteArray(dst.getAddress());

### 2014-06-10 app/src/org/smarte/tcptester/TcpTesterResults.java

@@ -98,7 +98,7 @@ public class TcpTesterResults extends ActionBarActivity
                 for (int i = 0; i < 4; i++) {
                     ip += Integer.toString((int)(result.extras[i] & 0xFF)) + (i < 3 ? "." : "");
                 }
-                globalIP.setText(R.string.your_IP + " " + ip);
+                globalIP.setText(getString(R.string.your_IP) + " " + ip);
                 continue;
             }
             if (result.result) {

### 2014-06-10 paper/paper.tex
a bit of restructure, giving draft to Jon

@@
 We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgment of the proposed scheme.
+Due to the problem with reserved bit allocation we would recommend using mechanisms we discuss later.
 
@@
 
-Our remaining option is crafting packets with specific values for initial sequence numbers, window size or a specific checksum value.
-Initial sequence numbers, however, must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security.
+Our remaining class of options, and by far the biggest, is defining protocol semantics based on specific values of certain header fields.
+It would necessitate generating crafting packets with specific values for initial sequence numbers, window size or a specific checksum value, and we posit that it would be an acceptable choice if such solution is deployable in today's network and can be implemented efficiently.
+
+Initial sequence numbers, must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security.
 
@@
 Secondly, checksum operations are efficient as they only involve one's complement addition.
-And finally, it is possible to generate checksums in a way that the intended value can recovered by the receiving end after traversing common types of middleboxes that recompute the checksum.
-
-In traditional NATs~\cite{Egevang:tu} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use.
-In our sample X networks only do that, as indicated by the fact that they allow packets with incorrect checksums through towards the endpoint (otherwise the packet would either be dropped or a valid checksum would be recomputed).
-It is not always the case and another alternative is to add payload to a packet to make checksum a specific value.
-It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards.
-A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
-
-\begin{quotation}
-	Although these
-	examples do not show connection synchronization using data-carrying
-	segments, this is perfectly legitimate...
-\end{quotation}
+And finally, it is possible to generate checksums in a way that the intended value can recovered by the receiving end after traversing common types of middleboxes that recompute the checksum as we discuss in the following sections.
 
@@
 
-To test what works on current networks we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes and without congestion or flow control since we only exchange two data packets once connection has been established (as well as any additional ones generated by middleboxes).
-We also currently do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
+To test what works on current networks we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes.
+
+The test has been implemented as an Android application and distributed through the \emph{Google Play Store} to collect data from volunteers.
+The overall structure of the tool is relatively straightforward:
+\begin{itemize}
+    \item UI which reports aggregated results to the end-users, sends results to the reporting server and manages the overall flow of the tests.
+    \item Native implementation of the simplified TCP stack using Raw Unix sockets (\emph{SOCK\_RAW}) to allow us to generate packets at the network layer.
+This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
+    \item Similar server implementation that generates specific answers to incoming packets based on header and payload values.
+\end{itemize}
+
+Potential limitations of our methodology include that we do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
+To make sure that it does not happen all our packets contain minimal amount of payload (up to 16 bytes during data exchange) and data exchange only consists of a single message sent each way, in addition to any extra acknowlegements generated by middleboxes.
+Due to the same reason we also do not take into account congestion control.
 
@@
 
-In this section we look at the different network behaviours as observed by our testing tool.
-
-We tested middlebox behavior by separately setting reserved bits in the handshake (SYN and SYNACK packets) and during data exchange, but we did not observe different behavior for the two cases: the bits were either untouched or cleared, but we did not observe cases when such packets are dropped.
-
 \begin{table*}[t]
@@
 \begin{center}
-\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
+\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
 \hline
@@
 }
-\caption{Network behaviour observed through tests generating custom TCP packets}
+\caption{Network behaviour observed through tests generating custom TCP packets.
+A single checkmark means that particular behaviour was observed or in case of normalization the field was reset or replaced with a valid value.
+Double checkmark means that specific type of packet was discarded.
+A dash means that we observed different cases on the same network.}
 \label{tab:networks}
@@
 
+We show our collected results in Table~\ref{tab:networks}.
+Since we were focusing on mobile applications, we collected data across a number of public, residential and university WiFi networks and cellular networks in XX different countries.
+The results outline for which networks we did not see any filtering, which ones discarded packets with invalid checksums, normalised particular fields (either dropped packets or reset to different values) and what type of mapping did NATs perform - only port number remapping or also changed sequence numbers.
+
+Checksum behaviour was particularly interesting.
+In traditional NATs~\cite{Egevang:tu} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use.
+In our sample X networks only do that, as indicated by the fact that they allow packets with incorrect checksums through towards the endpoint (otherwise the packet would either be dropped or a valid checksum would be recomputed).
+It is not always the case and another alternative is to add payload to a packet to make checksum a specific value.
+It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards.
+A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
+
+\begin{quotation}
+    Although these examples do not show connection synchronization using data-carrying segments, this is perfectly legitimate...
+\end{quotation}
+
+We also tested middlebox itneraction with reserved header bits.
+We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: the bits were either passwd through or cleared at both phases, but we did not observe cases when such packets are dropped.
+
 \subsection{Port-specific middlebox behavior}
 
-Interestingly, on certain networks with what appears to be proxies deployed the behavior is dependent on destination port used.
-In particular, they appear to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified.
+One column of our table of observations is especially interesting -- port-specific behavior.
+We have only observed it on one network, but the particular network appears to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified.
 In fact, most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped.
+There may be other ports configured to allow unmodified traffic through, but we only focused on commonly used ports.
+
+In addition to noting that our traffic was not encrypted but only sent to a specific destination port, it is also important to say that an increasing proportion of traffic is served over SSL.
+A commerciall deployed opt-in HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP, although varying with time~\footnote{http://db.awazza.com/users/global/}.
+
+All other traffic, however, is intercepted at the middlebox which we have separately confirmed to be a Bytemobile proxy~\footnote{Citrix Bytemobile - is it appropriate to reference this case accurately?} and it terminates clients' connections, establishing new ones to the destination, as well as adding options not present in original packets, changing window size, assigning new source port numbers sequentially and even postponing handshaking until the handshake with the client is complete.
 

### 2014-06-12 app/src/org/smarte/tcptester/RawSocketTester.java
version sent to various friends for testing

@@ -187,24 +187,24 @@ public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
 
     private ArrayList<TCPTest> buildTests(InetAddress serverAddress, int[] serverPorts) {
         ArrayList<TCPTest> basicTests = new ArrayList<TCPTest>();
-        // basicTests.add(new TCPTest("ACK-only", 2));
-        // basicTests.add(new TCPTest("URG-only", 3));
-        // // basicTests.add(new TCPTest("ACK-URG", 4));
-        // basicTests.add(new TCPTest("plain-URG", 5));
-        // basicTests.add(new TCPTest("ACK-checksum-incorrect", 6));
-        // basicTests.add(new TCPTest("ACK-checksum", 7));
-        // basicTests.add(new TCPTest("URG-URG", 8));
-        // basicTests.add(new TCPTest("URG-checksum", 9));
-        // basicTests.add(new TCPTest("URG-checksum-incorrect", 10));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 1));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 2));
-        // basicTests.add(new TCPTest("Reserved-syn", 11, 4));
-        // // basicTests.add(new TCPTest("Reserved-syn", 11, 8));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 1));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 2));
-        // basicTests.add(new TCPTest("Reserved-est", 12, 4));
-        // // basicTests.add(new TCPTest("Reserved-est", 12, 8));
-        // basicTests.add(new TCPTest("ACK-checksum-incorrect-seq", 13));
+        basicTests.add(new TCPTest("ACK-only", 2));
+        basicTests.add(new TCPTest("URG-only", 3));
+        // basicTests.add(new TCPTest("ACK-URG", 4));
+        basicTests.add(new TCPTest("plain-URG", 5));
+        basicTests.add(new TCPTest("ACK-checksum-incorrect", 6));
+        basicTests.add(new TCPTest("ACK-checksum", 7));
+        basicTests.add(new TCPTest("URG-URG", 8));
+        basicTests.add(new TCPTest("URG-checksum", 9));
+        basicTests.add(new TCPTest("URG-checksum-incorrect", 10));
+        basicTests.add(new TCPTest("Reserved-syn", 11, 1));
[…]

### 2014-06-13 paper/paper.tex
branch switching commit..

@@
 \fi
-\newcommand{\Ynote}[1]{{\authnote{Yan}{#1}}}
 \newcommand{\Anote}[1]{{\authnote{Andrius}{#1}}}
-\newcommand{\Nnote}[1]{{\authnote{Narseo}{#1}}}
-\newcommand{\Vnote}[1]{{\authnote{Vijay}{#1}}}
-\newcommand{\Dnote}[1]{{\authnote{Dina}{#1}}}
 
@@
 \begin{center}
-\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
+\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
 \hline
-    \textbf{Net ID} & \textbf{No filter} & \textbf{Global IP}   & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize ACK}  & \textbf{Normalize URG} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
+    \textbf{Net ID} & \textbf{Country} & \textbf{No filter} & \textbf{Global IP}   & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize ACK}  & \textbf{Normalize URG} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
     % Univ Helsinki WiFi
-    WiFi education 1 &                   &                      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    WiFi edu 1      & FI               &                    &                      &                        & \checkmark \checkmark      &                        &                         &                        &                             &                     \\ \hline
     % Eduroam
-    WiFi education 2 &                   & \checkmark           &                        &                            & \checkmark             &                         & \checkmark \checkmark  &                             & \checkmark          \\ \hline
+    WiFi edu 2      & UK               &                    & \checkmark           &                        &                            & \checkmark             &                         & \checkmark \checkmark  &                             & \checkmark          \\ \hline
     % wgb
-    WiFi public 1      & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    WiFi pub 1      & UK               & \checkmark         &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    % Virgin home
+    WiFi res 1      & UK               &                    &                      &                        & \checkmark \checkmark      &                        &                         &                        &                             &                     \\ \hline
+    \hline
     % Finland internet.saunalahti
-    Cellular Finland 1 & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
-    % Virgin home
-    WiFi residential 1 &                 &                      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    Cellular 1      & FI               & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
     % GiffGaff
-    Cellular UK 1      &                 &                      &  \checkmark (443, 993) & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    Cellular 2      & UK               &                 &                      &  \checkmark (443, 993) & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
 \end{tabular}
@@
 We show our collected results in Table~\ref{tab:networks}.
-Since we were focusing on mobile applications, we collected data across a number of public, residential and university WiFi networks and cellular networks in XX different countries.
-The results outline for which networks we did not see any filtering, which ones discarded packets with invalid checksums, normalised particular fields (either dropped packets or reset to different values) and what type of mapping did NATs perform - only port number remapping or also changed sequence numbers.
+Since we were focusing on mobile applications, we collected data across a number of public, residential and university WiFi networks and cellular networks in XX different countries: UK, USA, Finland (pending: Turkey, India, Lithuania, Spain...).
+We picked the most important dimensions that show what modifications to the protocol have the highest chance of being successull across the largest number of networks.
+In particular, the results include which networks provided clients a global IP address (very few), which ones we did not see any filtering for (a relatively large, but insufficient number to take any deployment for granted), and which ones discarded packets with invalid checksums, normalised particular fields - either dropped packets or reset to different values - and what type of mapping did NATs perform - only port number remapping or also changed sequence numbers.
 
@@
 \begin{quotation}
-    Although these examples do not show connection synchronization using data-carrying segments, this is perfectly legitimate...
+    Although these examples do not show connection synchronization using data-carrying segments, this is perfectly legitimate
 \end{quotation}
@@
 We also tested middlebox itneraction with reserved header bits.
-We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: the bits were either passwd through or cleared at both phases, but we did not observe cases when such packets are dropped.
+We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: all reserved bits were either allowed through or cleared at both phases.
+Importantly, we did not observe cases when such packets are dropped.
 

### 2014-06-13 app/src/org/smarte/tcptester/RawSocketTester.java
two bugfixes in collecting results and iptables printing contents unnecessarily

@@ -52,7 +52,7 @@ public class RawSocketTester extends AsyncTask<Void, Integer, Integer>
     public static final String TAG = "TCPTester";
     private static final String TESTER_BINARY = "tcptester";
     private static final String IPTABLES_CMD = 
-        "iptables -%c OUTPUT -p tcp --tcp-flags RST RST --sport %d --dport %d -d %s -j DROP && iptables --list";
+        "iptables -%c OUTPUT -p tcp --tcp-flags RST RST --sport %d --dport %d -d %s -j DROP";
 
     private SocketTesterServer mTesterServer;
     private Context mActivity;

### 2014-06-13 app/src/org/smarte/tcptester/TCPTest.java

@@ -87,8 +87,10 @@ public class TCPTest implements Parcelable {
         opcode = in.readByte();
         result = in.readByte() == 1 ? true : false;
         int extrasLength = in.readInt();
-        extras = new byte[extrasLength];
-        in.readByteArray(extras);
+        if (extrasLength > 0) {
+            extras = new byte[extrasLength];
+            in.readByteArray(extras);
+        }
         byte[] tempSrc = new byte[4];
         in.readByteArray(tempSrc);
         try {
@@ -112,8 +114,12 @@ public class TCPTest implements Parcelable {
         out.writeString(name);
         out.writeByte(opcode);
         out.writeByte((byte) (result ? 1 : 0));
-        out.writeInt(extras.length);
-        out.writeByteArray(extras);
+        if (extras != null) {
+            out.writeInt(extras.length);
+            out.writeByteArray(extras);
+        } else {
+            out.writeInt(0);
+        }
         out.writeByteArray(src.getAddress());
         out.writeInt(srcPort);
         out.writeByteArray(dst.getAddress());

### 2014-06-13 paper/paper.tex
minor changes and two more data points

@@
 The lack of extensibility has become acutely noticeable with the development of TCP extensions such as MPTCP and TCPCrypt, the former to make use of multi-homed endpoints, particularly useful in wireless networks and the latter to provide ubiquitous end-to-end traffic encryption.
+Even congestion control, the core feature of TCP, has required multiple extensions since protocol's inception and the large number of proposed congestion control schemes~\cite{5462976} suggests that more may be needed.
 These and other extensions are primarily designed around the limitations of existing middleboxes, but they demonstrate two problems: 1) any extension design is extremely constrained by prevalent middlebox behaviors\cite{Honda:2011ci} and 2) we have already run out of TCP option space, the built-in mechanism for extending TCP.
@@
 
+Instead of trying to infer which source port number would be assigned to a new connection by the kernel network stack without using it, we randomised the selection within the unprivileged port range.
+It was a simple optimisation that works on any device and minimises the probability of multiple tests failing due to another application using a particular port.
+We also repeated the tests on different destination ports, which we discuss in Sec.~\ref{sec:portspec}.
+
 \section{Network cahracteristics}
@@
     WiFi pub 1      & UK               & \checkmark         &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    % MKSW germany (carlos)
+    WiFi pub 2      & DE               &                    &                      &                        & \checkmark \checkmark      & \checkmark             &                         &                        &                             &                     \\ \hline
     % Virgin home
@@
     Cellular 2      & UK               &                    &                      &  \checkmark (443, 993) & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    % E-Plus germany
+    Cellular 3      & DE               &                    &                      &                        &                            & \checkmark             &                         & \checkmark \checkmark  &                             & \checkmark          \\ \hline
 \end{tabular}
@@
 It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards.
-A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
+A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets - a surprising 50\% of our tested networks - even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
 
@@
 \subsection{Port-specific middlebox behavior}
+\label{sec:portspec}
 
-One column of our table of observations is especially interesting -- port-specific behavior.
-We have only observed it on one network, but the particular network appears to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified.
-In fact, most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped.
+One column of our table of observations is especially interesting -- port-specific behavior. We run our tests on a number of different ports:
+\begin{itemize}
+    \item 80 - HTTP port allowed through most firewalls
+    \item 443 - HTTPS port
+    \item 993 - Secure IMAP port
+    \item 5228 - Google cloud messaging port
+    \item 6969 - random port number
+    \item 8000 - common HTTP proxy port number
+\end{itemize}
+
+So far we have only observed port-specific behavior on one network, \emph{Cellular 2} but the particular network appears to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified - most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped.
 There may be other ports configured to allow unmodified traffic through, but we only focused on commonly used ports.
+
+All other traffic, however, is intercepted at the middlebox which we have separately confirmed to be a Bytemobile proxy~\footnote{Citrix Bytemobile - is it appropriate to reference this case accurately?} and it terminates clients' connections, establishing new ones to the destination, as well as adding options not present in original packets, changing window size, assigning new source port numbers sequentially and even postponing handshaking until the handshake with the client is complete.
+Such behavior makes deployment of any protocol modifications or extensions very difficult (e.g.
+we verified that TCPCrypt does not work on this network) and although we did not see many cases of it, the product's website claims that they have over 160 clients, which suggests that it may be common.
 
@@
 
-All other traffic, however, is intercepted at the middlebox which we have separately confirmed to be a Bytemobile proxy~\footnote{Citrix Bytemobile - is it appropriate to reference this case accurately?} and it terminates clients' connections, establishing new ones to the destination, as well as adding options not present in original packets, changing window size, assigning new source port numbers sequentially and even postponing handshaking until the handshake with the client is complete.
+\subsection{Proxies and NATs}
+
+Based on results from Netalyzr?
 
@@
 Furthermore, any special protocol negotiation still needs to be standardized to ensure compatibility at the Internet scale...
+
+Nevertheless, it is important
 
@@
 
-%\section{Conclusion}
-%TBD
+\section{Conclusion}
+TBD
 %\input{conclusion}
@@
 
+\clearpage
 %Only show things we really cite :)

### 2014-07-08 paper/paper.tex
first draft

--- 2014-06-13 paper/paper.tex
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{noTCP: extend, extend, repurpose}
+\title{noTCP: extending across middleboxes through unspecified behavior}
 \ifnum\anon=1
@@
 \else
-\numberofauthors{2}
+\numberofauthors{3}
 \author{
@@
 \affaddr{University of Cambridge}
+\and
+\alignauthor Narseo Vallina-Rodriguez\\
+\affaddr{ICSI}
 }
@@
 \begin{abstract}
-Too abstract.
+TCP has become the de facto transport protocol in the Internet.
+The necessity to extend it as well as the difficulty of doing so has lead researchers to explore new ways of exchanging control information between endpoints.
+In this paper we explore the choices for such exchange without using more TCP options.
+To achieve it we design noTCP - a mechanism which uses TCP header fields in combinations that do not have specified semantics to exchange a small amount of control information between endpoints to bootstrap a control channel.
+We also present a preliminary study which shows the mechanism works across large cellular and WiFi networks.
 \end{abstract}
@@
 Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large.
-The problem primarily arises from the various middleboxes that perform packet content processing withing the network to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems and is already well documented~\cite{UntoldMiddlebox2011,Qian:2012bj,Honda:2011ci,Guha:2005tb}.
+The problem is caused by the various middleboxes that aim to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems.
+Even congestion control has required multiple extensions since protocol's inception and the large number of proposed congestion control schemes~\cite{5462976} suggests that more will be necessary.
 
-The lack of extensibility has become acutely noticeable with the development of TCP extensions such as MPTCP and TCPCrypt, the former to make use of multi-homed endpoints, particularly useful in wireless networks and the latter to provide ubiquitous end-to-end traffic encryption.
-These and other extensions are primarily designed around the limitations of existing middleboxes, but they demonstrate two problems: 1) any extension design is extremely constrained by prevalent middlebox behaviors\cite{Honda:2011ci} and 2) we have already run out of TCP option space, the built-in mechanism for extending TCP.
+In an ideal world all infrastructure is forwards compatible and changes together with the needs of the users.
+Instead, TCP has become the de facto transport protocol of the Internet and development of its extensions has been very challenging.
+Proposals including MPTCP~\cite{Handley:vj} for multi-homed endpoints and TCPCrypt~\cite{Mazieres:uz} for ubiquitous end-to-end traffic encryption are necessary for today's networks but deployment has been slow.
+
+Although TCP has been built to be extensible via options and reserved bits, all extensions are primarily designed around the limitations of existing network middleboxes\cite{Honda,Guha2005tb,UntoldMiddlebox2011}, but even if options were not filtered, we have run out of the 40-byte TCP option space.
 
 Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}.
-TCPCrypt goes even further and repurposes the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}.
+TCPCrypt uses the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}.
+An alternative way of looking at the problem is to hide the information from the middleboxes as done by steganographic approaches~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz}.
 
-We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified but at the same time to be compatible with most of existing, deployed network infrastructure and safely fall-back to vanilla TCP if such modifications fail.
+We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified.
+At the same time it is designed to work across most of deployed network infrastructure.
 At the core of the idea is modification of TCP header fields in a way that is not interfered with by middleboxes or so that such interference can be safely recovered or ignored.
+The idea builds up from the work on steganography using TCP~\cite{Frczek:2012dl,Murdoch:2005fz}, but we shift the focus from using sequence numbers and options to other fields as well.
 
@@
 \begin{enumerate}
-    \item Provide an analysis of current networks behavior in the presence of unspecified TCP protocol behavior.
-    \item Propose a robust method to negotiate endpoint-specific behavior across middleboxes.
-    \item Demonstrate that the method is effective across a number of cellular and WiFi networks.
+    \item Review the aspects of the TCP protocol that can potentially be used to extend it (Section~\ref{sec:protocolChanges}).
+    \item Provide an analysis of current mobile (WiFi and cellular) network behavior in the presence of such modifications (Section~\ref{sec:network}).
+    \item Propose a robust method to bootstrap control-stream between two endpoints across middleboxes (Section~\ref{sec:protocol}).
 \end{enumerate}
@@
 \section{Candidate protocol changes}
+\label{sec:protocolChanges}
 
@@
 \subsubsection*{Unset Flags}
+\label{sec:unset}
 
-The only header fields that may or may not have a meaning depending on the corresponding flag are \emph{urgent pointer} and \emph{acknowledgment number},with the latter only not being set at the initial SYN packet.
+The only header fields that may or may not have a meaning depending on the corresponding flag are \emph{urgent pointer} and \emph{acknowledgment number}, and the latter is only allowed to be not set for the initial SYN packet.
 
@@
 
-The main difficulty with TCP Reserved bits is that there are only 3 bits unallocated if we consider the three ECN bits added over 10 years ago~\cite{Floyd:up,Ely:uc} but still not universally implemented~\cite{}.
+The main difficulty with TCP Reserved bits is that there are only 3 bits unallocated if we consider the three ECN bits added over 10 years ago~\cite{Floyd:up,Ely:uc} but still not universally implemented~\cite{Kuhlewind:2013hr}.
 Any further reserved bit allocation would therefore be extremely difficult.
@@
 
-Initial sequence numbers, must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security.
+Initial sequence numbers must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security which has been problematic~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.
 
@@
 Higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of signaled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links.
-Finally, we observed many middleboxes to modify the field and there is a risk of the middleboxes using the initially negotiated one for assigning resources for the duration of the connection~\cite{}.
+Using the field risks interfering both with TCP flow control mechanism as well as window-size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed \eg by Citrix ByteMobile Proxy~\cite{Ha:2006td}.
 
 There are multiple benefits of using the checksum field.
-Firstly, it is (only) used for error detection as opposed to other fields that govern protocol behavior and is either not always necessary due to the layered design of the OSI stack (when error detection is done at the layer below or above), or can be replaced with extensions' separate (MPTCP DSS checksum) or even stronger (TCPCrypt MAC) mechanisms.
+Firstly, it is only used for error detection as opposed to other fields that govern protocol behavior and is either not always necessary due to the layered design of the OSI stack (when error detection is done at the layer below or above), or can be replaced with extensions' separate (MPTCP DSS checksum) or even stronger (TCPCrypt MAC) mechanisms.
 Secondly, checksum operations are efficient as they only involve one's complement addition.
@@
 This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
-    \item Similar server implementation that generates specific answers to incoming packets based on header and payload values.
+    \item Server implementation that generates specific answers to incoming packets based on header and payload values.
 \end{itemize}
@@
 Potential limitations of our methodology include that we do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
-To make sure that it does not happen all our packets contain minimal amount of payload (up to 16 bytes during data exchange) and data exchange only consists of a single message sent each way, in addition to any extra acknowlegements generated by middleboxes.
+To make sure that it does not happen all our packets contain minimal amount of payload (up to 16 bytes during data exchange) and data exchange only consists of a single message sent each way, in addition to any extra acknowledgments generated by middleboxes.
 Due to the same reason we also do not take into account congestion control.
 
-\section{Network cahracteristics}
+Instead of trying to infer which source port number would be assigned to a new connection by the kernel network stack without using it, we randomized the selection within the unprivileged port range.
+It was a simple optimization that works on any device and minimizes the probability of multiple tests failing due to another application using a particular port.
+We also repeated the tests on different destination ports, which we discuss in Sec.~\ref{sec:portspec}.
+
+Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks.
+The tests were run separately and we used it to discover non-transparent proxies.
+This test works by checking if proxies respond to a TCP connection from the client: SYN request is going to a server that is instrumented not to reply to the client and the client would not be able to finish the handshake without a proxy, otherwise there is a proxy.
+To see if proxying depend on the higher-level protocol, it also detects HTTP proxies by checking content and header changes, different IP address seen by the server for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed) and some other aspects.
+Results are presented in Sec.~\ref{sec:proxies}.
+
+\section{Network characteristics}
 \label{sec:network}
@@
 \begin{center}
-\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
+\begin{tabular}{| l | >{\centering\arraybackslash}m{1.2cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
 \hline
-    \textbf{Net ID} & \textbf{Country} & \textbf{No filter} & \textbf{Global IP}   & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize ACK}  & \textbf{Normalize URG} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
+    \textbf{Net ID} & \textbf{Country} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize Ack \#}  & \textbf{Normalize Urg ptr} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
     % Univ Helsinki WiFi
-    WiFi edu 1      & FI               &                    &                      &                        & \checkmark \checkmark      &                        &                         &                        &                             &                     \\ \hline
+    WiFi edu 1      & FI               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
     % Eduroam
-    WiFi edu 2      & UK               &                    & \checkmark           &                        &                            & \checkmark             &                         & \checkmark \checkmark  &                             & \checkmark          \\ \hline
+    WiFi edu 2      & UK               &                        &                            & \checkmark             &                         & \checkmark             &                             & \checkmark          \\ \hline
     % wgb
-    WiFi pub 1      & UK               & \checkmark         &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    WiFi pub 1      & UK               &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    % MKSW germany (carlos)
+    WiFi pub 2      & DE               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     % Virgin home
-    WiFi res 1      & UK               &                    &                      &                        & \checkmark \checkmark      &                        &                         &                        &                             &                     \\ \hline
+    WiFi res 1      & UK               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    % TEO Lt
+    WiFi res 2      & LT               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    % Telefonica Free
+    WiFi ent 1      & ES               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     \hline
     % Finland internet.saunalahti
-    Cellular 1      & FI               & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    Cellular 1      & FI               &                        &                            &                        &                         &                        &                             &                     \\ \hline
     % GiffGaff
-    Cellular 2      & UK               &                 &                      &  \checkmark (443, 993) & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    Cellular 2      & UK               &  443, 993              & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    % E-Plus germany
+    Cellular 3      & DE               &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
+    % EE/T-Mobile
+    Cellular 4      & UK               &               80, 443  & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    % Three UK
+    Cellular 5      & UK               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    % Omnitel LT
+    Cellular 6      & LT               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
 \end{tabular}
@@
 }
-\caption{Network behaviour observed through tests generating custom TCP packets.
-A single checkmark means that particular behaviour was observed or in case of normalization the field was reset or replaced with a valid value.
-Double checkmark means that specific type of packet was discarded.
-A dash means that we observed different cases on the same network.}
+\caption{Network behavior observed through tests generating custom TCP packets. A dash means that we observed different cases on the same network.}
 \label{tab:networks}
@@
 We show our collected results in Table~\ref{tab:networks}.
-Since we were focusing on mobile applications, we collected data across a number of public, residential and university WiFi networks and cellular networks in XX different countries: UK, USA, Finland (pending: Turkey, India, Lithuania, Spain...).
-We picked the most important dimensions that show what modifications to the protocol have the highest chance of being successull across the largest number of networks.
-In particular, the results include which networks provided clients a global IP address (very few), which ones we did not see any filtering for (a relatively large, but insufficient number to take any deployment for granted), and which ones discarded packets with invalid checksums, normalised particular fields - either dropped packets or reset to different values - and what type of mapping did NATs perform - only port number remapping or also changed sequence numbers.
+Since we were focusing on mobile applications, in our preliminary study we collected results across a few public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries: UK, Finland, Germany and Lithuania.
 
-Checksum behaviour was particularly interesting.
-In traditional NATs~\cite{Egevang:tu} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use.
-In our sample X networks only do that, as indicated by the fact that they allow packets with incorrect checksums through towards the endpoint (otherwise the packet would either be dropped or a valid checksum would be recomputed).
-It is not always the case and another alternative is to add payload to a packet to make checksum a specific value.
-It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards.
-A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
+The results include which networks did not filter any modifications (one WiFi and two cellular nets), and which ones discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers.
+Only one network provided clients with a global IP (WiFi edu 2).
+
+In traditional NATs~\cite{Egevang:tu} checksum recalculation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use.
+The networks that deliver packets with invalid checksum only do that as otherwise a packet with invalid checksum would be dropped.
+
+A bigger issue we have discovered is that middleboxes do not always respect the standard and simply drop SYN-ACK packets with payload.
+We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
 
@@
 
-We also tested middlebox itneraction with reserved header bits.
+As discussed in Section~\ref{sec:unset}, we can squeeze in a few extra bits of information by defining meaning to header fields when corresponding flags are not set.
+None of our tested networks filtered Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with Urgent Pointer set without URG flag.
+
+Finally we tested interaction reserved header bits.
 We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: all reserved bits were either allowed through or cleared at both phases.
@@
 \subsection{Port-specific middlebox behavior}
+\label{sec:portspec}
 
-One column of our table of observations is especially interesting -- port-specific behavior.
-We have only observed it on one network, but the particular network appears to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified.
-In fact, most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped.
-There may be other ports configured to allow unmodified traffic through, but we only focused on commonly used ports.
+We repeat our tests on a few ports:
+\begin{itemize}
+    \item 80, 443 - HTTP and HTTPS
+    \item 993 - Secure IMAP port
+    \item 5228 - Google cloud messaging port
+    \item 6969 - random port number
+    \item 8000 - common HTTP proxy port number
+\end{itemize}
 
-In addition to noting that our traffic was not encrypted but only sent to a specific destination port, it is also important to say that an increasing proportion of traffic is served over SSL.
-A commerciall deployed opt-in HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP, although varying with time~\footnote{http://db.awazza.com/users/global/}.
+We have only observed port-specific behavior on two networks: \emph{Cellular 2} and \emph{Cellular 4}.
+In the first case, traffic travels over SSL ports (443, 993) unmodified except for packets with invalid checksum being dropped.
+Our traffic was not encrypted and an increasing proportion of traffic is served over SSL.
+A commercially deployed HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP\footnote{\url{http://db.awazza.com/users/global/}}, making the value of such proxying questionable.
 
-All other traffic, however, is intercepted at the middlebox which we have separately confirmed to be a Bytemobile proxy~\footnote{Citrix Bytemobile - is it appropriate to reference this case accurately?} and it terminates clients' connections, establishing new ones to the destination, as well as adding options not present in original packets, changing window size, assigning new source port numbers sequentially and even postponing handshaking until the handshake with the client is complete.
+We confirmed the middlebox to be an optimizing proxy (ByteMobile).
+It replaces the original connections with a new one which adds options not present in original packets, changes window size and resets other fields to new values.
+Such behavior makes deployment of any protocol modifications or extensions very difficult (e.g.
+we verified that TCPCrypt does not work on this network) and although we did not see more cases of it, the company claims over 160 clients.
+
+In the second case, traffic was only modified when going to ports 80 and 443. As we discuss below, the network optimizes HTTP traffic and since our traffic does not resemble valid HTTP, it is dropped.
+
+\subsection{Non-transparent proxies}
+\label{sec:proxies}
+
+Non-transparent proxies are one of the biggest problems for TCP extension deployment.
+They terminate a connection between an initiator and destination and replaces it with a new one towards the destination, often deliberately changing connection parameters to optimize for certain behavior (\eg ByteMobile, HTTP proxies).
+We checked for non-transparent proxies on in our tested networks on a number of popular port numbers:
+
+21, 22, 25, 80, 110,
+  135,
+  139,
+  143,
+  161,
+  443,
+  445,
+  465,
+  585,
+  587,
+  993,
+  995,
+ 1194,
+ 1723,
+ 5060,
+ 6881,
+ 9001
+
+\begin{table}[t!]
+{\small
+\begin{center}
+\begin{tabular}{| l | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{3.7cm} |}
+\hline
+    \textbf{Net ID} & \textbf{Radio Technology} & \textbf{Proxied Ports} \\ \hline \hline
+    Cellular 2      & HSPA                      & 21, 22, 25, 80, 110, 135, 143, 161, 465, 585, 587, 995, 1194, 1723, 5060, 6881, 9001 \\ \hline
+    Cellular 4      & GPRS                      & 80, 110, 143 \\ \hline
+    Cellular 4      & HSPA                      & 80, 110, 143 \\ \hline
+    Cellular 4      & LTE                       & 80, 110, 143, 443, 993, 995 \\ \hline
+    % T-Mobile, part of EE
+    Cellular 4\*     & HSPA                     & 25, 80, 110, 143 \\ \hline
+    % Virgin on T-Mobile
+    Cellular 4\*\*    & HSPA                    & 25, 80, 110, 143 \\ \hline
+\end{tabular}
+\end{center}
+}
+\caption{Proxies}
+\label{tab:proxies}
+\end{table}
+
+Results are summarized in Table~\ref{tab:proxies}. Cellular 2 proxies almost all of our tested ports except for 443 and 993 as observed before. 
+Cellular 4, is a more interesting case since reported proxying is not the same across the whole network.
+It is not obvious why ports used for encrypted communication are proxied over LTE and in our dataset there was only 1 case for HTTPS, but the other two cases seem related to the operator being formed by merging two operators.
+These differences are evident in \emph{whois} records, reported network name (\textbf{\*} shows the old operator name) and MVNO setup (\textbf{\*\*}).
+It is not safe to estimate proxying just from \emph{whois} records either: they do not necessarily match the operator names and we saw the same address ranges used with all Cellular 4 reported network names for HSPA and LTE radios.
+
+Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions.
+On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 (47 sessions) in two cases non-HTTP traffic is filtered on port 80, signaling that there still may be some proxying within the networks.
+Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but allow non-HTTP traffic.
+
+We did not have data about Cellular 6, but results discussed above suggest that there are no non-transparent proxies present.
+All other networks did not show non-transparent proxy behavior, but did not necessarily allow for TCP modifications either.
+In addition, the tests on port 80 show that a number of networks contain middleboxes that optimize HTTP, but such protocol-specific behavior is not uniform even within a single network making it more difficult to implement any changes.
+
+%    an_op_name    | an_cell_net_type | an_country | port |    ip_global (simplified to network address. Included the owner according to whois
+% db)
+% -----------------+------------------+------------+------+-----------------
+%  EE              | GPRS             | gb         |   80 | 31.72/19 (EE)
+%  EE              | GPRS             | gb         |  110 | 31.72/19 (EE)
+%  EE              | GPRS             | gb         |  143 | 31.72/19 (EE) 
+%  EE              | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE)
+%  EE              | HSPA             | gb         |   80 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
+%  EE              | HSPA             | gb         |  110 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
+%  EE              | HSPA             | gb         |  143 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
+%  EE              | LTE              | gb         |   80 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
+%  EE              | LTE              | gb         |  110 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
+%  EE              | LTE              | gb         |  143 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
+%  EE              | LTE              | gb         |  443 | 213.205.224/12 (EE)
+%  EE              | LTE              | gb         |  993 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
+%  EE              | LTE              | gb         |  995 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
+%  T-Mobile        | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile)
+%  T-Mobile        | HSPA             | gb         |   80 | 31.96/21 (EE), 149.254/16 (T-Mobile)
+%  T-Mobile        | HSPA             | gb         |  110 | 31.96/21 (EE), 149.254/16 (T-Mobile)
+%  T-Mobile        | HSPA             | gb         |  143 | 31.96/21 (EE), 149.254/16 (T-Mobile)
+%  Virgin          | HSPA             | gb         |   25 | 149.254/16 (T-Mobile)
+%  Virgin          | HSPA             | gb         |   80 | 149.254/16 (T-Mobile)
+%  Virgin          | HSPA             | gb         |  110 | 149.254/16 (T-Mobile)
+%  Virgin          | HSPA             | gb         |  143 | 149.254/16 (T-Mobile)
+%  giffgaff        | HSPA             | gb         |   21 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |   22 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |   25 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |   80 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  110 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  135 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  143 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  161 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  465 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  585 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  587 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         |  995 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         | 1194 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         | 1723 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         | 5060 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         | 6881 | 82.132/16 (O2)
+%  giffgaff        | HSPA             | gb         | 9001 | 82.132/16 (O2)
 
 \section{Protocol negotiation}
+\label{sec:protocol}
 
-What is common to all candidate protocol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information.
-Such information must instead be exchanged through other means - either implicit to the opcode, or explicit by redefining the meaning of part of the data payload or previously exchanged information.
-We therefore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.
+What is common to all candidate protocol changes described above is that they are very limited in size and therefore sufficient to exchange an opcode or an ID.
+Other protocol details must instead be exchanged either out of band, implicit to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information.
+
+\subsection{Opcode embedding}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figs/tcp-header}
+% \vspace{-2mm}
+\caption{TCP Header}
+\label{fig:header}
+% \vspace{-4mm}
+\end{figure}
+
+A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections.
+In addition, it is a 32-bit value and we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant.
+The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Fig.~\ref{fig:header}), carrying the desired opcode.
+For simultaneous open we can use the same mechanism in both directions.
+
+For acknowledging extensions in SYN-ACK packet we use Urgent Pointer and Checksum fields.
+Our tests indicate that neither choice is guaranteed to succeed, therefore we combine the two by redefining their values when SYN=1.
+We define Urgent Pointer to be $Opcode_U$ when URG=0 and Checksum to carry a meaning when Urgent Pointer = 0 to get around the cases where one is blocked but the other one is not.
+Across all networks we have tested in our preliminary study at least one of the choices succeeds.
+
+Checksum value can not be used directly as an opcode since it changes at every level of NAT.
+The traditional NAT does a simple recalculation of the checksum~\cite{Egevang:tu}: subtracts the original source address and port number (in some cases also the sequence number) and adds the new values.
+The received Checksum after \(n\) NAT hops is therefore
+
+\begin{align*}
+Checksum_2 & = Checksum_1 - src_1 + src_2 \\
+Checksum_i & = Checksum_{i-1} - src_{i-1} + src_i \\
+ ... \\
+Checksum_n & = Checksum_1 - src_1 + src_n
+\end{align*}
+
+Therefore we encode the Opcode as a target checksum so that it is recoverable by a simple binary subtraction of source port number and address and sequence numbers when they are remapped by the NAT (3-7 16-bit operations in total):
+
+\begin{align*}
+Checksum_{target} & = Opcode + src_1 \\
+Checksum_n = Opcode_C & = Opcode + src_n
+\end{align*}
+
+There are two choices for setting checksum to a specific value: invalid checksum or adding padding payload.
+The latter can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, but keeps a packet valid from network's perspective.
+
+For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum, in line with other extensions which reuse part of data payload for option negotiation and maintaining validity of a packet.
+This Checksum Correction field is not delivered to the application.
+TLV-Encoded Control Stream can be present in data part~\cite{Bonaventure:wx} of the SYN-ACK packet after the Checksum Correction field.
+
+Sending data with SYN-ACK packet works on all cellular networks we have tested, but does not on most WiFi.
+However, for all three cases (Urgent pointer and checksum) the information either reaches the destination as intended, or the packet is dropped.
+On all of our tested networks at least one case succeeds if there is no non-transparent proxy.
+
+\subsection{Bootstrapping control stream}
+
+We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream.
+We extend the three-way handshake with fields shown in Fig.~\ref{fig:header}:
+
+\begin{enumerate}
+\item \textbf{SYN}: the active opener sends SYN packet with \emph{$Opcode Namespace$ + $Opcode$} and ACK unset.
+\item \textbf{SYN-ACK}: if passive opener recognizes the namespace and the $Opcode$, it replies with SYN-ACK with $Opcode_U$ = $Opcode$ (URG unset) to acknowledge the Opcode.
+\item If passive opener does not receive an ACK, retransmit with $Opcode_C$ and \emph{Checksum Correction}.
+If that fails as well, retransmit SYN-ACK with an invalid checksum.
+Finally, fall back to vanilla TCP and retransmit a standard SYN-ACK.
+\item \textbf{ACK}: when active opener receives a packet, it checks for $Opcode_U$ (if URG is unset).
+Otherwise, it checks if Checksum = $Opcode_C$ as described above.
+If no Opcode is recoverable, fall back to vanilla TCP and transmit a standard ACK.
+\item Otherwise both openers recognize the $Opcode$. Active opener sends an ACK packet with TLV-encoded control information in packet payload. The control stream has been successfully bootstrapped.
+\end{enumerate}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=.9\columnwidth]{figs/handshake}
+% \vspace{-2mm}
+\caption{Example handshake with extension negotiation}
+\label{fig:handshake}
+% \vspace{-4mm}
+\end{figure}
+
+An example handshake is shown in Fig.~\ref{fig:handshake}.
+In this case a client (active opener) sends a SYN packet as show above.
+The packet traverses a middlebox which does not remove the opcode and reaches the Server (passive opener).
+The Server replies with SYN-ACK with U\_Opcode which is discarded by the middlebox.
+It then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, which reaches the Client.
+The client recovers and verifies the received opcode.
+Both endpoints are now assumed to understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging reception of the $Opcode$ from the Server.
+
+An issue with putting control stream into packet payload is that some middleboxes split ACK packets with data into two: one only acknowledging the received data and the other one with the actual payload.
+It is not an issue with TLV-encoded control stream in the payload, because it will be delivered to the receiver first.
+Hence if the passive opener receives data during or immediately after the handshake but can not decode that as valid TLV-encoded control information, it treats that as application data and falls back to vanilla TCP.
+
+After the first data packets there are potential problems with packet resegmentation: control stream may reach the destination in the middle of a payload of a packet and would be impossible to find without scanning the whole payload.
+Either SCTP style chunking or MPTCP DSS option can be reused to get around the problem.
+We leave the it as an open question.
 
@@
 Furthermore, any special protocol negotiation still needs to be standardized to ensure compatibility at the Internet scale...
+
+The exact protocol semantics are subject to further testing and discussions.
+The main goal of this work is to identify a potential path for negotiating TCP extensions between two endpoints across currently deployed networks and we show how a full control stream for TCP can be bootstrapped by defining or overloading the meaning of underspecified header fields.
+
+\Anote{unfinished}
 
@@
 
+A very recent proposal for a more accurate ECN feedback~\cite{Kuhlewind:2014vd} in uses an additional reserved bit, overloads the meaning of already assigned ECN and NS bits and that of the urgent pointer when URG is not set.\Anote{finish this}
+
 In the context of MPTCP one proposal to add a generic control stream is to map such stream into a separate sequence number space and exchange control data over established subflows, only modifying MPTCP DSS option to use one of the reserved bits to differentiate the control stream from the data stream.
 Arguably the specified options are complex as well as extensible to enable it are precisely because there are very few ways in which TCP itself can be extended.
-the underlying assumption, however, is that theyt MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
+the underlying assumption, however, is that MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
 
@@
 
+\clearpage
 %Only show things we really cite :)
 % \nocite{*}
-{\footnotesize 
+{
+
 \bibliographystyle{abbrv}
-%\footnotesize
+\small 
 \bibliography{references}

### 2014-07-10 paper/paper.tex
more references and discussion; related work on tcp steganography

@@
 \usepackage{amsmath}
+\usepackage{lipsum}
 \usepackage{algorithm}
@@
 \usepackage{listings}
+
+\captionsetup[figure]{labelfont={small},textfont={small}}
 
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{noTCP: extending across middleboxes through unspecified behavior}
+\title{NoTCP: steganopgraphic tricks to bypass middleboxes}
 \ifnum\anon=1
-\author{[Paper: \#XXX]}% \hspace{0.2cm} \ampmtime ]}
+\author{[Paper: \#137]}% \hspace{0.2cm} \ampmtime ]}
 \else
@@
 \begin{abstract}
-TCP has become the de facto transport protocol in the Internet.
-The necessity to extend it as well as the difficulty of doing so has lead researchers to explore new ways of exchanging control information between endpoints.
-In this paper we explore the choices for such exchange without using more TCP options.
-To achieve it we design noTCP - a mechanism which uses TCP header fields in combinations that do not have specified semantics to exchange a small amount of control information between endpoints to bootstrap a control channel.
-We also present a preliminary study which shows the mechanism works across large cellular and WiFi networks.
+TCP is the de facto transport protocol in the Internet.
+However it has many limitations that make it ill suited to the modern world of high mobility and multi-homing.
+Protocol optimizations and filtering performed by middleboxes make extension deployment challenging which has led researchers to explore new ways of exchanging control information between endpoints.
+In this paper we explore steganographic approaches to smuggle information between endpoints.
+We design NoTCP -- a mechanism which uses the TCP header fields in combinations that are currently have no defined semantics to hide a small amount of control information from middleboxes and hence bootstrap a control channel.
+We also present a preliminary study showing how the mechanism works across cellular and WiFi networks.
 \end{abstract}
@@
 Instead, TCP has become the de facto transport protocol of the Internet and development of its extensions has been very challenging.
-Proposals including MPTCP~\cite{Handley:vj} for multi-homed endpoints and TCPCrypt~\cite{Mazieres:uz} for ubiquitous end-to-end traffic encryption are necessary for today's networks but deployment has been slow.
+Proposals including MPTCP~\cite{Handley:vj} for multi-homed endpoints and TCPCrypt~\cite{Mazieres:uz} for ubiquitous end-to-end traffic encryption are necessary for today's networks but deployment has been slow and difficult.
+Previous work on MPTCP overviews the major problems with different types of middleboxes~\cite{Raiciu:2012ur}.
 
@@
 \begin{itemize}
-    \item Setting header fields to specific values without setting the corresponding flags so that they do not have a meaning to legacy implementations.
+    \item Setting header fields to specific values without setting the corresponding flags so they do not have a meaning to legacy implementations.
     \item Assigning one or more of the reserved field bits.
@@
 The tests were run separately and we used it to discover non-transparent proxies.
-This test works by checking if proxies respond to a TCP connection from the client: SYN request is going to a server that is instrumented not to reply to the client and the client would not be able to finish the handshake without a proxy, otherwise there is a proxy.
-To see if proxying depend on the higher-level protocol, it also detects HTTP proxies by checking content and header changes, different IP address seen by the server for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed) and some other aspects.
+We primarily use previously developed technique~\cite{Weaver:RHwbx82O} --  non-responsive server test.
+To see proxying dependence on higher-level protocols we also detect HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed).
 Results are presented in Sec.~\ref{sec:proxies}.
@@
 \begin{center}
-\begin{tabular}{| l | >{\centering\arraybackslash}m{1.2cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
+\begin{tabular}{| l | >{\centering\arraybackslash}m{1.2cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
 \hline
-    \textbf{Net ID} & \textbf{Country} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize Ack \#}  & \textbf{Normalize Urg ptr} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
+    \textbf{Net ID} & \textbf{Country} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop SYN-ACK data} & \textbf{Normalize SYN Ack number}  & \textbf{Normalize Urgent Pointer} & \textbf{Normalize Reserved} & \textbf{Remap Sequence} \\ \hline \hline
     % Univ Helsinki WiFi
-    WiFi edu 1      & FI               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    WiFi edu 1      & Finland               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
     % Eduroam
@@
     % MKSW germany (carlos)
-    WiFi pub 2      & DE               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    WiFi pub 2      & Germany               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     % Virgin home
@@
     % TEO Lt
-    WiFi res 2      & LT               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    WiFi res 2      & Lithuania               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     % Telefonica Free
-    WiFi ent 1      & ES               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    WiFi ent 1      & Spain               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     \hline
     % Finland internet.saunalahti
-    Cellular 1      & FI               &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    Cellular 1      & Finland               &                        &                            &                        &                         &                        &                             &                     \\ \hline
     % GiffGaff
@@
     % E-Plus germany
-    Cellular 3      & DE               &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
+    Cellular 3      & Germany               &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
     % EE/T-Mobile
@@
     % Omnitel LT
-    Cellular 6      & LT               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    Cellular 6      & Lithuania               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
 \end{tabular}
@@
 }
-\caption{Network behavior observed through tests generating custom TCP packets. A dash means that we observed different cases on the same network.}
+\caption{Network behavior observed through tests generating custom TCP packets. A dash means different cases observed based on port numbers.}
 \label{tab:networks}
@@
 
-We have only observed port-specific behavior on two networks: \emph{Cellular 2} and \emph{Cellular 4}.
+We only observed port-specific behavior on two networks: \emph{Cellular 2} and \emph{Cellular 4}.
 In the first case, traffic travels over SSL ports (443, 993) unmodified except for packets with invalid checksum being dropped.
@@
 
-We confirmed the middlebox to be an optimizing proxy (ByteMobile).
-It replaces the original connections with a new one which adds options not present in original packets, changes window size and resets other fields to new values.
+We confirmed the middlebox to be a ByteMobile proxy.
+It replaces the original connections with a new one which adds options not present in original packets, changes window size and resets other fields to new values to optimize TCP for cellular networks~\cite{Ha:2006td}.
 Such behavior makes deployment of any protocol modifications or extensions very difficult (e.g.
-we verified that TCPCrypt does not work on this network) and although we did not see more cases of it, the company claims over 160 clients.
+we verified that TCPCrypt does not work on this network).
+Although we did not see more cases of it, the company claims over 160 clients.
 
@@
 
-Non-transparent proxies are one of the biggest problems for TCP extension deployment.
-They terminate a connection between an initiator and destination and replaces it with a new one towards the destination, often deliberately changing connection parameters to optimize for certain behavior (\eg ByteMobile, HTTP proxies).
-We checked for non-transparent proxies on in our tested networks on a number of popular port numbers:
-
-21, 22, 25, 80, 110,
-  135,
-  139,
-  143,
-  161,
-  443,
-  445,
-  465,
-  585,
-  587,
-  993,
-  995,
- 1194,
- 1723,
- 5060,
- 6881,
- 9001
+Since proxies make new changes very difficult, we checked for non-transparent oes on in our tested networks on a number of popular port numbers:
+21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.
 
@@
     % T-Mobile, part of EE
-    Cellular 4\*     & HSPA                     & 25, 80, 110, 143 \\ \hline
+    Cellular 4*     & HSPA                     & 25, 80, 110, 143 \\ \hline
     % Virgin on T-Mobile
-    Cellular 4\*\*    & HSPA                    & 25, 80, 110, 143 \\ \hline
+    Cellular 4**    & HSPA                    & 25, 80, 110, 143 \\ \hline
 \end{tabular}
@@
 }
-\caption{Proxies}
+\caption{Non-transparent proxies and proxied port numbers by radio technology}
 \label{tab:proxies}
@@
 Results are summarized in Table~\ref{tab:proxies}. Cellular 2 proxies almost all of our tested ports except for 443 and 993 as observed before. 
-Cellular 4, is a more interesting case since reported proxying is not the same across the whole network.
-It is not obvious why ports used for encrypted communication are proxied over LTE and in our dataset there was only 1 case for HTTPS, but the other two cases seem related to the operator being formed by merging two operators.
-These differences are evident in \emph{whois} records, reported network name (\textbf{\*} shows the old operator name) and MVNO setup (\textbf{\*\*}).
-It is not safe to estimate proxying just from \emph{whois} records either: they do not necessarily match the operator names and we saw the same address ranges used with all Cellular 4 reported network names for HSPA and LTE radios.
+Cellular 4 shows non-uniform behavior across the network.
+It is not obvious why ports commonly used for encrypted communication are proxied over LTE, we only recorded 1 case for HTTPS.
+The other two cases (* and **) seem related to the operator being formed from two merged operators.
+These differences are evident in \emph{whois} records, device-reported network name (\textbf{*} shows the old operator name) and MVNO setup (\textbf{**} setup with the old operator).
+
+It is not safe to estimate proxying just from \emph{whois} records: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios.
+An active test is needed reliably detect proxies.
 
@@
 On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 (47 sessions) in two cases non-HTTP traffic is filtered on port 80, signaling that there still may be some proxying within the networks.
-Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but allow non-HTTP traffic.
+Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic.
+We did not have data about Cellular 6, but results discussed above suggest that there are no non-transparent proxies present.
 
-We did not have data about Cellular 6, but results discussed above suggest that there are no non-transparent proxies present.
-All other networks did not show non-transparent proxy behavior, but did not necessarily allow for TCP modifications either.
-In addition, the tests on port 80 show that a number of networks contain middleboxes that optimize HTTP, but such protocol-specific behavior is not uniform even within a single network making it more difficult to implement any changes.
+All other networks did not show non-transparent proxy behavior, but the ones who did pose a challenge that is not easily solved.
+Proxies potentially change every aspect of the original connection and hiding any information within the packet becomes extremely difficult since any value that a proxy does not recognize is reset.
 
@@
 \centering
+\small{
 \includegraphics[width=\columnwidth]{figs/tcp-header}
 % \vspace{-2mm}
-\caption{TCP Header}
+\caption{NoTCP Header with overloaded fields and extra Checksum Correction / Control Stream fields within payload}
 \label{fig:header}
+}
 % \vspace{-4mm}
@@
 The received Checksum after \(n\) NAT hops is therefore
-
+\vspace{-2mm}
 \begin{align*}
@@
 \end{align*}
+\vspace{-2mm}
 
-Therefore we encode the Opcode as a target checksum so that it is recoverable by a simple binary subtraction of source port number and address and sequence numbers when they are remapped by the NAT (3-7 16-bit operations in total):
-
+We encode the Opcode as a target checksum so that it is recoverable by a simple binary subtraction of source port number, address and sequence numbers in case they are remapped by the NAT (3-7 16-bit operations in total):
+\vspace{-2mm}
 \begin{align*}
@@
 \end{align*}
+\vspace{-2mm}
 
@@
 This Checksum Correction field is not delivered to the application.
-TLV-Encoded Control Stream can be present in data part~\cite{Bonaventure:wx} of the SYN-ACK packet after the Checksum Correction field.
+TLV-Encoded (Type-Length-Value) Control Stream can be present in data part~\cite{Bonaventure:wx} of the SYN-ACK packet after the Checksum Correction field.
 
@@
 
-After the first data packets there are potential problems with packet resegmentation: control stream may reach the destination in the middle of a payload of a packet and would be impossible to find without scanning the whole payload.
+After the first data packets there are potential problems with segment splitting and coalescing : control stream may reach the destination in the middle of a payload of a packet and would be impossible to find without scanning the whole payload.
 Either SCTP style chunking or MPTCP DSS option can be reused to get around the problem.
-We leave the it as an open question.
+The obvious drawback of such choice is that control stream becomes subject to flow control.
+We leave refinement of the full control channel for future work.
 
@@
 
-The main points of discussion on \emph{noTCP} is the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
-Furthermore, any special protocol negotiation still needs to be standardized to ensure compatibility at the Internet scale...
+The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
+We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.
+It is an extremely hard sell for standards bodies.
+However the main motivation of our work is precisely that options are no longer a usable choice for new extensions hence such position is justified.
 
 The exact protocol semantics are subject to further testing and discussions.
-The main goal of this work is to identify a potential path for negotiating TCP extensions between two endpoints across currently deployed networks and we show how a full control stream for TCP can be bootstrapped by defining or overloading the meaning of underspecified header fields.
+The main goal of this work is to identify potential methods for exchanging control information by hiding it from existing middleboxes and we show how a full control stream for TCP can be bootstrapped by overloading the meaning of underspecified header fields.
 
-\Anote{unfinished}
+We also need to extend our study to more networks to make sure that our proposed method of exchanging control information works in diverse scenarios.
+More importantly, we have left open the question of finding a method to hide such information from non-transparent proxies.
+The easiest albeit very limited solution is to identify and cache port numbers that allow unmodified traffic.
+Our goal instead is to study commonly found proxy implementations in detail in order to quickly identify them as well as encode information within their own operating patterns.
 
@@
 \label{sec:related}
+
+Steganography has been used in the past to hide information within network protocols.
+The main focus has been on covert channels that violate system's security policy and initial sequence numbers~\cite{Rowland:1997vq}, TCP Timestamps~\cite{Giffin:2002wh} and combinations with IP flags~\cite{Murdoch:2005fz}.
+There are detection techniques for these methods and our main goal is to exchange a small amount of control information rather than a genuine covert channel, asking for a different design.
 
@@
 
-A very recent proposal for a more accurate ECN feedback~\cite{Kuhlewind:2014vd} in uses an additional reserved bit, overloads the meaning of already assigned ECN and NS bits and that of the urgent pointer when URG is not set.\Anote{finish this}
+AccECN proposal~\cite{Kuhlewind:2014vd} uses an additional reserved bit, overloads the meaning of already assigned ECN~\cite{Floyd:up} and NS~\cite{Ely:uc} bits and redefines Urgent Pointer as Non-Urgent when URG is not set.
+The idea is highly related to ours, but reserving the field solely for ECN use prevents other extensions from using.
+Furthermore, we saw that some networks block packets with such field, making it especially difficult to use for ECN.
 
-In the context of MPTCP one proposal to add a generic control stream is to map such stream into a separate sequence number space and exchange control data over established subflows, only modifying MPTCP DSS option to use one of the reserved bits to differentiate the control stream from the data stream.
-Arguably the specified options are complex as well as extensible to enable it are precisely because there are very few ways in which TCP itself can be extended.
-the underlying assumption, however, is that MPTCP is already deployed as well as that existing deployments will be forwards-compatible with the specification.
+Generic control stream for MPTCP~\cite{Bonaventure:wx} can be reused once it has been negotiated.
+It proposes mapping such stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams.
+The underlying assumption, however, is that MPTCP is already deployed and that existing deployments will be forwards-compatible with the new specification.
 
-Conceptually close to MPTCP's subflows is SCTP's \emph{data chunks}, where a packet can contain multiple TLV (Type-Length-Value) encoded chunks to separate user data and control information.
-Despite its challenged deployment in the Internet, SCTP over DTLS is being used as the basis of WebRTC data protocol~\cite{Tuexen:wv} and we suggest that such chunking could be used on top of \emph{noTCP} once such behavior has been negotiated as described above.
+%More generally SCTP uses \emph{data chunks} to differentiate between different subflows.
+Each packet can contain multiple TLV-encoded chunks to separate user data and control information.
+Despite its challenged deployment in the Internet, SCTP over DTLS is being used as the basis of WebRTC data protocol~\cite{Tuexen:wv} and we suggest that such chunking could be used on top of \emph{noTCP} once control channel has been bootstrapped.
 
@@
 
-\clearpage
+% \clearpage
 %Only show things we really cite :)

### 2014-07-14 paper/paper.tex
toby's intro + rewrote large parts of the paper. Discussion and conclusions with Jon's ideas

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{NoTCP: steganopgraphic tricks to bypass middleboxes}
+\title{NoTCP: steganographic tricks to bypass middleboxes}
 \ifnum\anon=1
@@
 
-Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large.
+TCP is everywhere in the Internet.
+The overwhelming majority of bytes are carried in TCP streams~\cite{John:2007dv}, and TCP is used for all types of traffic from bulk download to video streaming to instant messaging.
+Over the last decade a number of new protocols have been proposed to replace TCP.
+However many middle boxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like.
+Consequently few of these protocols have found any leverage.
+
+One of the main requirements to extend and modify TCP is the ability to exchange control information.
+Traditionally this is what TCP option space was used for.
+However option space is limited, heavily contended~\cite{Handley:vj,Mazieres:uz} and troubled by network middleboxes~\cite{Honda,Guha2005tb,UntoldMiddlebox2011}.
+
+Instead we believe that the solution is to work round the middle boxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header.
+This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that SYN Cookies~\cite{Eddy:2007to} do currently.
+
+A practical method of exchanging control information outside of TCP options could not only speed up deployment of existing TCP extensions (\eg MPTCP~\cite{Handley:vj}, TCPCrypt~\cite{Mazieres:uz}), but also help add a generic control stream within TCP.
+Potential applications range from new secure extensions~\cite{Rescorla:R080FoGB} to exchanging information about endpoints for connection parameter tuning (\eg initial congestion window~\cite{Dukkipati:2010hs}) to designing better coordination for long-lived connections to save mobile device energy~\cite{Aucinas:2013uk}.
+Most importantly, development of new functionality should not be limited by network middleboxes.
+
+This paper makes three main contributions. 
+\begin{enumerate}
+\item We discuss the possible ways to conceal information in the TCP header and give some examples of the extensions this could enable
+\item We describe a measurement application that we have written to run on Android phones.  
+\item We present the results of a small scale deployment of our application.
+\end{enumerate}
+
+Our hope is to motivate the research community into exploring the full range of possibilities in this space and to encourage more people to install our application so that we can build a comprehensive database of the impact of middleboxes across different network providers.
+
+% Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large.
 The problem is caused by the various middleboxes that aim to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems.
@@
 
-In an ideal world all infrastructure is forwards compatible and changes together with the needs of the users.
+% In an ideal world all infrastructure is forwards compatible and changes together with the needs of the users.
 Instead, TCP has become the de facto transport protocol of the Internet and development of its extensions has been very challenging.
@@
 
-Although TCP has been built to be extensible via options and reserved bits, all extensions are primarily designed around the limitations of existing network middleboxes\cite{Honda,Guha2005tb,UntoldMiddlebox2011}, but even if options were not filtered, we have run out of the 40-byte TCP option space.
+% Although TCP has been built to be extensible via options and reserved bits, all extensions are primarily designed around the limitations of existing network middleboxes\cite{Honda,Guha2005tb,UntoldMiddlebox2011}, but even if options were not filtered, we have run out of the 40-byte TCP option space.
 
-Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}.
+% Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}.
 TCPCrypt uses the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}.
@@
 
-We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified.
+% We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified.
 At the same time it is designed to work across most of deployed network infrastructure.
@@
 
-With the focus on cellular and WiFi networks due to their renown middlebox peculiarities, we make the following contributions in this work:
+% With the focus on cellular and WiFi networks due to their renown middlebox peculiarities, we make the following contributions in this work:
 
-\begin{enumerate}
-    \item Review the aspects of the TCP protocol that can potentially be used to extend it (Section~\ref{sec:protocolChanges}).
-    \item Provide an analysis of current mobile (WiFi and cellular) network behavior in the presence of such modifications (Section~\ref{sec:network}).
-    \item Propose a robust method to bootstrap control-stream between two endpoints across middleboxes (Section~\ref{sec:protocol}).
-\end{enumerate}
+% \begin{enumerate}
+%     \item Review the aspects of the TCP protocol that can potentially be used to extend it (Section~\ref{sec:protocolChanges}).
+%     \item Provide an analysis of current mobile (WiFi and cellular) network behavior in the presence of such modifications (Section~\ref{sec:network}).
+%     \item Propose a robust method to bootstrap control-stream between two endpoints across middleboxes (Section~\ref{sec:protocol}).
+% \end{enumerate}
 
@@
 
-Without using the options field and only changing the meaning of header fields in a backwards-compatible manner we only have a few options:
+\begin{figure}[t!]
+\centering
+\small{
+\includegraphics[width=\columnwidth]{figs/tcp-header}
+\vspace{-3mm}
+\caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload}
+\label{fig:header}
+}
+\vspace{-2mm}
+\end{figure}
+
+Without using the options field we only have a few options for control information exchange:
 \begin{itemize}
-    \item Setting header fields to specific values without setting the corresponding flags so they do not have a meaning to legacy implementations.
+    \item Setting header fields to specific values without setting the corresponding flags to have no meaning to legacy implementations.
     \item Assigning one or more of the reserved field bits.
-    \item Defining semantics for specific values of currently valid header fields.
+    \item Defining new semantics for valid header fields.
 \end{itemize}
 
-In this section we overview these different options, discuss their pros and cons and show how real networks interact with them.
+In this section we discuss them and their pros and cons in depth. Fig.~\ref{fig:header} shows TCP header with fields that we propose for reuse and define in Section~\ref{sec:protocol}.
 
@@
 
-The only header fields that may or may not have a meaning depending on the corresponding flag are \emph{urgent pointer} and \emph{acknowledgment number}, and the latter is only allowed to be not set for the initial SYN packet.
+The only header fields that may not have a meaning depending on the corresponding flag are \emph{Urgent Pointer} and \emph{Acknowledgment Number}.
+The latter must always be set after the initial SYN packet.
 
-Urgent pointer is both a good and a bad candidate at the same time: it is already recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be claimed that it is taking up precious space in the TCP header.
-On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) to reset the field as it makes it difficult to track the application-layer data~\cite{seolma}.
-We present our observations of how often the field is filtered in Section~\ref{sec:network}, however although we witnessed the field to be passed through on many of our studied networks and hence it is a usable option, it is not without drawbacks.
+It is recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
+On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
+Better yet, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and can be redefined.
 
-Acknowledgment field, on the other hand, is double the length and we saw it successfully pass through the networks in more cases.
-Nevertheless, unless specific value is signaled out of bound between the two endpoints to use the field with TCP Simultaneous Open, only the initiating end can use it and the responding side needs to acknowledge the understanding of it in some other method.
+Acknowledgment field is only available in the initial SYN packet.
+It must therefore be combined with another mechanism to acknowledge control information unless TCP Simultaneous Open is used.
+Simultaneous open is still viable in most cellular networks~\cite{UntoldMiddlebox2011}, but we also explore other options.
 
@@
 
-The main difficulty with TCP Reserved bits is that there are only 3 bits unallocated if we consider the three ECN bits added over 10 years ago~\cite{Floyd:up,Ely:uc} but still not universally implemented~\cite{Kuhlewind:2013hr}.
-Any further reserved bit allocation would therefore be extremely difficult.
+Only 3 TCP Reserved bis have been left unallocated with the three ECN bits (ECN, CWR and experimental NS)~\cite{Floyd:up,Ely:uc} that are still not universally implemented~\cite{Kuhlewind:2013hr}.
 
-Nevertheless, there are proposals that make use of them.
-One of the proposals for reducing Web latency~\cite{Flach:2013uy} suggests sending multiple copies of packets, but to avoid triggering duplicate acknowledgments add a reserved bit to the header.
-We confirmed the authors observation that middleboxes do not seem to discard packets with non-compliant TCP flags, however we also saw a number of middleboxes to clear such flags, hence interfering with the duplicate acknowledgment of the proposed scheme.
-Due to the problem with reserved bit allocation we would recommend using mechanisms we discuss later.
+Due to simplicity and non-interference with legacy implementations there are proposals that make use of them.
+One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets.
+To avoid triggering duplicate acknowledgments reserved bit is used to mark such packets as duplicate and should be ignored by fast retransmit.
 
@@
 
-Our remaining class of options, and by far the biggest, is defining protocol semantics based on specific values of certain header fields.
-It would necessitate generating crafting packets with specific values for initial sequence numbers, window size or a specific checksum value, and we posit that it would be an acceptable choice if such solution is deployable in today's network and can be implemented efficiently.
+The remaining choice is to encode information within valid header fields.
+It involves generating packets with specific values for initial sequence numbers, window size or checksums.
+We posit that it would be an acceptable choice as long as it is deployable in today's network.
 
-Initial sequence numbers must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security which has been problematic~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.
+Initial sequence numbers must be hard to predict as they provide a measure (however weak) of security and until strong encryption, \eg TCPCrypt, is universally adopted.
+Adding any information to the field in a way that is easily decoded by the receiver makes it predictable and weakens security  which has already been problematic~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.
 
-Window size field is potentially a better candidate, especially if used only during connection setup phase, since data is not typically transmitted during the handshake.
-We could further claim that even during data exchange a few of the least-significant bits could be sacrificed for carrying other information, however that would prevent flow control from operating properly.
-Higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of signaled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links.
-Using the field risks interfering both with TCP flow control mechanism as well as window-size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed \eg by Citrix ByteMobile Proxy~\cite{Ha:2006td}.
+Window size is potentially a better candidate, especially if used during connection setup phase since data is not typically transmitted during the handshake.
+Using lower-order bits for control information would prevent flow control from operating properly, but higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of advertised window size.
+Using the field risks interfering both with TCP flow control mechanism as well as window size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed \eg by Citrix ByteMobile Proxy~\cite{Ha:2006td}.
 
-There are multiple benefits of using the checksum field.
-Firstly, it is only used for error detection as opposed to other fields that govern protocol behavior and is either not always necessary due to the layered design of the OSI stack (when error detection is done at the layer below or above), or can be replaced with extensions' separate (MPTCP DSS checksum) or even stronger (TCPCrypt MAC) mechanisms.
+There are multiple benefits to using the checksum field.
+Firstly, it is only used for error detection.
+It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown unreliable~\cite{Stone:2000fc}.
+Instead, it may be replaced with extensions' separate, potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code) or left for the upper layer.
+
 Secondly, checksum operations are efficient as they only involve one's complement addition.
-And finally, it is possible to generate checksums in a way that the intended value can recovered by the receiving end after traversing common types of middleboxes that recompute the checksum as we discuss in the following sections.
+We show later in the paper (Section~\ref{sec:protocol}) how checksums can be generated in a way that the intended value is recovered by the receiver after traversing multiple layers of Network Address Translation (NAT).
 
@@
 
-To test what works on current networks we have developed a simplified implementation of TCP that exchanges packets in the standard order (three-way handshake, data exchange and connection shutdown) but with various header fields modified to our purposes.
+To verify current network behavior we have developed an Android application distributed through the \emph{Google Play Store} to collect data from volunteers.
+It contains a simplified implementation of TCP to generate packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown.
+It allows us to modify various header fields to our purposes and check the values received by the other endpoint.
 
-The test has been implemented as an Android application and distributed through the \emph{Google Play Store} to collect data from volunteers.
-The overall structure of the tool is relatively straightforward:
+The testing setup is:
 \begin{itemize}
-    \item UI which reports aggregated results to the end-users, sends results to the reporting server and manages the overall flow of the tests.
-    \item Native implementation of the simplified TCP stack using Raw Unix sockets (\emph{SOCK\_RAW}) to allow us to generate packets at the network layer.
-This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
-    \item Server implementation that generates specific answers to incoming packets based on header and payload values.
+    \item Android UI to launch tests, report aggregated results to the end-users, send results to the reporting server and manage the tests.
+    \item Simplified TCP stack using Raw Unix sockets for full control over the TCP header (\emph{SOCK\_RAW}).
+%This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
+    \item Server instrumented to generate specific responses to incoming packets based on header and payload values.
+We currently use a dedicated test server in a single location and no firewalls or NATs.
 \end{itemize}
 
-Potential limitations of our methodology include that we do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together.
-To make sure that it does not happen all our packets contain minimal amount of payload (up to 16 bytes during data exchange) and data exchange only consists of a single message sent each way, in addition to any extra acknowledgments generated by middleboxes.
-Due to the same reason we also do not take into account congestion control.
+Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks.
+The tests were run separately, using previously developed techniques~\cite{Weaver:RHwbx82O}.
+The primary one was non-responsive server test: the server is instrumented not to respond to SYN, therefore if connection succeeds it is to a proxy on the path.
+We repeated these tests on different destination ports and found that proxying can be specific to certain port numbers (Section~\ref{sec:portspec})
 
-Instead of trying to infer which source port number would be assigned to a new connection by the kernel network stack without using it, we randomized the selection within the unprivileged port range.
-It was a simple optimization that works on any device and minimizes the probability of multiple tests failing due to another application using a particular port.
-We also repeated the tests on different destination ports, which we discuss in Sec.~\ref{sec:portspec}.
+Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed).
+We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.
+Results are presented in Section~\ref{sec:proxies}.
 
-Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks.
-The tests were run separately and we used it to discover non-transparent proxies.
-We primarily use previously developed technique~\cite{Weaver:RHwbx82O} --  non-responsive server test.
-To see proxying dependence on higher-level protocols we also detect HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed).
-Results are presented in Sec.~\ref{sec:proxies}.
+\subsection{Limitations}
+
+Potential limitations of our methodology include that we do not consider packet splitting and coalescing.
+Our goal is to exchange control information within packet headers that correct behavior does not depend on how packets are resegmented.
+To ensure that it does not happen during our tests we do not send more than a few bytes of data and the exchange only consists of a single message sent each way.
+Due to the same reason we do not take congestion control into account either.
+
+Raw sockets operate in parallel to the kernel network stack.
+We had to suppress reset packets generated by the kernel in response to TCP packets from sessions it is not aware of.
+We also had to choose local port numbers unused by the kernel.
+We chose them randomly from the unprivileged port range and relied on other error detection mechanisms to prevent invalid packets from being delivered in the unlikely case of coinciding connection parameters.
 
@@
 A bigger issue we have discovered is that middleboxes do not always respect the standard and simply drop SYN-ACK packets with payload.
-We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:
+We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}.
 
-\begin{quotation}
-    Although these examples do not show connection synchronization using data-carrying segments, this is perfectly legitimate
-\end{quotation}
+% \begin{quotation}
+%     Although these examples do not show connection synchronization using data-carrying segments, this is perfectly legitimate
+% \end{quotation}
 
@@
 }
+\vspace{-2mm}
 \caption{Non-transparent proxies and proxied port numbers by radio technology}
 \label{tab:proxies}
+\vspace{-4mm}
 \end{table}
@@
 The other two cases (* and **) seem related to the operator being formed from two merged operators.
-These differences are evident in \emph{whois} records, device-reported network name (\textbf{*} shows the old operator name) and MVNO setup (\textbf{**} setup with the old operator).
+These differences are evident in \emph{whois} records and device-reported network name --- \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO), which rent network infrastructure from the mobile network operator.
 
@@
 Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions.
-On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 (47 sessions) in two cases non-HTTP traffic is filtered on port 80, signaling that there still may be some proxying within the networks.
+While other networks did not interfere with our tests, they do not appear to be completely free of proxies.
+On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80.
 Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic.
@@
 
-\section{Protocol negotiation}
+\section{Control information exchange}
 \label{sec:protocol}
@@
 Other protocol details must instead be exchanged either out of band, implicit to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information.
+Here we propose one way of exchanging control information as an opcode to bootstrap a full control stream.
 
 \subsection{Opcode embedding}
-
-\begin{figure}[t!]
-\centering
-\small{
-\includegraphics[width=\columnwidth]{figs/tcp-header}
-% \vspace{-2mm}
-\caption{NoTCP Header with overloaded fields and extra Checksum Correction / Control Stream fields within payload}
-\label{fig:header}
-}
-% \vspace{-4mm}
-\end{figure}
 
@@
 The received Checksum after \(n\) NAT hops is therefore
-\vspace{-2mm}
+
+\vspace{-4mm}
 \begin{align*}
-Checksum_2 & = Checksum_1 - src_1 + src_2 \\
-Checksum_i & = Checksum_{i-1} - src_{i-1} + src_i \\
+Checksum_2 & = Checksum_1 - header_1 + header_2 \\
+Checksum_i & = Checksum_{i-1} - header_{i-1} + header_i \\
  ... \\
-Checksum_n & = Checksum_1 - src_1 + src_n
+Checksum_n & = Checksum_1 - header_1 + header_n
 \end{align*}
-\vspace{-2mm}
 
-We encode the Opcode as a target checksum so that it is recoverable by a simple binary subtraction of source port number, address and sequence numbers in case they are remapped by the NAT (3-7 16-bit operations in total):
-\vspace{-2mm}
+Here $header$ refers to the fields of the header that change, namely the destination port number, address and sequence number.
+Previous study~\cite{UntoldMiddlebox2011} found no NATs mangling sequence numbers, but we witnessed a few cases of such behavior.
+The Opcode can then be encoded as a target checksum so that it is recoverable by a simple binary subtraction of the changing fields (3 to 7 16-bit arithmetic operations in total):
+
+\vspace{-4mm}
 \begin{align*}
-Checksum_{target} & = Opcode + src_1 \\
-Checksum_n = Opcode_C & = Opcode + src_n
+Checksum_{target} & = Opcode + header_1 \\
+Checksum_n = Opcode_C & = Opcode + header_n
 \end{align*}
-\vspace{-2mm}
 
@@
 
-Sending data with SYN-ACK packet works on all cellular networks we have tested, but does not on most WiFi.
-However, for all three cases (Urgent pointer and checksum) the information either reaches the destination as intended, or the packet is dropped.
-On all of our tested networks at least one case succeeds if there is no non-transparent proxy.
+Sending data with SYN-ACK packet works on all tested cellular networks, but does not on most WiFi.
+For the above header modifications the information either reaches the destination as intended, or the packet is dropped.
+Finally, on all of our tested networks at least one case succeeds as long as there is no non-transparent proxy for the chosen destination port.
 
@@
 \label{fig:handshake}
-% \vspace{-4mm}
+\vspace{-4mm}
 \end{figure}
@@
 
-An issue with putting control stream into packet payload is that some middleboxes split ACK packets with data into two: one only acknowledging the received data and the other one with the actual payload.
+Putting control stream into packet payload suffers from much the same issues as other proposals to extend TCP option space~\cite{Ramaiah:2012wa}.
+The premise of our work is that middleboxes should not be aware of the new control streams in the first place and therefore not interfere with them.
+
+We witnessed a case of resegmentation where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload.
 It is not an issue with TLV-encoded control stream in the payload, because it will be delivered to the receiver first.
-Hence if the passive opener receives data during or immediately after the handshake but can not decode that as valid TLV-encoded control information, it treats that as application data and falls back to vanilla TCP.
+The mechanism of \emph{Checksum Correction} would, however, be compromised for the first segment.
+If the passive opener receives data during or immediately after the handshake but can not decode that as valid TLV-encoded control information, it treats that as application data and falls back to vanilla TCP.
 
@@
 We leave refinement of the full control channel for future work.
-
-\section{Discussion}
-
-The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
-We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.
-It is an extremely hard sell for standards bodies.
-However the main motivation of our work is precisely that options are no longer a usable choice for new extensions hence such position is justified.
-
-The exact protocol semantics are subject to further testing and discussions.
-The main goal of this work is to identify potential methods for exchanging control information by hiding it from existing middleboxes and we show how a full control stream for TCP can be bootstrapped by overloading the meaning of underspecified header fields.
-
-We also need to extend our study to more networks to make sure that our proposed method of exchanging control information works in diverse scenarios.
-More importantly, we have left open the question of finding a method to hide such information from non-transparent proxies.
-The easiest albeit very limited solution is to identify and cache port numbers that allow unmodified traffic.
-Our goal instead is to study commonly found proxy implementations in detail in order to quickly identify them as well as encode information within their own operating patterns.
 
@@
 TCP implements an \emph{urgent mechanism} at its core that allows the sending user to stimulate the receiving user to accept some \emph{urgent data}.
-The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is specified that it is not a mechanism for sending such data.
+The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is explicitly not designed as a mechanism for sending such data.
 Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and in general it is recommended against the use of the mechanism~\cite{Gont:2011vi}.
@@
 
-Generic control stream for MPTCP~\cite{Bonaventure:wx} can be reused once it has been negotiated.
-It proposes mapping such stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams.
-The underlying assumption, however, is that MPTCP is already deployed and that existing deployments will be forwards-compatible with the new specification.
+Generic control stream for MPTCP~\cite{Bonaventure:wx} can be reused as well.
+It proposes the mapping control stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams.
+The underlying assumption, however, is that MPTCP is already deployed and that existing deployments will be compatible with the new specification.
 
@@
 
+\section{Discussion}
+
+The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
+We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.
+
+The exact protocol semantics are subject to further testing and discussions.
+The main goal of this work is to identify potential methods for exchanging control information by hiding it from existing middleboxes.
+
+In the long term and beyond our example control information exchange we believe that each protocol-relate standards (TCP extensions in this case) should explicitly define which header fields it interferes with and how.
+Examples include extending sequence numbers and using stronger message integrity checks~\cite{Mazieres:uz}, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple checksum or window size modifications.
+
+More importantly, we have left open the question of finding a method to hide such information from non-transparent proxies.
+The easiest, albeit very limited solution is to identify and cache port numbers that allow unmodified traffic.
+Our goal instead is to study commonly found proxy implementations in detail in order to quickly identify them and encode information within their own operating patterns.
+
+The final goal is to map header field modifications by the various extensions as well as network middleboxes into an algebraic form as we have done in the simple case of checksum.
+This way we would not only be able to exchange control information across the network reliably, but also identify network path behavior by looking at its performed operations.
+This way extending a protocol would become a matter of defining a new set of such header algebra rules and simplify the process of design and testing.
+
 \section{Conclusion}
-TBD
-%\input{conclusion}
+
+We hope to motivate the research community to explore the full range of possibilities in extending existing protocols.
+We have developed an Android application to collect data about how network middleboxes interfere with packet header fields that may not currently have a meaning.
+We encourage more people to install our application to build a comprehensive database of network behavior.
+The goal of the study is to explore new ways of exchanging control information resilient to major classes of currently deployed middleboxes by using steganographic approaches.
+
+Finally, we would like to use the results of the study to develop a set of rules for how packet headers are modified by middleboxes as well as standard protocol extensions.
+If this were true, extending a protocol could become near automatic.
 

### 2014-07-14 paper/paper.tex
cleaned up references

@@
 
-\vspace{-4mm}
+% \vspace{-2mm}
 \begin{align*}
@@
 
-After the first data packets there are potential problems with segment splitting and coalescing : control stream may reach the destination in the middle of a payload of a packet and would be impossible to find without scanning the whole payload.
-Either SCTP style chunking or MPTCP DSS option can be reused to get around the problem.
-The obvious drawback of such choice is that control stream becomes subject to flow control.
-We leave refinement of the full control channel for future work.
+The focus of our work is to show the possibility of exchanging control information within the packet header. We therefore leave refinement of the full control channel for future work.
 
@@
 
-% \clearpage
+\clearpage
 %Only show things we really cite :)

### 2014-07-15 paper/paper.tex
another iteration, Narseo's comments

@@
 Protocol optimizations and filtering performed by middleboxes make extension deployment challenging which has led researchers to explore new ways of exchanging control information between endpoints.
-In this paper we explore steganographic approaches to smuggle information between endpoints.
+In this paper we explore steganographic approaches to smuggle information between endpoints to avoid middlebox interference.
 We design NoTCP -- a mechanism which uses the TCP header fields in combinations that are currently have no defined semantics to hide a small amount of control information from middleboxes and hence bootstrap a control channel.
@@
 
-TCP is everywhere in the Internet.
-The overwhelming majority of bytes are carried in TCP streams~\cite{John:2007dv}, and TCP is used for all types of traffic from bulk download to video streaming to instant messaging.
+TCP is everywhere in the Internet and it carries the overwhelming majority of bytes~\cite{John:2007dv}.
+TCP is used for all types of traffic from bulk download to video streaming to instant messaging.
 Over the last decade a number of new protocols have been proposed to replace TCP.
@@
 Consequently few of these protocols have found any leverage.
+
+Modern networks necessitate new functionality, such as Multipath Connectivity~\cite{Handley:vj} for multi-homed devices and ubiquitous traffic encryption~\cite{Mazieres:uz}).
+The applications of exchanging control information between endpoints range from new secure extensions~\cite{Rescorla:R080FoGB} to exchanging information about endpoints for connection parameter tuning (\eg initial congestion window~\cite{Dukkipati:2010hs}) to designing better coordination for long-lived connections to save mobile device energy~\cite{Aucinas:2013uk}.
 
@@
 
-Instead we believe that the solution is to work round the middle boxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header.
+We believe that the solution is to work round the middle boxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header.
 This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that SYN Cookies~\cite{Eddy:2007to} do currently.
-
-A practical method of exchanging control information outside of TCP options could not only speed up deployment of existing TCP extensions (\eg MPTCP~\cite{Handley:vj}, TCPCrypt~\cite{Mazieres:uz}), but also help add a generic control stream within TCP.
-Potential applications range from new secure extensions~\cite{Rescorla:R080FoGB} to exchanging information about endpoints for connection parameter tuning (\eg initial congestion window~\cite{Dukkipati:2010hs}) to designing better coordination for long-lived connections to save mobile device energy~\cite{Aucinas:2013uk}.
-Most importantly, development of new functionality should not be limited by network middleboxes.
 
@@
 \item We discuss the possible ways to conceal information in the TCP header and give some examples of the extensions this could enable
-\item We describe a measurement application that we have written to run on Android phones.  
-\item We present the results of a small scale deployment of our application.
+\item We describe a measurement application that we have written to run on Android phones for a large-scale study.  
+\item We present the results of an initial small scale deployment of our application.
 \end{enumerate}
 
-Our hope is to motivate the research community into exploring the full range of possibilities in this space and to encourage more people to install our application so that we can build a comprehensive database of the impact of middleboxes across different network providers.
-
-% Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large.
-The problem is caused by the various middleboxes that aim to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems.
-Even congestion control has required multiple extensions since protocol's inception and the large number of proposed congestion control schemes~\cite{5462976} suggests that more will be necessary.
-
-% In an ideal world all infrastructure is forwards compatible and changes together with the needs of the users.
-Instead, TCP has become the de facto transport protocol of the Internet and development of its extensions has been very challenging.
-Proposals including MPTCP~\cite{Handley:vj} for multi-homed endpoints and TCPCrypt~\cite{Mazieres:uz} for ubiquitous end-to-end traffic encryption are necessary for today's networks but deployment has been slow and difficult.
-Previous work on MPTCP overviews the major problems with different types of middleboxes~\cite{Raiciu:2012ur}.
-
-% Although TCP has been built to be extensible via options and reserved bits, all extensions are primarily designed around the limitations of existing network middleboxes\cite{Honda,Guha2005tb,UntoldMiddlebox2011}, but even if options were not filtered, we have run out of the 40-byte TCP option space.
-
-% Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}.
-TCPCrypt uses the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}.
-An alternative way of looking at the problem is to hide the information from the middleboxes as done by steganographic approaches~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz}.
-
-% We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified.
-At the same time it is designed to work across most of deployed network infrastructure.
-At the core of the idea is modification of TCP header fields in a way that is not interfered with by middleboxes or so that such interference can be safely recovered or ignored.
-The idea builds up from the work on steganography using TCP~\cite{Frczek:2012dl,Murdoch:2005fz}, but we shift the focus from using sequence numbers and options to other fields as well.
-
-% With the focus on cellular and WiFi networks due to their renown middlebox peculiarities, we make the following contributions in this work:
-
-% \begin{enumerate}
-%     \item Review the aspects of the TCP protocol that can potentially be used to extend it (Section~\ref{sec:protocolChanges}).
-%     \item Provide an analysis of current mobile (WiFi and cellular) network behavior in the presence of such modifications (Section~\ref{sec:network}).
-%     \item Propose a robust method to bootstrap control-stream between two endpoints across middleboxes (Section~\ref{sec:protocol}).
-% \end{enumerate}
+We use our initial study to design a way of exchanging a small amount of control information successfully across all of our tested networks and show a potential mechanism to use it for bootstrapping a full control stream currently missing from TCP.
+We hope that our successful results will motivate the research community into exploring the full range of possibilities in this space.
 
@@
 
-Without using the options field we only have a few options for control information exchange:
+Figure~\ref{fig:header} shows TCP header with fields that we propose for reuse and define later in the paper.
+However, the structure of TCP header is very simple and gives are few choices for hiding extra control information:
 \begin{itemize}
-    \item Setting header fields to specific values without setting the corresponding flags to have no meaning to legacy implementations.
+    \item Header fields without corresponding flags to have no meaning to legacy implementations.
     \item Assigning one or more of the reserved field bits.
@@
 \end{itemize}
-
-In this section we discuss them and their pros and cons in depth. Fig.~\ref{fig:header} shows TCP header with fields that we propose for reuse and define in Section~\ref{sec:protocol}.
 
@@
 
-It is recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
+The general recommendation is to not use It is recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
 On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
@@
 
-Only 3 TCP Reserved bis have been left unallocated with the three ECN bits (ECN, CWR and experimental NS)~\cite{Floyd:up,Ely:uc} that are still not universally implemented~\cite{Kuhlewind:2013hr}.
+Only 3 TCP Reserved bits have not been allocated since the three ECN bits (ECN, CWR and experimental NS)~\cite{Floyd:up,Ely:uc} that are still not universally implemented~\cite{Kuhlewind:2013hr}.
 
@@
 One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets.
-To avoid triggering duplicate acknowledgments reserved bit is used to mark such packets as duplicate and should be ignored by fast retransmit.
+The solution avoids triggering fast retransmit by such duplicate acknowledgments by marking them as duplicates in a reserved flag.
 
@@
 It involves generating packets with specific values for initial sequence numbers, window size or checksums.
-We posit that it would be an acceptable choice as long as it is deployable in today's network.
+Arguably it is an acceptable choice if deployable in today's networks.
 
-Initial sequence numbers must be hard to predict as they provide a measure (however weak) of security and until strong encryption, \eg TCPCrypt, is universally adopted.
-Adding any information to the field in a way that is easily decoded by the receiver makes it predictable and weakens security  which has already been problematic~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.
+Initial sequence numbers must be hard to predict as they provide a measure of security and until strong encryption.
+Stronger security measures such as TCPCrypt are better suited for this purpose, but are not universally adopted.
+Adding information to the field to be easily decoded by the receiver makes it predictable and weakens already problematic security~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.
 
@@
 It allows us to modify various header fields to our purposes and check the values received by the other endpoint.
+We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.
 
@@
 \begin{itemize}
-    \item Android UI to launch tests, report aggregated results to the end-users, send results to the reporting server and manage the tests.
-    \item Simplified TCP stack using Raw Unix sockets for full control over the TCP header (\emph{SOCK\_RAW}).
-%This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
+    \item Android application with a simplified TCP stack using Raw Unix sockets for full control over the TCP header (\emph{SOCK\_RAW}). It manages the tests and sends results back to our servers. 
     \item Server instrumented to generate specific responses to incoming packets based on header and payload values.
@@
 Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks.
-The tests were run separately, using previously developed techniques~\cite{Weaver:RHwbx82O}.
+We implemented previously developed techniques~\cite{Weaver:RHwbx82O} and run the test separately.
 The primary one was non-responsive server test: the server is instrumented not to respond to SYN, therefore if connection succeeds it is to a proxy on the path.
-We repeated these tests on different destination ports and found that proxying can be specific to certain port numbers (Section~\ref{sec:portspec})
+We repeated these tests on different destination ports and found that proxying can be specific to certain port numbers.
 
-Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed).
-We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.
-Results are presented in Section~\ref{sec:proxies}.
+%Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic and whether non-HTTP traffic on HTTP port is filtered.
 
@@
 Potential limitations of our methodology include that we do not consider packet splitting and coalescing.
-Our goal is to exchange control information within packet headers that correct behavior does not depend on how packets are resegmented.
-To ensure that it does not happen during our tests we do not send more than a few bytes of data and the exchange only consists of a single message sent each way.
+Our goal is to exchange control information within packet headers to avoid dealing with it altogether.
+To ensure that splitting does not happen during our tests we do not send more than a few bytes of data and the exchange only consists of a single message sent each way.
 Due to the same reason we do not take congestion control into account either.
@@
 We also had to choose local port numbers unused by the kernel.
-We chose them randomly from the unprivileged port range and relied on other error detection mechanisms to prevent invalid packets from being delivered in the unlikely case of coinciding connection parameters.
+We chose them randomly from the unprivileged port range and relied on other error detection mechanisms preventing invalid packet delivery to the applications.
 
@@
 }
+\vspace{-3mm}
 \caption{Network behavior observed through tests generating custom TCP packets. A dash means different cases observed based on port numbers.}
 \label{tab:networks}
+\vspace{-2mm}
 \end{table*}
@@
 We show our collected results in Table~\ref{tab:networks}.
-Since we were focusing on mobile applications, in our preliminary study we collected results across a few public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries: UK, Finland, Germany and Lithuania.
+Since we were focusing on mobile applications, in our preliminary study we collected results across public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries.
 
-The results include which networks did not filter any modifications (one WiFi and two cellular nets), and which ones discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers.
+The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and ones that discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers.
 Only one network provided clients with a global IP (WiFi edu 2).
 
-In traditional NATs~\cite{Egevang:tu} checksum recalculation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use.
-The networks that deliver packets with invalid checksum only do that as otherwise a packet with invalid checksum would be dropped.
+Traditional NATs~\cite{Egevang:tu} perform a simple checksum recalculation: subtract the old header fields from the checksum and add the new ones.
+The networks that deliver packets with invalid checksum only do that.
+Otherwise they would detect and drop an invalid packet.
 
-A bigger issue we have discovered is that middleboxes do not always respect the standard and simply drop SYN-ACK packets with payload.
+A bigger issue we have discovered is that middleboxes do not always respect the standard and drop SYN-ACK packets with payload.
 We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}.
@@
 Finally we tested interaction reserved header bits.
-We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: all reserved bits were either allowed through or cleared at both phases.
-Importantly, we did not observe cases when such packets are dropped.
+We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences.
+All networks appear to pass reserved bits on non-proxied connections.
 
@@
 
-We repeat our tests on a few ports:
+To differentiate between general behavior of particular network and application-specific optimizations (\eg HTTP acceleration) we repeat our tests on a few application-specific ports as well as a random one:
 \begin{itemize}
@@
 In the first case, traffic travels over SSL ports (443, 993) unmodified except for packets with invalid checksum being dropped.
-Our traffic was not encrypted and an increasing proportion of traffic is served over SSL.
-A commercially deployed HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP\footnote{\url{http://db.awazza.com/users/global/}}, making the value of such proxying questionable.
+Our traffic was not encrypted.
+An increasing proportion of traffic is served over SSL, making the value of such proxies questionable: a commercially deployed HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP\footnote{\url{http://db.awazza.com/users/global/}}.
 
-We confirmed the middlebox to be a ByteMobile proxy.
-It replaces the original connections with a new one which adds options not present in original packets, changes window size and resets other fields to new values to optimize TCP for cellular networks~\cite{Ha:2006td}.
-Such behavior makes deployment of any protocol modifications or extensions very difficult (e.g.
-we verified that TCPCrypt does not work on this network).
-Although we did not see more cases of it, the company claims over 160 clients.
+We confirmed the middlebox on Cellular 2 to be a ByteMobile proxy.
+It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to new values to optimize TCP for cellular networks~\cite{Ha:2006td}.
+Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that \eg TCPCrypt does not work on this network.
+We do not currently have an estimate of how widespread ByteMobile proxies are as we only observed them on one network.
 
-In the second case, traffic was only modified when going to ports 80 and 443. As we discuss below, the network optimizes HTTP traffic and since our traffic does not resemble valid HTTP, it is dropped.
+In the second case, traffic was only modified when going to ports 80 and 443.
+The network likely optimizes HTTP traffic and drops any non-HTTP traffic.
+In general port-specific behavior appears related to proxies.
 
@@
 
-Since proxies make new changes very difficult, we checked for non-transparent oes on in our tested networks on a number of popular port numbers:
+Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize and hiding any information within the packet becomes difficult.
+We checked for them in our tested networks on a number of popular port numbers:
 21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.
@@
 \vspace{-2mm}
-\caption{Non-transparent proxies and proxied port numbers by radio technology}
+\caption{Non-transparent proxies and their proxied port numbers by radio technology}
 \label{tab:proxies}
@@
 
-Results are summarized in Table~\ref{tab:proxies}. Cellular 2 proxies almost all of our tested ports except for 443 and 993 as observed before. 
+We summarize results in Table~\ref{tab:proxies}.
+3G and 4G networks use different IP core networks, therefore we split the results by radio technology.
+We did not detect proxies on Cellular 1 and Cellular 3 networks and did not have data about Cellular 6, but results of our TCP testsuite did not indicate any proxies.
+Cellular 2 proxies almost all of our tested ports except for 443 and 993.
+
 Cellular 4 shows non-uniform behavior across the network.
-It is not obvious why ports commonly used for encrypted communication are proxied over LTE, we only recorded 1 case for HTTPS.
-The other two cases (* and **) seem related to the operator being formed from two merged operators.
-These differences are evident in \emph{whois} records and device-reported network name --- \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO), which rent network infrastructure from the mobile network operator.
+It is not obvious why the network proxies connections to ports commonly used for encrypted communications.
+Currently we have only recorded 1 such case for HTTPS.
+The differences between the other two cases (* and **) seem related to forming the operator from two merged ones.
+Their infrastructure may have not been completely consolidated: \emph{whois} records and device-reported network names are different.
+Case \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO).
+MVNOs buy network services from operators at wholesale rates and typically have little infrastructure of their own, and expect to see the same behavior as that of the underlying network.
 
-It is not safe to estimate proxying just from \emph{whois} records: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios.
-An active test is needed reliably detect proxies.
+Estimating proxying from \emph{whois} records alone is not safe: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios.
+Reliable detection needs active testing, potentially combined with caching for repeated connections.
 
-Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions.
-While other networks did not interfere with our tests, they do not appear to be completely free of proxies.
-On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80.
+%Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions.
+While some networks did not interfere with our tests, they do not appear to be completely free of proxies.
+On Cellular 1 (15 sessions) two sessions show DNS proxies (over UDP) and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80.
 Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic.
-We did not have data about Cellular 6, but results discussed above suggest that there are no non-transparent proxies present.
-
-All other networks did not show non-transparent proxy behavior, but the ones who did pose a challenge that is not easily solved.
-Proxies potentially change every aspect of the original connection and hiding any information within the packet becomes extremely difficult since any value that a proxy does not recognize is reset.
 
@@
 In addition, it is a 32-bit value and we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant.
-The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Fig.~\ref{fig:header}), carrying the desired opcode.
+The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}), carrying the desired opcode.
 For simultaneous open we can use the same mechanism in both directions.
@@
 
-% \vspace{-2mm}
+\vspace{-2mm}
 \begin{align*}
@@
 We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream.
-We extend the three-way handshake with fields shown in Fig.~\ref{fig:header}:
+We extend the three-way handshake with fields shown in Figure~\ref{fig:header}:
 
@@
 
-An example handshake is shown in Fig.~\ref{fig:handshake}.
+Figure~\ref{fig:handshake} shows an example handshake.
 In this case a client (active opener) sends a SYN packet as show above.
-The packet traverses a middlebox which does not remove the opcode and reaches the Server (passive opener).
-The Server replies with SYN-ACK with U\_Opcode which is discarded by the middlebox.
-It then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, which reaches the Client.
+The packet traverses a middlebox that does not remove the opcode and reaches the Server (passive opener).
+The Server replies with SYN-ACK with U\_Opcode, but the middlebox discards it.
+The server then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, delivered to the Client.
 The client recovers and verifies the received opcode.
-Both endpoints are now assumed to understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging reception of the $Opcode$ from the Server.
+Hence both endpoints understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging reception of the $Opcode$ from the Server.
+
+We witnessed a case of packet splitting where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload.
+We did not observe a case where a middlebox both passes payload in SYN-ACK packet and splits ACK segments.
+If the handshake succeeds we need to take a little extra care with the fallback mechanism -- we need to wait for an ACK segment with payload before falling back to vanilla TCP.
+If the passive opener receives data it can not decode as valid TLV-encoded control information, it falls back to vanilla TCP.
 
@@
 The premise of our work is that middleboxes should not be aware of the new control streams in the first place and therefore not interfere with them.
+Currently we focus on exchanging control information within the packet header and leave refinement of the full control channel for future work.
 
-We witnessed a case of resegmentation where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload.
-It is not an issue with TLV-encoded control stream in the payload, because it will be delivered to the receiver first.
-The mechanism of \emph{Checksum Correction} would, however, be compromised for the first segment.
-If the passive opener receives data during or immediately after the handshake but can not decode that as valid TLV-encoded control information, it treats that as application data and falls back to vanilla TCP.
+\subsection{Control channel bitrate}
 
-The focus of our work is to show the possibility of exchanging control information within the packet header. We therefore leave refinement of the full control channel for future work.
+In progress...
 
@@
 The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is explicitly not designed as a mechanism for sending such data.
-Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and in general it is recommended against the use of the mechanism~\cite{Gont:2011vi}.
+Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and the general recommendation is against the use of the mechanism~\cite{Gont:2011vi}.
 
@@
 
-Generic control stream for MPTCP~\cite{Bonaventure:wx} can be reused as well.
+Generic control stream for MPTCP~\cite{Bonaventure:wx} is another alternative.
 It proposes the mapping control stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams.
@@
 
-%More generally SCTP uses \emph{data chunks} to differentiate between different subflows.
-Each packet can contain multiple TLV-encoded chunks to separate user data and control information.
-Despite its challenged deployment in the Internet, SCTP over DTLS is being used as the basis of WebRTC data protocol~\cite{Tuexen:wv} and we suggest that such chunking could be used on top of \emph{noTCP} once control channel has been bootstrapped.
-
 \section{Discussion}
 
-The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible?
+The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, deployable and forwards-compatible?
 We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.
@@
 
-In the long term and beyond our example control information exchange we believe that each protocol-relate standards (TCP extensions in this case) should explicitly define which header fields it interferes with and how.
+In the long term and beyond our example control information exchange we believe that each protocol-relate standards (TCP extensions in this case) should explicitly define all header fields it interferes with and how.
 Examples include extending sequence numbers and using stronger message integrity checks~\cite{Mazieres:uz}, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple checksum or window size modifications.
@@
 More importantly, we have left open the question of finding a method to hide such information from non-transparent proxies.
-The easiest, albeit very limited solution is to identify and cache port numbers that allow unmodified traffic.
-Our goal instead is to study commonly found proxy implementations in detail in order to quickly identify them and encode information within their own operating patterns.
+The easiest, albeit limited solution is to identify and cache port numbers that allow unmodified traffic.
+Instead we plan to study commonly found proxy implementations in detail in order to quickly identify them and encode information within their own operating patterns.
 
-The final goal is to map header field modifications by the various extensions as well as network middleboxes into an algebraic form as we have done in the simple case of checksum.
+Finally, we hope to map header field modifications by existing extensions and network middleboxes into an algebraic form as we have done in the simple case of checksum.
 This way we would not only be able to exchange control information across the network reliably, but also identify network path behavior by looking at its performed operations.
@@
 
-We hope to motivate the research community to explore the full range of possibilities in extending existing protocols.
-We have developed an Android application to collect data about how network middleboxes interfere with packet header fields that may not currently have a meaning.
-We encourage more people to install our application to build a comprehensive database of network behavior.
-The goal of the study is to explore new ways of exchanging control information resilient to major classes of currently deployed middleboxes by using steganographic approaches.
+Our hope is to motivate the research community into exploring the full range of possibilities in this space and to encourage more people to install our application so that we can build a comprehensive database of middlebox interference with protocols across different network providers.
 
-Finally, we would like to use the results of the study to develop a set of rules for how packet headers are modified by middleboxes as well as standard protocol extensions.
-If this were true, extending a protocol could become near automatic.
+The study will explore new ways of exchanging control information resilient to major classes of deployed middleboxes by using steganographic approaches.
+We would like to use the results of the study to develop a set of rules for how protocol extensions redefine and how middleboxes modify packet headers.
+Development of such ruleset would  mean that extending a protocol in the future could become near automatic.
 

### 2014-07-16 paper/paper.tex
channel bitrates part and comments from toby

@@
 TCP is the de facto transport protocol in the Internet.
-However it has many limitations that make it ill suited to the modern world of high mobility and multi-homing.
-Protocol optimizations and filtering performed by middleboxes make extension deployment challenging which has led researchers to explore new ways of exchanging control information between endpoints.
-In this paper we explore steganographic approaches to smuggle information between endpoints to avoid middlebox interference.
-We design NoTCP -- a mechanism which uses the TCP header fields in combinations that are currently have no defined semantics to hide a small amount of control information from middleboxes and hence bootstrap a control channel.
+However it has many limitations that make it ill-suited to the modern world of high mobility and multi-homing.
+Protocol optimizations and filtering performed by middleboxes make extension deployment challenging.
+This has led researchers to explore new ways of exchanging control information between endpoints.
+We present NoTCP -- a mechanism which uses the TCP header fields in combinations that have no defined semantics in order to hide a small amount of control information from middleboxes and hence bootstrap a control channel.
 We also present a preliminary study showing how the mechanism works across cellular and WiFi networks.
@@
 Over the last decade a number of new protocols have been proposed to replace TCP.
-However many middle boxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like.
-Consequently few of these protocols have found any leverage.
-
-Modern networks necessitate new functionality, such as Multipath Connectivity~\cite{Handley:vj} for multi-homed devices and ubiquitous traffic encryption~\cite{Mazieres:uz}).
-The applications of exchanging control information between endpoints range from new secure extensions~\cite{Rescorla:R080FoGB} to exchanging information about endpoints for connection parameter tuning (\eg initial congestion window~\cite{Dukkipati:2010hs}) to designing better coordination for long-lived connections to save mobile device energy~\cite{Aucinas:2013uk}.
+However many middleboxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like.
+Consequently, few of these protocols have found any leverage.
 
@@
 Traditionally this is what TCP option space was used for.
-However option space is limited, heavily contended~\cite{Handley:vj,Mazieres:uz} and troubled by network middleboxes~\cite{Honda,Guha2005tb,UntoldMiddlebox2011}.
+However option space is limited, heavily contended~\cite{Handley:vj,Mazieres:uz} and network middleboxes make it difficult to use~\cite{Honda,Guha2005tb,UntoldMiddlebox2011}.
+On the other hand, modern networks demand new protocols for mobile and multi-homed devices~\cite{Mazieres:uz}, ubiquitous traffic encryption~\cite{Handley:vj,Rescorla:R080FoGB}, energy efficiency~\cite{Aucinas:2013uk}, connection parameters tuning for specific workloads and environments~\cite{Dukkipati:2010hs,Erman:2013uz} that were unknown 30 years ago.
 
-We believe that the solution is to work round the middle boxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header.
+We believe that the solution is to work round the middleboxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header.
 This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that SYN Cookies~\cite{Eddy:2007to} do currently.
@@
 \small{
-\includegraphics[width=\columnwidth]{figs/tcp-header}
-\vspace{-3mm}
+\includegraphics[width=.99\columnwidth]{figs/tcp-header}
+\vspace{-5mm}
 \caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload}
@@
 }
-\vspace{-2mm}
+\vspace{-4mm}
 \end{figure}
 
-Figure~\ref{fig:header} shows TCP header with fields that we propose for reuse and define later in the paper.
-However, the structure of TCP header is very simple and gives are few choices for hiding extra control information:
+Figure~\ref{fig:header} shows the TCP header with fields that we propose for reuse and define later in the paper.
+However, the structure of TCP header is very simple and gives few choices for hiding extra control information:
 \begin{itemize}
-    \item Header fields without corresponding flags to have no meaning to legacy implementations.
+    \item Setting header fields without the corresponding flags so they have no meaning to legacy implementations.
     \item Assigning one or more of the reserved field bits.
-    \item Defining new semantics for valid header fields.
+    \item Overloading the semantics for valid header fields.
 \end{itemize}
@@
 
-The only header fields that may not have a meaning depending on the corresponding flag are \emph{Urgent Pointer} and \emph{Acknowledgment Number}.
+Only 2 header fields have their semantics set by corresponding flags.
+These are the \emph{Acknowledgment Number} and \emph{Urgent Pointer} fields.
+The former must also always be set after the initial SYN packet.
+%The only header fields that may not have a meaning depending on the corresponding flag are \emph{Urgent Pointer} and \emph{Acknowledgment Number}.
 The latter must always be set after the initial SYN packet.
 
-The general recommendation is to not use It is recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
+The general recommendation is to not use the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
 On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
-Better yet, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and can be redefined.
+Importantly, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and so can be redefined.
 
-Acknowledgment field is only available in the initial SYN packet.
+The Acknowledgment Number field is only available in the initial SYN packet.
 It must therefore be combined with another mechanism to acknowledge control information unless TCP Simultaneous Open is used.
@@
 \subsubsection*{Reserved bits}
+\label{sec:reserved}
 
-Only 3 TCP Reserved bits have not been allocated since the three ECN bits (ECN, CWR and experimental NS)~\cite{Floyd:up,Ely:uc} that are still not universally implemented~\cite{Kuhlewind:2013hr}.
-
-Due to simplicity and non-interference with legacy implementations there are proposals that make use of them.
+3 TCP Reserved bits remain unallocated.
+One further bit is only allocated to the experimental ECN nonce\cite{Ely:uc}.
+Despite their limited size, there are proposals that make use of new reserved bits due to implementation simplicity and non-interference with legacy implementations.
 One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets.
-The solution avoids triggering fast retransmit by such duplicate acknowledgments by marking them as duplicates in a reserved flag.
+The proposal deliberately sends copies, but uses a reserved flag to explicitly mark solution avoids triggering fast retransmit by such duplicate acknowledgments by marking them as duplicates in a reserved flag.
 
@@
 
-The remaining choice is to encode information within valid header fields.
+The remaining choice is to encode information within otherwise valid header fields.
 It involves generating packets with specific values for initial sequence numbers, window size or checksums.
 Arguably it is an acceptable choice if deployable in today's networks.
+Indeed SYN Cookies already do this~\cite{Eddy:2007to}
 
-Initial sequence numbers must be hard to predict as they provide a measure of security and until strong encryption.
-Stronger security measures such as TCPCrypt are better suited for this purpose, but are not universally adopted.
-Adding information to the field to be easily decoded by the receiver makes it predictable and weakens already problematic security~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.
+Initial sequence numbers must be hard to predict as they provide a measure of security against TCP injection attacks.
+Adding information to the field to be easily decoded by the receiver makes it predictable and weakens already problematic security~\cite{Bellovin:uz,Qian:2012wb}.
+Some stateful middleboxes are known to interfere with sequence numbers to improve security instead making it worse~\cite{Qian:2012bj}.
+Stronger security measures such as TCPCrypt~\cite{Mazieres:uz} are better suited for connection protection.
 
@@
 Firstly, it is only used for error detection.
-It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown unreliable~\cite{Stone:2000fc}.
-Instead, it may be replaced with extensions' separate, potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code) or left for the upper layer.
+It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown to be unreliable~\cite{Stone:2000fc}.
+Instead, it may be replaced with potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code) or left for the upper layer.
 
@@
 
-\section{Methodology}
+\section{Measurement Methodology}
 
-To verify current network behavior we have developed an Android application distributed through the \emph{Google Play Store} to collect data from volunteers.
-It contains a simplified implementation of TCP to generate packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown.
+To measure current network behavior we have developed an Android application distributed through the \emph{Google Play Store}\footnote{Link omitted to anonymize submission} to collect data from volunteers.
+It contains a simplified implementation of TCP which generates packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown.
 It allows us to modify various header fields to our purposes and check the values received by the other endpoint.
@@
 
-The testing setup is:
+The test setup is:
 \begin{itemize}
@@
 We implemented previously developed techniques~\cite{Weaver:RHwbx82O} and run the test separately.
-The primary one was non-responsive server test: the server is instrumented not to respond to SYN, therefore if connection succeeds it is to a proxy on the path.
+The primary one was non-responsive server test: the server is instrumented not to respond to SYN, therefore if the connection succeeds it is to a proxy on the path.
 We repeated these tests on different destination ports and found that proxying can be specific to certain port numbers.
@@
 Potential limitations of our methodology include that we do not consider packet splitting and coalescing.
-Our goal is to exchange control information within packet headers to avoid dealing with it altogether.
-To ensure that splitting does not happen during our tests we do not send more than a few bytes of data and the exchange only consists of a single message sent each way.
-Due to the same reason we do not take congestion control into account either.
+By exchanging information within packet headers and using small packets we avoid packet splitting.
+We also prevent splitting during data exchanging by not send more than a few bytes of data.
+Coalescing is prevented by only exchanging a single data message each way.
+For the same reason we do not take congestion control into account.
 
-Raw sockets operate in parallel to the kernel network stack.
+\emph{Raw sockets} operate in parallel to the kernel network stack.
 We had to suppress reset packets generated by the kernel in response to TCP packets from sessions it is not aware of.
 We also had to choose local port numbers unused by the kernel.
-We chose them randomly from the unprivileged port range and relied on other error detection mechanisms preventing invalid packet delivery to the applications.
+We chose them randomly from the unprivileged port range and relied on other error detection mechanisms preventing invalid packet delivery to any applications.
 
-\section{Network characteristics}
+\section{Preliminary Results}
 \label{sec:network}
@@
 \begin{center}
-\begin{tabular}{| l | >{\centering\arraybackslash}m{1.2cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
+\begin{tabular}{| l | >{\centering\arraybackslash}m{1.3cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{2.4cm} | >{\centering\arraybackslash}m{2.4cm} | >{\centering\arraybackslash}m{2.3cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
 \hline
-    \textbf{Net ID} & \textbf{Country} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop SYN-ACK data} & \textbf{Normalize SYN Ack number}  & \textbf{Normalize Urgent Pointer} & \textbf{Normalize Reserved} & \textbf{Remap Sequence} \\ \hline \hline
+    \textbf{Net ID} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop SYN-ACK data} & \textbf{Normalize SYN Ack number}  & \textbf{Normalize Urgent Pointer} & \textbf{Normalize Reserved} & \textbf{Remap Sequence} \\ \hline \hline
     % Univ Helsinki WiFi
-    WiFi edu 1      & Finland               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    WiFi edu 1      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
     % Eduroam
-    WiFi edu 2      & UK               &                        &                            & \checkmark             &                         & \checkmark             &                             & \checkmark          \\ \hline
+    WiFi edu 2      &                        &                            & \checkmark             &                         & \checkmark             &                             & \checkmark          \\ \hline
     % wgb
-    WiFi pub 1      & UK               &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    WiFi pub 1      &                        &                            &                        &                         &                        &                             &                     \\ \hline
     % MKSW germany (carlos)
-    WiFi pub 2      & Germany               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    WiFi pub 2      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     % Virgin home
-    WiFi res 1      & UK               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    WiFi res 1      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
     % TEO Lt
-    WiFi res 2      & Lithuania               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    WiFi res 2      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     % Telefonica Free
-    WiFi ent 1      & Spain               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
+    WiFi ent 1      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
     \hline
     % Finland internet.saunalahti
-    Cellular 1      & Finland               &                        &                            &                        &                         &                        &                             &                     \\ \hline
+    Cellular 1      &                        &                            &                        &                         &                        &                             &                     \\ \hline
     % GiffGaff
-    Cellular 2      & UK               &  443, 993              & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    Cellular 2      &  443, 993              & *                          & *                      & *                       & *                      & *                           & *                   \\ \hline
     % E-Plus germany
-    Cellular 3      & Germany               &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
+    Cellular 3      &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
     % EE/T-Mobile
-    Cellular 4      & UK               &               80, 443  & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
+    Cellular 4      &               80, 443  & *                          & *                      & *                       & *                      & *                           & *                   \\ \hline
     % Three UK
-    Cellular 5      & UK               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    Cellular 5      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
     % Omnitel LT
-    Cellular 6      & Lithuania               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
+    Cellular 6      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
 \end{tabular}
@@
 }
-\vspace{-3mm}
-\caption{Network behavior observed through tests generating custom TCP packets. A dash means different cases observed based on port numbers.}
+\vspace{-4mm}
+\caption{Network behavior observed through tests generating custom TCP packets. * means different cases observed based on port numbers.}
 \label{tab:networks}
-\vspace{-2mm}
+\vspace{-4mm}
 \end{table*}
 
-We show our collected results in Table~\ref{tab:networks}.
-Since we were focusing on mobile applications, in our preliminary study we collected results across public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries.
+For our preliminary study we have focused on mobile networks including public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries.
+We present our results in Table~\ref{tab:networks}.
 
-The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and ones that discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers.
-Only one network provided clients with a global IP (WiFi edu 2).
+The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and others that interfered with our traffic.
+The interference we tested for includes resetting fields without corresponding flags set (\emph{Acknowledgment Number} and \emph{Urgent Pointer}) or dropping them, discarding packets with invalid checksums, and resetting reserved flags.
+Only one network provided clients with a global IP (WiFi edu 2) and we also tested whether NATs rewrite sequence numbers in addition to addresses and port numbers.
+
+As discussed in Section~\ref{sec:unset}, we can squeeze in a few extra bits of information by using certain header fields when the corresponding flags are not set.
+None of our tested networks filtered the Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with the Urgent Pointer set without the URG flag.
+
+We also tested middlebox interaction with reserved header bits (Section~\ref{sec:reserved}).
+We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences.
+All networks appear to pass reserved bits on non-proxied connections.
 
@@
 
-A bigger issue we have discovered is that middleboxes do not always respect the standard and drop SYN-ACK packets with payload.
+A bigger issue we have discovered is that middleboxes do not always respect the standard~\cite{Postel:3EDyoxP_} and drop SYN-ACK packets with payload.
 We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}.
-
-% \begin{quotation}
-%     Although these examples do not show connection synchronization using data-carrying segments, this is perfectly legitimate
-% \end{quotation}
-
-As discussed in Section~\ref{sec:unset}, we can squeeze in a few extra bits of information by defining meaning to header fields when corresponding flags are not set.
-None of our tested networks filtered Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with Urgent Pointer set without URG flag.
-
-Finally we tested interaction reserved header bits.
-We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences.
-All networks appear to pass reserved bits on non-proxied connections.
 
@@
 We confirmed the middlebox on Cellular 2 to be a ByteMobile proxy.
-It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to new values to optimize TCP for cellular networks~\cite{Ha:2006td}.
-Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that \eg TCPCrypt does not work on this network.
+It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to different values to optimize TCP for cellular networks~\cite{Ha:2006td}.
+Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that TCPCrypt fails on this network.
 We do not currently have an estimate of how widespread ByteMobile proxies are as we only observed them on one network.
 
-In the second case, traffic was only modified when going to ports 80 and 443.
+In the case of \emph{Cellular 4}, traffic was only modified when going to ports 80 and 443.
 The network likely optimizes HTTP traffic and drops any non-HTTP traffic.
-In general port-specific behavior appears related to proxies.
+Generally port-specific behavior appears related to proxies.
 
@@
 }
-\vspace{-2mm}
-\caption{Non-transparent proxies and their proxied port numbers by radio technology}
+\vspace{-4mm}
+\caption{Non-transparent proxies and their proxied port numbers by radio technology. */** - see text for details.}
 \label{tab:proxies}
@@
 
-What is common to all candidate protocol changes described above is that they are very limited in size and therefore sufficient to exchange an opcode or an ID.
+What is common to all candidate protocol changes described above is that they are very limited in size and therefore only sufficient to exchange an opcode.
 Other protocol details must instead be exchanged either out of band, implicit to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information.
@@
 A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections.
-In addition, it is a 32-bit value and we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant.
+As it is a 32-bit value we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant.
 The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}), carrying the desired opcode.
@@
 
-For acknowledging extensions in SYN-ACK packet we use Urgent Pointer and Checksum fields.
+For acknowledging extensions in SYN-ACK packet we use the Urgent Pointer and Checksum fields.
 Our tests indicate that neither choice is guaranteed to succeed, therefore we combine the two by redefining their values when SYN=1.
-We define Urgent Pointer to be $Opcode_U$ when URG=0 and Checksum to carry a meaning when Urgent Pointer = 0 to get around the cases where one is blocked but the other one is not.
-Across all networks we have tested in our preliminary study at least one of the choices succeeds.
+We define the Urgent Pointer to be $Opcode_U$ when URG=0 and Checksum to carry a meaning when Urgent Pointer = 0 to get around the cases where one is blocked but the other one is not.
+Across all networks we tested in our preliminary study, at least one of the choices succeeds.
 
@@
 The traditional NAT does a simple recalculation of the checksum~\cite{Egevang:tu}: subtracts the original source address and port number (in some cases also the sequence number) and adds the new values.
-The received Checksum after \(n\) NAT hops is therefore
+The received Checksum after \(n\) NAT hops is therefore:
 
@@
 Here $header$ refers to the fields of the header that change, namely the destination port number, address and sequence number.
-Previous study~\cite{UntoldMiddlebox2011} found no NATs mangling sequence numbers, but we witnessed a few cases of such behavior.
+A previous study~\cite{UntoldMiddlebox2011} found no NATs mangling sequence numbers, but we witnessed a few cases of such behavior.
 The Opcode can then be encoded as a target checksum so that it is recoverable by a simple binary subtraction of the changing fields (3 to 7 16-bit arithmetic operations in total):
@@
 There are two choices for setting checksum to a specific value: invalid checksum or adding padding payload.
-The latter can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, but keeps a packet valid from network's perspective.
+The latter can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, but this approach keeps a packet valid from network's perspective.
 
-For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum, in line with other extensions which reuse part of data payload for option negotiation and maintaining validity of a packet.
+For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum to maintain validity of a packet.
+The choice is in line with other extensions which reuse part of data payload~\cite{Mazieres:uz,Bonaventure:wx}.
 This Checksum Correction field is not delivered to the application.
@@
 
-Sending data with SYN-ACK packet works on all tested cellular networks, but does not on most WiFi.
+Sending data with the SYN-ACK packet works on all tested cellular networks, but not on most WiFi.
 For the above header modifications the information either reaches the destination as intended, or the packet is dropped.
@@
 
-\subsection{Bootstrapping control stream}
+\subsection{Bootstrapping the control stream}
 
@@
 \item If passive opener does not receive an ACK, retransmit with $Opcode_C$ and \emph{Checksum Correction}.
-If that fails as well, retransmit SYN-ACK with an invalid checksum.
+If that fails as well, retransmit the SYN-ACK with an invalid checksum.
 Finally, fall back to vanilla TCP and retransmit a standard SYN-ACK.
@@
 
-Putting control stream into packet payload suffers from much the same issues as other proposals to extend TCP option space~\cite{Ramaiah:2012wa}.
+Putting control stream information into packet payload suffers from much the same issues as other proposals to extend TCP option space~\cite{Ramaiah:2012wa}.
 The premise of our work is that middleboxes should not be aware of the new control streams in the first place and therefore not interfere with them.
@@
 
-In progress...
+%Covert channels are typically low bitrate in order to make it difficult to detect them.
+Application-specific ones, especially based on media transcoding can achieve rates from 270 b/s using $\mu TP$ torrent protocol, to 2 kbit/s by encoding data in silent Skype voice call periods, to megabits in wireless environments by inserting hidden data into the padding of frames~\cite{Zielinska:2014fn}.
+
+Our control channel does not need to protection from steganalysis as it is only used to avoid middlebox interference with end-to-end control information exchange.
+Bitrate of the channel only using packet headers (\ie without redefining part of packet payload for control information) is therefore bounded by the size of the TCP header (20 bytes without options), the round trip time (RTT) and the number of packets in flight per RTT.
+Depending on fields usable for a particular connection as inferred during the handshake, we can use between 16 (either $Opcode_C$ or $Opcode_U$) and 35 bits when using both fields and reserved flags.
+
+For any given operating window $cwnd$ (in bytes) a TCP flow has $cwnd/MSS$ data packets per RTT, hence that many headers with covert control data.
+If \emph{window scaling} option is not used, the maximum \emph{receive window} is 65536 bytes.
+Using the typical HSPA network RTT of about 100ms the maximum data rate is therefore 5 Mbit/s, reasonable for current networks~\cite{Chen:2012vj}.
+In this case covert channel bitrate is up to 7.0 to 15.4 kbit/s, depending on the fields used.
+
+Otherwise, we can estimate channel bitrate from previously studied 3G and 4G network characteristics~\cite{Chen:2012vj}.
+The average bandwidth measured for a large TCP transmission ranges from $1.2\pm 0.2$ Mbit/s for a 3G network and $348\pm 141$ ms RTT to $15.6\pm 1.0$ Mbit/s and $125\pm 41$ ms RTT for LTE.
+Covert channel bitrates for these cases range from $3.7 \pm 0.6$ kbit/s for 3G with one 16-bit field usable to $49.0 \pm 3.1$ kbit/s for LTE and both fields as well as reserved bits usable.
+
+Finally, the actual amount of information exchanged will depend on the coding scheme and the required reliability guarantees.
+For reliability in particular, a scheme described in Nusha~\cite{Rutkowska:2004ul} can be used.
+
+For comparison, bitrates achievable by detection resistant schemes using initial sequence numbers are limited by only using the first packet in each connection to fit up to 15 bits of data~\cite{Murdoch:2005fz}.
+Using the same assumptions as before and establishing connections sequentially results in 150 bit/s when only SYN and SYN-ACK are exchanged, to 75 bit/s with complete handshake and teardown.
 
@@
 The main focus has been on covert channels that violate system's security policy and initial sequence numbers~\cite{Rowland:1997vq}, TCP Timestamps~\cite{Giffin:2002wh} and combinations with IP flags~\cite{Murdoch:2005fz}.
-There are detection techniques for these methods and our main goal is to exchange a small amount of control information rather than a genuine covert channel, asking for a different design.
+There are detection techniques for these methods, however our main goal is to exchange a small amount of control information rather than a genuine covert channel, calling for a different design.
 
-TCP implements an \emph{urgent mechanism} at its core that allows the sending user to stimulate the receiving user to accept some \emph{urgent data}.
+TCP implements an \emph{urgent mechanism} at its core that allows the sending user to make the receiving user to accept some data as \emph{urgent}.
 The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is explicitly not designed as a mechanism for sending such data.
@@
 AccECN proposal~\cite{Kuhlewind:2014vd} uses an additional reserved bit, overloads the meaning of already assigned ECN~\cite{Floyd:up} and NS~\cite{Ely:uc} bits and redefines Urgent Pointer as Non-Urgent when URG is not set.
-The idea is highly related to ours, but reserving the field solely for ECN use prevents other extensions from using.
-Furthermore, we saw that some networks block packets with such field, making it especially difficult to use for ECN.
+The idea is highly related to ours, but reserving the field solely for ECN use prevents other extensions from using it.
+Furthermore, we saw that some networks block such packets, making it especially difficult to use for ECN.
 
 Generic control stream for MPTCP~\cite{Bonaventure:wx} is another alternative.
-It proposes the mapping control stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams.
+It proposes mapping the control stream into a separate sequence number space and exchanging control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams.
 The underlying assumption, however, is that MPTCP is already deployed and that existing deployments will be compatible with the new specification.
@@
 
-The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, deployable and forwards-compatible?
+The main points of discussion for \emph{NoTCP} are the same as with all TCP extensions: is it necessary, deployable and forwards-compatible?
 We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.
@@
 
-In the long term and beyond our example control information exchange we believe that each protocol-relate standards (TCP extensions in this case) should explicitly define all header fields it interferes with and how.
-Examples include extending sequence numbers and using stronger message integrity checks~\cite{Mazieres:uz}, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple checksum or window size modifications.
-
 More importantly, we have left open the question of finding a method to hide such information from non-transparent proxies.
@@
 
-Finally, we hope to map header field modifications by existing extensions and network middleboxes into an algebraic form as we have done in the simple case of checksum.
+In the long term we believe that each protocol-related standard should explicitly define how it interferes with all header fields.
+Examples include implicitly extending sequence number space~\cite{Mazieres:uz}, allowing encrypted TCP session to omit the timestamp option, using stronger message integrity checks making checksum field redundant, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple window size modifications~\cite{Crowcroft:1998dk}.
+
+We hope to map such header field modifications by existing extensions as well as network middlebox behavior patterns into an algebraic form as we have done in the simple case of checksum.
 This way we would not only be able to exchange control information across the network reliably, but also identify network path behavior by looking at its performed operations.
-This way extending a protocol would become a matter of defining a new set of such header algebra rules and simplify the process of design and testing.
+Thus extending a protocol would become a matter of defining a new set of such header algebra rules hence simplifying the process of design and testing.
 
@@
 
-Our hope is to motivate the research community into exploring the full range of possibilities in this space and to encourage more people to install our application so that we can build a comprehensive database of middlebox interference with protocols across different network providers.
+This research is still in early stages.
+The end goal is to develop a set of rules for how protocol extensions redefine and how middleboxes modify packet headers and use them to hide information so that middleboxes do not interfere with endpoint behavior.
+Development of such ruleset would mean that extending a protocol in the future could become near automatic.
 
-The study will explore new ways of exchanging control information resilient to major classes of deployed middleboxes by using steganographic approaches.
-We would like to use the results of the study to develop a set of rules for how protocol extensions redefine and how middleboxes modify packet headers.
-Development of such ruleset would  mean that extending a protocol in the future could become near automatic.
+More importantly, this paper hopes to motivate the research community to explore the full range of possibilities in this space as well as contribute to our work by downloading our research app and helping to build a comprehensive database of middlebox interference with protocols across different networks.
 

### 2014-07-16 paper/paper.tex
another iteration of comments from toby, version given to Anil

@@
 \small{
-\includegraphics[width=.99\columnwidth]{figs/tcp-header}
-\vspace{-5mm}
+\includegraphics[width=\columnwidth]{figs/tcp-header}
+% \vspace{-5mm}
 \caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload}
@@
 }
-\vspace{-4mm}
+% \vspace{-4mm}
 \end{figure}
 
-Figure~\ref{fig:header} shows the TCP header with fields that we propose for reuse and define later in the paper.
-However, the structure of TCP header is very simple and gives few choices for hiding extra control information:
+Figure~\ref{fig:header} shows the TCP header with fields that we propose for reuse (defined later in the paper).
+The structure of the TCP header is very simple and gives few choices for hiding extra control information:
 \begin{itemize}
@@
 
-\subsubsection*{Unset Flags}
+\subsubsection*{Flag-dependent semantics}
 \label{sec:unset}
 
-Only 2 header fields have their semantics set by corresponding flags.
+Only two header fields have their semantics set by corresponding flags.
 These are the \emph{Acknowledgment Number} and \emph{Urgent Pointer} fields.
 The former must also always be set after the initial SYN packet.
-%The only header fields that may not have a meaning depending on the corresponding flag are \emph{Urgent Pointer} and \emph{Acknowledgment Number}.
-The latter must always be set after the initial SYN packet.
 
 The general recommendation is to not use the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
-On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
+On the other hand, Network Intrusion Detection Systems (NIDS) often reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
 Importantly, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and so can be redefined.
@@
 
-3 TCP Reserved bits remain unallocated.
+Three TCP Reserved bits remain unallocated.
 One further bit is only allocated to the experimental ECN nonce\cite{Ely:uc}.
@@
 The remaining choice is to encode information within otherwise valid header fields.
-It involves generating packets with specific values for initial sequence numbers, window size or checksums.
+It involves generating packets with specific values for \emph{Initial Sequence Numbers}, \emph{Window Size} or \emph{Checksums}.
 Arguably it is an acceptable choice if deployable in today's networks.
@@
 
-Initial sequence numbers must be hard to predict as they provide a measure of security against TCP injection attacks.
-Adding information to the field to be easily decoded by the receiver makes it predictable and weakens already problematic security~\cite{Bellovin:uz,Qian:2012wb}.
-Some stateful middleboxes are known to interfere with sequence numbers to improve security instead making it worse~\cite{Qian:2012bj}.
+\emph{Initial Sequence Numbers} must be hard to predict as they provide a measure of security against TCP injection attacks.
+Adding information to the field to be easily decoded by the receiver makes it more predictable and weakens already problematic security~\cite{Bellovin:uz,Qian:2012wb}.
+Also some stateful middleboxes are known to interfere with sequence numbers in an attempt to improve security~\cite{Qian:2012bj}.
 Stronger security measures such as TCPCrypt~\cite{Mazieres:uz} are better suited for connection protection.
 
-Window size is potentially a better candidate, especially if used during connection setup phase since data is not typically transmitted during the handshake.
-Using lower-order bits for control information would prevent flow control from operating properly, but higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of advertised window size.
-Using the field risks interfering both with TCP flow control mechanism as well as window size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed \eg by Citrix ByteMobile Proxy~\cite{Ha:2006td}.
+\emph{Window Size} is potentially a better candidate, especially if used during connection setup phase since data is not typically transmitted during the handshake.
+Using lower-order bits for control information would prevent flow control from operating properly, but higher-order bits could instead be given up when the \emph{window scaling} option is used to compensate for the decrease of advertised window size.
+Using the field risks interfering both with TCP flow control mechanism as well as window size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed by \eg Citrix ByteMobile Proxy~\cite{Ha:2006td}.
 
@@
 It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown to be unreliable~\cite{Stone:2000fc}.
-Instead, it may be replaced with potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code) or left for the upper layer.
-
+Instead, it may be replaced with potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code or left for the upper layer.
 Secondly, checksum operations are efficient as they only involve one's complement addition.
-We show later in the paper (Section~\ref{sec:protocol}) how checksums can be generated in a way that the intended value is recovered by the receiver after traversing multiple layers of Network Address Translation (NAT).
+We show later in the paper (\S\ref{sec:protocol}) how checksums can be generated in a way that the intended value is recovered by the receiver after traversing multiple layers of Network Address Translation (NAT).
 
@@
 It contains a simplified implementation of TCP which generates packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown.
-It allows us to modify various header fields to our purposes and check the values received by the other endpoint.
+It allows us to modify various header fields and check the values received by the other endpoint.
 We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.
@@
 
-%Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic and whether non-HTTP traffic on HTTP port is filtered.
-
-\subsection{Limitations}
-
-Potential limitations of our methodology include that we do not consider packet splitting and coalescing.
-By exchanging information within packet headers and using small packets we avoid packet splitting.
-We also prevent splitting during data exchanging by not send more than a few bytes of data.
+Our methodology avoids packet splitting by only using small packets.
 Coalescing is prevented by only exchanging a single data message each way.
@@
 }
-\vspace{-4mm}
+% \vspace{-4mm}
 \caption{Network behavior observed through tests generating custom TCP packets. * means different cases observed based on port numbers.}
 \label{tab:networks}
-\vspace{-4mm}
+% \vspace{-6mm}
 \end{table*}
 
-For our preliminary study we have focused on mobile networks including public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries.
+Our preliminary study focused on mobile networks including public, residential, university and enterprise WiFi networks as well as six cellular networks in four different countries.
 We present our results in Table~\ref{tab:networks}.
@@
 The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and others that interfered with our traffic.
-The interference we tested for includes resetting fields without corresponding flags set (\emph{Acknowledgment Number} and \emph{Urgent Pointer}) or dropping them, discarding packets with invalid checksums, and resetting reserved flags.
-Only one network provided clients with a global IP (WiFi edu 2) and we also tested whether NATs rewrite sequence numbers in addition to addresses and port numbers.
+We looked for dropped packets and reset or altered fields: (\emph{Acknowledgment Number} and \emph{Urgent Pointer}, potentially invalid checksums and reserved flags.
+Only one network provided clients with a global IP (WiFi edu 2) so we also tested whether NATs rewrite sequence numbers in addition to addresses and port numbers.
 
-As discussed in Section~\ref{sec:unset}, we can squeeze in a few extra bits of information by using certain header fields when the corresponding flags are not set.
-None of our tested networks filtered the Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with the Urgent Pointer set without the URG flag.
+As discussed in \S\ref{sec:unset}, we can squeeze in a few extra bits of information by using certain header fields when the corresponding flags are not set.
+None of our tested networks filtered the Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with the Urgent Pointer set without the corresponding URG flag.
 
-We also tested middlebox interaction with reserved header bits (Section~\ref{sec:reserved}).
+We also tested middlebox interaction with reserved header bits (\S\ref{sec:reserved}).
 We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences.
@@
 
-We confirmed the middlebox on Cellular 2 to be a ByteMobile proxy.
+We confirmed the presence of a ByteMobile proxy on Cellular 2.
 It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to different values to optimize TCP for cellular networks~\cite{Ha:2006td}.
@@
 In the case of \emph{Cellular 4}, traffic was only modified when going to ports 80 and 443.
-The network likely optimizes HTTP traffic and drops any non-HTTP traffic.
+The network likely optimizes HTTP traffic and drops non-HTTP traffic.
 Generally port-specific behavior appears related to proxies.
@@
 
-Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize and hiding any information within the packet becomes difficult.
-We checked for them in our tested networks on a number of popular port numbers:
+Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize so hiding any information within the packet becomes difficult.
+We tested for them on a number of popular port numbers:
 21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.
@@
 }
-\vspace{-4mm}
+% \vspace{-4mm}
 \caption{Non-transparent proxies and their proxied port numbers by radio technology. */** - see text for details.}
 \label{tab:proxies}
-\vspace{-4mm}
+% \vspace{-4mm}
 \end{table}
-
-We summarize results in Table~\ref{tab:proxies}.
-3G and 4G networks use different IP core networks, therefore we split the results by radio technology.
-We did not detect proxies on Cellular 1 and Cellular 3 networks and did not have data about Cellular 6, but results of our TCP testsuite did not indicate any proxies.
-Cellular 2 proxies almost all of our tested ports except for 443 and 993.
-
-Cellular 4 shows non-uniform behavior across the network.
-It is not obvious why the network proxies connections to ports commonly used for encrypted communications.
-Currently we have only recorded 1 such case for HTTPS.
-The differences between the other two cases (* and **) seem related to forming the operator from two merged ones.
-Their infrastructure may have not been completely consolidated: \emph{whois} records and device-reported network names are different.
-Case \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO).
-MVNOs buy network services from operators at wholesale rates and typically have little infrastructure of their own, and expect to see the same behavior as that of the underlying network.
-
-Estimating proxying from \emph{whois} records alone is not safe: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios.
-Reliable detection needs active testing, potentially combined with caching for repeated connections.
-
-%Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions.
-While some networks did not interfere with our tests, they do not appear to be completely free of proxies.
-On Cellular 1 (15 sessions) two sessions show DNS proxies (over UDP) and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80.
-Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic.
-
-%    an_op_name    | an_cell_net_type | an_country | port |    ip_global (simplified to network address. Included the owner according to whois
-% db)
-% -----------------+------------------+------------+------+-----------------
-%  EE              | GPRS             | gb         |   80 | 31.72/19 (EE)
-%  EE              | GPRS             | gb         |  110 | 31.72/19 (EE)
-%  EE              | GPRS             | gb         |  143 | 31.72/19 (EE) 
-%  EE              | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE)
-%  EE              | HSPA             | gb         |   80 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
-%  EE              | HSPA             | gb         |  110 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
-%  EE              | HSPA             | gb         |  143 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
-%  EE              | LTE              | gb         |   80 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
-%  EE              | LTE              | gb         |  110 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
-%  EE              | LTE              | gb         |  143 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
-%  EE              | LTE              | gb         |  443 | 213.205.224/12 (EE)
-%  EE              | LTE              | gb         |  993 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
-%  EE              | LTE              | gb         |  995 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
-%  T-Mobile        | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile)
-%  T-Mobile        | HSPA             | gb         |   80 | 31.96/21 (EE), 149.254/16 (T-Mobile)
-%  T-Mobile        | HSPA             | gb         |  110 | 31.96/21 (EE), 149.254/16 (T-Mobile)
-%  T-Mobile        | HSPA             | gb         |  143 | 31.96/21 (EE), 149.254/16 (T-Mobile)
-%  Virgin          | HSPA             | gb         |   25 | 149.254/16 (T-Mobile)
-%  Virgin          | HSPA             | gb         |   80 | 149.254/16 (T-Mobile)
-%  Virgin          | HSPA             | gb         |  110 | 149.254/16 (T-Mobile)
-%  Virgin          | HSPA             | gb         |  143 | 149.254/16 (T-Mobile)
-%  giffgaff        | HSPA             | gb         |   21 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |   22 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |   25 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |   80 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  110 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  135 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  143 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  161 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  465 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  585 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  587 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         |  995 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         | 1194 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         | 1723 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         | 5060 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         | 6881 | 82.132/16 (O2)
-%  giffgaff        | HSPA             | gb         | 9001 | 82.132/16 (O2)
 
@@
 
-What is common to all candidate protocol changes described above is that they are very limited in size and therefore only sufficient to exchange an opcode.
-Other protocol details must instead be exchanged either out of band, implicit to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information.
+What is common to all candidate protocol changes described above is that they are limited in size and therefore only sufficient to exchange an opcode.
+Other protocol details must be exchanged either out of band, implicitly to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information.
 Here we propose one way of exchanging control information as an opcode to bootstrap a full control stream.
@@
 A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections.
-As it is a 32-bit value we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant.
-The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}), carrying the desired opcode.
-For simultaneous open we can use the same mechanism in both directions.
+As it is a 32-bit value we can use part of it to signal that the rest of the field is significant.
+The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}).
+For \emph{simultaneous open} we can use the same mechanism in both directions.
 
-For acknowledging extensions in SYN-ACK packet we use the Urgent Pointer and Checksum fields.
+For acknowledging extensions in the SYN-ACK packet we use the Urgent Pointer and Checksum fields.
 Our tests indicate that neither choice is guaranteed to succeed, therefore we combine the two by redefining their values when SYN=1.
-We define the Urgent Pointer to be $Opcode_U$ when URG=0 and Checksum to carry a meaning when Urgent Pointer = 0 to get around the cases where one is blocked but the other one is not.
-Across all networks we tested in our preliminary study, at least one of the choices succeeds.
+We define the Urgent Pointer to be $Opcode_U$ when URG=0.
+If Urgent Pointer is 0, Checksum is the opcode to get around the cases where one is blocked but the other one is not.
+Across all networks tested in our preliminary study, at least one of the choices succeeds.
+In the cases when both are usable data can be encoded across them as a single field once the handshake is completed.
 
@@
 The traditional NAT does a simple recalculation of the checksum~\cite{Egevang:tu}: subtracts the original source address and port number (in some cases also the sequence number) and adds the new values.
-The received Checksum after \(n\) NAT hops is therefore:
+The Checksum after \(n\) NAT hops is:
 
-\vspace{-2mm}
+% \vspace{-4mm}
 \begin{align*}
@@
 \end{align*}
+% \vspace{-2mm}
 
@@
 
-\vspace{-2mm}
+% \vspace{-4mm}
 \begin{align*}
@@
 For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum to maintain validity of a packet.
+The field value is not considered as part of application data.
 The choice is in line with other extensions which reuse part of data payload~\cite{Mazieres:uz,Bonaventure:wx}.
-This Checksum Correction field is not delivered to the application.
-TLV-Encoded (Type-Length-Value) Control Stream can be present in data part~\cite{Bonaventure:wx} of the SYN-ACK packet after the Checksum Correction field.
+Sending data with the SYN-ACK packet works on all tested cellular networks, but not on most WiFi.
 
-Sending data with the SYN-ACK packet works on all tested cellular networks, but not on most WiFi.
 For the above header modifications the information either reaches the destination as intended, or the packet is dropped.
@@
 We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream.
-We extend the three-way handshake with fields shown in Figure~\ref{fig:header}:
+After the handshake a higher-bandwidth \emph{Control Channel} can be defined by using TLV (Type-Length-Value) encoding to delimit control information within packet payload~\cite{Bonaventure:wx}.
+The channel is bootstrapped using an extended three-way handshake with fields shown in Figure~\ref{fig:header}:
 
@@
 \label{fig:handshake}
-\vspace{-4mm}
+% \vspace{-4mm}
 \end{figure}
@@
 
-We witnessed a case of packet splitting where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload.
+We witnessed one case of packet splitting where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload.
 We did not observe a case where a middlebox both passes payload in SYN-ACK packet and splits ACK segments.
@@
 
-%Covert channels are typically low bitrate in order to make it difficult to detect them.
-Application-specific ones, especially based on media transcoding can achieve rates from 270 b/s using $\mu TP$ torrent protocol, to 2 kbit/s by encoding data in silent Skype voice call periods, to megabits in wireless environments by inserting hidden data into the padding of frames~\cite{Zielinska:2014fn}.
-
 Our control channel does not need to protection from steganalysis as it is only used to avoid middlebox interference with end-to-end control information exchange.
-Bitrate of the channel only using packet headers (\ie without redefining part of packet payload for control information) is therefore bounded by the size of the TCP header (20 bytes without options), the round trip time (RTT) and the number of packets in flight per RTT.
+The bitrate of the channel without redefining part of packet payload for control information is bounded by the size of the TCP header (20 bytes without options), the round trip time (RTT) and the number of packets in flight per RTT.
 Depending on fields usable for a particular connection as inferred during the handshake, we can use between 16 (either $Opcode_C$ or $Opcode_U$) and 35 bits when using both fields and reserved flags.
 
-For any given operating window $cwnd$ (in bytes) a TCP flow has $cwnd/MSS$ data packets per RTT, hence that many headers with covert control data.
+For an operating window $cwnd$ (in bytes) a TCP flow has $cwnd/MSS$ data packets per RTT, hence that many headers can carry control data.
 If \emph{window scaling} option is not used, the maximum \emph{receive window} is 65536 bytes.
 Using the typical HSPA network RTT of about 100ms the maximum data rate is therefore 5 Mbit/s, reasonable for current networks~\cite{Chen:2012vj}.
-In this case covert channel bitrate is up to 7.0 to 15.4 kbit/s, depending on the fields used.
+In this case covert channel bitrate is 7.0 to 15.4 kbit/s, depending on the fields used.
 
-Otherwise, we can estimate channel bitrate from previously studied 3G and 4G network characteristics~\cite{Chen:2012vj}.
+We can estimate channel bitrate from previously studies of 3G and 4G network characteristics~\cite{Chen:2012vj}.
 The average bandwidth measured for a large TCP transmission ranges from $1.2\pm 0.2$ Mbit/s for a 3G network and $348\pm 141$ ms RTT to $15.6\pm 1.0$ Mbit/s and $125\pm 41$ ms RTT for LTE.
@@
 
-TCP implements an \emph{urgent mechanism} at its core that allows the sending user to make the receiving user to accept some data as \emph{urgent}.
-The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is explicitly not designed as a mechanism for sending such data.
-Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and the general recommendation is against the use of the mechanism~\cite{Gont:2011vi}.
+TCP implements an \emph{urgent mechanism} that allows the sender to make the receiver accept some \emph{urgent data}.
+The mechanism is often quoted as providing ``out-of-bound'' data delivery and may be compelling to use for control information even though it is explicitly not designed as such.
+There are ambiguities regarding the semantics of the urgent pointer and Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer.
+Hence the general recommendation is against the use of the mechanism~\cite{Gont:2011vi}.
 
-AccECN proposal~\cite{Kuhlewind:2014vd} uses an additional reserved bit, overloads the meaning of already assigned ECN~\cite{Floyd:up} and NS~\cite{Ely:uc} bits and redefines Urgent Pointer as Non-Urgent when URG is not set.
-The idea is highly related to ours, but reserving the field solely for ECN use prevents other extensions from using it.
-Furthermore, we saw that some networks block such packets, making it especially difficult to use for ECN.
+AccECN~\cite{Kuhlewind:2014vd} uses an additional reserved bit, overloads the meaning of already assigned ECN~\cite{Floyd:up} and NS~\cite{Ely:uc} bits and redefines Urgent Pointer as Non-Urgent when URG is not set.
+The idea is related to ours, but reserving the field solely for ECN use prevents other extensions from using it.
+Also, we saw that some networks block such packets, making it especially difficult to use for ECN.
 
@@
 
-\section{Discussion}
+\section{Discussion and Future Work}
 
@@
 We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.
-
 The exact protocol semantics are subject to further testing and discussions.
@@
 
-More importantly, we have left open the question of finding a method to hide such information from non-transparent proxies.
+We have left open the question of finding a method to hide such information from non-transparent proxies.
 The easiest, albeit limited solution is to identify and cache port numbers that allow unmodified traffic.
@@
 
-In the long term we believe that each protocol-related standard should explicitly define how it interferes with all header fields.
+In the long term we believe that each protocol-related standard should explicitly define how it affects header fields.
 Examples include implicitly extending sequence number space~\cite{Mazieres:uz}, allowing encrypted TCP session to omit the timestamp option, using stronger message integrity checks making checksum field redundant, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple window size modifications~\cite{Crowcroft:1998dk}.
@@
 Thus extending a protocol would become a matter of defining a new set of such header algebra rules hence simplifying the process of design and testing.
-
-\section{Conclusion}
-
-This research is still in early stages.
-The end goal is to develop a set of rules for how protocol extensions redefine and how middleboxes modify packet headers and use them to hide information so that middleboxes do not interfere with endpoint behavior.
-Development of such ruleset would mean that extending a protocol in the future could become near automatic.
 

### 2014-07-16 paper/paper.tex
first submitted version

@@
 
-TCP is everywhere in the Internet and it carries the overwhelming majority of bytes~\cite{John:2007dv}.
-TCP is used for all types of traffic from bulk download to video streaming to instant messaging.
 Over the last decade a number of new protocols have been proposed to replace TCP.
-However many middleboxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like.
-Consequently, few of these protocols have found any leverage.
+However many middleboxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like~\cite{Handley:2001vp,Vutukuru:2008fc,Guha2005tb}.
+Consequently, few of these protocols have seen real deployment.
 
@@
 Traditionally this is what TCP option space was used for.
-However option space is limited, heavily contended~\cite{Handley:vj,Mazieres:uz} and network middleboxes make it difficult to use~\cite{Honda,Guha2005tb,UntoldMiddlebox2011}.
+However option space is limited, heavily contended~\cite{Handley:vj,Mazieres:uz} and network middleboxes make it difficult to use because of inconsistent interpretation of the specifications~\cite{Honda,Guha2005tb,UntoldMiddlebox2011}.
 On the other hand, modern networks demand new protocols for mobile and multi-homed devices~\cite{Mazieres:uz}, ubiquitous traffic encryption~\cite{Handley:vj,Rescorla:R080FoGB}, energy efficiency~\cite{Aucinas:2013uk}, connection parameters tuning for specific workloads and environments~\cite{Dukkipati:2010hs,Erman:2013uz} that were unknown 30 years ago.
 
-We believe that the solution is to work round the middleboxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header.
-This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that SYN Cookies~\cite{Eddy:2007to} do currently.
+One solution is to work round the middleboxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header beyond the sight of middleboxes.
+This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that TCP Fast Open~\cite{Chu:2011tn} does for application data delivery.
 
-This paper makes three main contributions. 
-\begin{enumerate}
-\item We discuss the possible ways to conceal information in the TCP header and give some examples of the extensions this could enable
-\item We describe a measurement application that we have written to run on Android phones for a large-scale study.  
-\item We present the results of an initial small scale deployment of our application.
-\end{enumerate}
+Steganographic communication channels are typically low bandwidth and statistical analysis of a sufficient amount of data usually exposes the existence of such channels~\cite{Murdoch:2005fz}.
+Nevertheless, it is the perfect choice for exchanging a small amount of end-to-end control data in real-time and it does not compromise network security since the channel can be detected using existing techniques.
+In our work we discuss the possible ways to conceal information in the TCP header and present the results of a preliminary study on the suitability of various header fields for the purpose.
 
 We use our initial study to design a way of exchanging a small amount of control information successfully across all of our tested networks and show a potential mechanism to use it for bootstrapping a full control stream currently missing from TCP.
+
+Our approach is not a short-term hack.
+Many recent proposals have demonstrated that control over end-to-end protocol behavior should remain at the endpoints.
+These include TCPCrypt which offers improved communication confidentiality and security and TCP Fast Open for better communication performance.
 We hope that our successful results will motivate the research community into exploring the full range of possibilities in this space.
@@
 % \vspace{-5mm}
-\caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload}
+\caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload.
+Fields in gray used for control information.
+Light-gray fields are optional and are within payload.}
 \label{fig:header}
 }
-% \vspace{-4mm}
+\vspace{-4mm}
 \end{figure}
 
-Figure~\ref{fig:header} shows the TCP header with fields that we propose for reuse (defined later in the paper).
-The structure of the TCP header is very simple and gives few choices for hiding extra control information:
+Design of a control channel is constrained by protocol specification and its interpretation by middleboxes.
+In the past, TCP sequence numbers~\cite{Murdoch:2005fz}, timestamps~\cite{Giffin:2002wh} and IP ID~\cite{Ahsan:2002wc} have been used for encoding hidden information.
+In our work we focus on a different set of TCP header fields (Fig.~\ref{fig:header}, in gray, defined in \S\ref{sec:protocol}).
+Overall, the header leaves few options for hiding control information:
 \begin{itemize}
-    \item Setting header fields without the corresponding flags so they have no meaning to legacy implementations.
+    \item Setting fields without the corresponding flags so they have no meaning to legacy implementations.
     \item Assigning one or more of the reserved field bits.
@@
 Only two header fields have their semantics set by corresponding flags.
-These are the \emph{Acknowledgment Number} and \emph{Urgent Pointer} fields.
-The former must also always be set after the initial SYN packet.
+These are the \emph{Acknowledgment Number} (which must always be set after the initial SYN packet) and the \emph{Urgent Pointer}.
 
-The general recommendation is to not use the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics.
-On the other hand, Network Intrusion Detection Systems (NIDS) often reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
-Importantly, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and so can be redefined.
-
-The Acknowledgment Number field is only available in the initial SYN packet.
+The Acknowledgment Number field is only available in the initial SYN packet, but is usable across our tested networks (\S\ref{sec:network}).
 It must therefore be combined with another mechanism to acknowledge control information unless TCP Simultaneous Open is used.
 Simultaneous open is still viable in most cellular networks~\cite{UntoldMiddlebox2011}, but we also explore other options.
+
+The general recommendation is to not use the urgent mechanism~\cite{Gont:2011vi}.
+The \emph{Urgent Pointer} could therefore be reused for other purposes as long as both endpoints understand the new semantics and middleboxes permit it.
+On the other hand, Network Intrusion Detection Systems (NIDS) often reset the field as it makes it difficult to track application-layer data~\cite{seolma}.
+Importantly, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and so can be redefined if middleboxes permit it as they often do (\S\ref{sec:network}).
 
@@
 
-Three TCP Reserved bits remain unallocated.
+Three TCP reserved bits remain unallocated.
 One further bit is only allocated to the experimental ECN nonce\cite{Ely:uc}.
@@
 It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown to be unreliable~\cite{Stone:2000fc}.
-Instead, it may be replaced with potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code or left for the upper layer.
+Instead, it may be replaced with potentially stronger mechanisms such as MPTCP DSS checksum and TCPCrypt Message Authentication Code or left for the upper layer.
 Secondly, checksum operations are efficient as they only involve one's complement addition.
@@
 
-To differentiate between general behavior of particular network and application-specific optimizations (\eg HTTP acceleration) we repeat our tests on a few application-specific ports as well as a random one:
+To differentiate between general behavior of particular network and application-specific optimizations (\eg HTTP acceleration) we repeat our tests on a few application ports as well as a random one:
 \begin{itemize}
@@
 It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to different values to optimize TCP for cellular networks~\cite{Ha:2006td}.
-Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that TCPCrypt fails on this network.
+Such behavior makes deployment of protocol extensions difficult -- we verified that TCPCrypt fails on this network.
 We do not currently have an estimate of how widespread ByteMobile proxies are as we only observed them on one network.
@@
 Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize so hiding any information within the packet becomes difficult.
-We tested for them on a number of popular port numbers:
+
+We tested for them on a number of popular port numbers, from those use for HTTP and mail to typically encrypted ones:
 21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.
@@
 
-We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream.
-After the handshake a higher-bandwidth \emph{Control Channel} can be defined by using TLV (Type-Length-Value) encoding to delimit control information within packet payload~\cite{Bonaventure:wx}.
+Covert channel is only needed to ensure that both endpoints understand the same protocol semantics.
+Once the handshake with hidden control information is complete, any aspect of a protocol can be redefined.
+For example, a higher-bandwidth \emph{Control Channel} can use TLV (Type-Length-Value) encoding to delimit control information within packet payload~\cite{Bonaventure:wx}.
 The channel is bootstrapped using an extended three-way handshake with fields shown in Figure~\ref{fig:header}:
@@
 \begin{enumerate}
-\item \textbf{SYN}: the active opener sends SYN packet with \emph{$Opcode Namespace$ + $Opcode$} and ACK unset.
-\item \textbf{SYN-ACK}: if passive opener recognizes the namespace and the $Opcode$, it replies with SYN-ACK with $Opcode_U$ = $Opcode$ (URG unset) to acknowledge the Opcode.
-\item If passive opener does not receive an ACK, retransmit with $Opcode_C$ and \emph{Checksum Correction}.
+\item \textbf{SYN}: the active opener (client) sends SYN packet with \emph{$Opcode Namespace$ + $Opcode$} and ACK unset.
+\item \textbf{SYN-ACK}: if passive opener (server) recognizes the  $Namespace$ and the $Opcode$, it replies with SYN-ACK with $Opcode_U$ = $Opcode$ (URG unset) to acknowledge the Opcode.
+\item If the server does not receive an ACK, retransmit with $Opcode_C$ and \emph{Checksum Correction}.
 If that fails as well, retransmit the SYN-ACK with an invalid checksum.
 Finally, fall back to vanilla TCP and retransmit a standard SYN-ACK.
-\item \textbf{ACK}: when active opener receives a packet, it checks for $Opcode_U$ (if URG is unset).
+\item \textbf{ACK}: when the client receives a packet, it checks for $Opcode_U$ (if URG is unset).
 Otherwise, it checks if Checksum = $Opcode_C$ as described above.
 If no Opcode is recoverable, fall back to vanilla TCP and transmit a standard ACK.
-\item Otherwise both openers recognize the $Opcode$. Active opener sends an ACK packet with TLV-encoded control information in packet payload. The control stream has been successfully bootstrapped.
+\item Otherwise both openers recognize the $Opcode$.
+Active opener sends an ACK packet with control information in the field it received it on and potentially TLV-encoded control information in packet payload.
+The control stream has been successfully bootstrapped.
 \end{enumerate}
@@
 Figure~\ref{fig:handshake} shows an example handshake.
-In this case a client (active opener) sends a SYN packet as show above.
-The packet traverses a middlebox that does not remove the opcode and reaches the Server (passive opener).
-The Server replies with SYN-ACK with U\_Opcode, but the middlebox discards it.
-The server then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, delivered to the Client.
+In this case a client sends a SYN packet as described above.
+The packet traverses a middlebox that does not remove the opcode and reaches the server.
+The server replies with SYN-ACK with $Opcode_U$, but the middlebox discards it.
+The server then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, delivered to the client.
 The client recovers and verifies the received opcode.
-Hence both endpoints understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging reception of the $Opcode$ from the Server.
+Hence both endpoints understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging $Opcode$ reception.
 
@@
 There are ambiguities regarding the semantics of the urgent pointer and Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer.
-Hence the general recommendation is against the use of the mechanism~\cite{Gont:2011vi}.
+Hence the general recommendation is not to use of the mechanism~\cite{Gont:2011vi}.
 
@@
 
-In the long term we believe that each protocol-related standard should explicitly define how it affects header fields.
+In the long term each protocol-related standard should explicitly define how it affects header fields.
 Examples include implicitly extending sequence number space~\cite{Mazieres:uz}, allowing encrypted TCP session to omit the timestamp option, using stronger message integrity checks making checksum field redundant, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple window size modifications~\cite{Crowcroft:1998dk}.
@@
 We hope to map such header field modifications by existing extensions as well as network middlebox behavior patterns into an algebraic form as we have done in the simple case of checksum.
-This way we would not only be able to exchange control information across the network reliably, but also identify network path behavior by looking at its performed operations.
+Such definitions would allow us to go beyond the simple opcode exchange and signal detailed protocol information between endpoints using a low-bandwidth channel.
 Thus extending a protocol would become a matter of defining a new set of such header algebra rules hence simplifying the process of design and testing.
 
-More importantly, this paper hopes to motivate the research community to explore the full range of possibilities in this space as well as contribute to our work by downloading our research app and helping to build a comprehensive database of middlebox interference with protocols across different networks.
+We have demonstrated the feasibility of using steganography to exchange end-to-end control information across current networks.
+Importantly, this is not a short term solution: TCP Fast Open is already being deployed and offers even more opportunities to hide information in the initial packets within higher-level protocol information, such as HTTP headers and cipher suites~\cite{Zielinska:2014fn}.
+Even if TCP was completely replaced by another protocol, opportunities for control information exchange remain.
 
